---
title: "Fish traits"
output: github_document
editor_options: 
  chunk_output_type: console
---
#### R Markdown

## Load packages
- These are the packages that are needed to extract data from online repositories and manipulate the data
```{r load packages and files}
library(rfishbase) # data manipulation
library(tidyr) # data manipulation
library(dplyr) # data manipulation
library(stringr) # data manipulation
```

## Possible trait cateogries
- We won't use this but it is a helpful tool to check out to find categories of traits
```{r, eval = FALSE}
# fb_tables(server = "fishbase", version = "latest")
```

## Find certain traits
- Can be useful to see the variety of traits available from FishBase
```{r, eval = FALSE}
# # Use these to explore data tables for possible traits
# fb_tbl("spawnagg", server = "fishbase")
# 
# spawnagg <- rfishbase:::endpoint("spawnagg")
# spawnagg <- spawnagg(existent_species)
# 
# larvaepresence <- rfishbase:::endpoint("larvaepresence")
# larvaepresence <- larvaepresence(existent_species)
```

## Load in the species list
```{r, extract species pool traits from rfishbase}
existing_species <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/existing_species.csv")

existing_species <- existing_species[,-1]
```

## Extract traits for species
```{r}
## Species Traits
# Download traits for fish and drop any traits you do not want
trait <- species(existing_species)
#trait <- trait[, c(2:3, 15, 20:23, 38)]

## Morphological traits 
# Download traits for fish and drop any traits you do not want
morphology <- morphology(existing_species)

## Estimated traits
# Download traits for fish and drop any traits you do not want
# Trophic levels
estimate <- estimate(existing_species)

## Ecological traits 
# Download traits for fish and drop any traits you do not want
ecology <- ecology(existing_species)

## Reproduction traits
# Download traits for fish and drop any traits you do not want
reproduction <- reproduction(existing_species)

# Bind the four trait dataframes by columns, duplicated column names will be deleted with the second line of code.
traits <- full_join(trait, morphology, by = "Species")
traits <- full_join(traits, estimate, by = "Species")
traits <- full_join(traits, ecology, by = "Species")
traits <- full_join(traits, reproduction, by = "Species")

# traits <- cbind(trait, morphology, estimate, ecology, reproduction)
traits <- traits[,!duplicated(colnames(traits))]
```


## Completeness Function
- You will run a function written by Gio Rapacciuolo that will search out traits that are a certain percentage complete for all species. #From trait-beta-div-functions.R
```{r, completeness function}
##### -- Check for the completeness of vectors -- #####
completeness <- function(vec){
  if (is.list(vec)) return(apply(vec, 2, function(x) (sum(complete.cases(x))/length(x)) * 100))
  else return((sum(complete.cases(vec))/length(vec)) * 100)
}
```

## Filter for 90% complete traits
- Now isolate for completeness of traits because you need to have traits that are recorded for most species to look at the diversity. This will get rid of traits that are lacking sufficient data.
```{r, Isolate the most useful traits that have information}
# Isolate the most complete traits (completeness > 75% - 90%)
# complete_traits_75 <- traits[, which(completeness(traits) > 75)]
# 
# complete_traits_80 <- traits[, which(completeness(traits) > 80)]
# 
# complete_traits_85 <- traits[, which(completeness(traits) > 85)]

complete_traits_90 <- traits[, which(completeness(traits) > 90)]

# check <- complete_traits_75[setdiff(names(complete_traits_75), names(complete_traits_90))]

# Isolate traits that include variation
# complete_traits_75 <- complete_traits_75[, which(apply(complete_traits_75, 2, function(x) length(unique(na.omit(x)))) > 1)]
# 
# complete_traits_80 <- complete_traits_80[, which(apply(complete_traits_80, 2, function(x) length(unique(na.omit(x)))) > 1)]
# 
# complete_traits_85 <- complete_traits_85[, which(apply(complete_traits_85, 2, function(x) length(unique(na.omit(x)))) > 1)]

complete_traits_90 <- complete_traits_90[, which(apply(complete_traits_90, 2, function(x) length(unique(na.omit(x)))) > 1)]
```

## Select specific traits
- Edit how traits are viewed by R. Also fishbase records for environmental data presence is -1 however, programs have trouble with that so we need to convert all those -1 to 1 so that they can be better dealt with in R.
```{r, edit how data is viewed and change columns with -1 to 1}
complete_traits <- complete_traits_90

# # Calculate the estimated Weight with the following columns.
# complete_traits$Weight <- complete_traits$a * complete_traits$MaxLengthTL^complete_traits$b
# 
# # Calculate the estimated caudal fin length with the following columns.
# complete_traits$CaudalFinLength <- complete_traits$MaxLengthTL - complete_traits$MaxLengthSL

# Get rid of unnecessary or irrelevant traits
keep <- c("Species", "Genus", "BodyShapeI.x", "Fresh", "Brack", "Saltwater", "DemersPelag", "OperculumPresent", "DorsalSpinesMax", "MaxLengthTL", "Troph", "DepthMin", "DepthMax", "TempPrefMin", "TempPrefMax", "FeedingPath")
complete_traits <- complete_traits[,keep]
# Traits we are looking to keep:
# Species	Genus	BodyShapeI.x Fresh Brack Saltwater	DemersPelag	OperculumPresent DorsalSpinesMax MaxLengthTL Troph	DepthMin DepthMax TempPrefMin TempPrefMax	FeedingPath

# complete_traits <- complete_traits[, c(1,3,10,13:16,21,55,57,59:62,68,70:71,73:76,83,84,87:91,93,95,97,98)]
# Traits we are looking to keep:
# Species	Family	Genus	BodyShapeI.x	Fresh	Brack	Saltwater	DemersPelag	Vulnerability	OperculumPresent	DorsalSpinesMin	DorsalSpinesMax	DorsalSoftRaysMin	DorsalSoftRaysMax	Araymin	Araymax	MaxLengthTL	Troph	seTroph	a	sd_log10a	b	sd_b	DepthMin	DepthMax	PredPreyRatioMin	PredPreyRatioMax	TempPrefMin	TempPrefMean	TempPrefMax	FeedingPath	MaxLengthSL	Weight	CaudalFinLength	RepGuild1	RepGuild2	ParentalCare

# Make -1 in 1 by multiplication
complete_traits$OperculumPresent <- complete_traits$OperculumPresent*-1

# RepGuild1	RepGuild2	ParentalCare
keep <- c("Species", "RepGuild1", "RepGuild2", "ParentalCare")
repro <- reproduction[,keep]

# DorsalSpinesMin
keep <- c("Species", "DorsalSpinesMin")
dorsal <- morphology[,keep]

complete_traits <- full_join(complete_traits, repro, by = "Species")
complete_traits <- full_join(complete_traits, dorsal, by = "Species")

# Make integer fields numeric
numerical <- c("DorsalSpinesMax", "MaxLengthTL", "Troph", "DepthMin", "DepthMax", "TempPrefMin", "TempPrefMax", "DorsalSpinesMin")
complete_traits[,numerical] <- sapply(complete_traits[,numerical], as.numeric)

# Make character fields into factors to use in trait analyses
factor <- c("BodyShapeI.x", "Fresh", "Brack", "Saltwater", "DemersPelag", "OperculumPresent", "FeedingPath", "RepGuild1", "RepGuild2", "ParentalCare")
# Convert to lowercase and then to factors
complete_traits[,factor] <- lapply(complete_traits[,factor], function(x) as.factor(tolower(trimws(x))))

# Identify which rows have NAs and will be filled in with Genus level data
fromGenus <- apply(complete_traits, 1, function(x){any(is.na(x))})
sum(fromGenus) 

complete_traits <-  cbind(complete_traits, fromGenus)
```

## Find and fill in missing data using genera averages
- Next we are going to replace an NA in species trait data information with a summary of trait from all species within that genus. That way we can make sure we are not missing information from species. We do much of what we did above but this is now information only on species that are lacking information. So we download all the data for the species with that Genus and then summarize it to give an estimate to the missing results.
```{r, message=FALSE}
# Run to check for duplicates
duplicates_check <- which(duplicated(complete_traits$Species))
duplicates_check

# Add genera of species with incomplete data
# Last time I went through this, 1083 species, plus the 48 species only identified to genera had NAs and needed averaged data from the genus.
species_incomplete_genus <- complete_traits %>% dplyr::filter(!complete.cases(.)) %>% dplyr::select(Genus)
species_incomplete_genus <- unique(species_incomplete_genus$Genus)

## Download fishbase data from all species in those additional genera
# Generate the list of all species from those genera
species_incomplete_genus <- unlist(lapply(species_incomplete_genus, function(x) species_list(Genus = x)))

## Extract traits from fishbase
# Main traits
trait_incomplete_genus <- species(species_incomplete_genus)

## Morphology
morphology_incomplete_genus <- morphology(species_incomplete_genus)
# Used to determine duplicates of Species in a column
duplicates_check <- which(duplicated(morphology_incomplete_genus$Species))
duplicates_check 
# Removed duplicates
morphology_incomplete_genus <- morphology_incomplete_genus %>%
  distinct(Species, .keep_all = TRUE)

# Estimates
estimate_incomplete_genus <- estimate(species_incomplete_genus)

## Ecology
ecology_incomplete_genus <- ecology(species_incomplete_genus)
# Used to determine duplicates of Species in a column
duplicates_check <- which(duplicated(ecology_incomplete_genus$Species))
duplicates_check 
# Removed duplicates
ecology_incomplete_genus <- ecology_incomplete_genus %>%
  distinct(Species, .keep_all = TRUE)

# Reproduction
reproduction_incomplete_genus <- reproduction(species_incomplete_genus)

# Merge the five data.frames by columns
traits_incomplete_genus <- full_join(trait_incomplete_genus, morphology_incomplete_genus, by = "Species")
traits_incomplete_genus <- full_join(traits_incomplete_genus, estimate_incomplete_genus, by = "Species")
traits_incomplete_genus <- full_join(traits_incomplete_genus, ecology_incomplete_genus, by = "Species")
traits_incomplete_genus <- full_join(traits_incomplete_genus, reproduction_incomplete_genus, by = "Species")

# Get rid of duplicated columns
traits_incomplete_genus <- traits_incomplete_genus[,!duplicated(colnames(traits_incomplete_genus))]

# # Calculate the estimated Weight with the following columns.
# traits_incomplete_genus$Weight <- traits_incomplete_genus$a * traits_incomplete_genus$MaxLengthTL^traits_incomplete_genus$b
# 
# # Calculate the estimated caudal fin length with the following columns.
# traits_incomplete_genus$CaudalFinLength <- traits_incomplete_genus$MaxLengthTL - traits_incomplete_genus$MaxLengthSL

# Make -1 in 1 by multiplication
traits_incomplete_genus$OperculumPresent <- traits_incomplete_genus$OperculumPresent*-1

# Filter the fields of interest
traits_incomplete_genus <- traits_incomplete_genus[names(complete_traits)[-ncol(complete_traits)]]

# Check columns to see data types
str(traits_incomplete_genus)

## Calculate genus-level summary traits
# Split species by genera
traits_incomplete_genus_list <- split(traits_incomplete_genus, as.factor(traits_incomplete_genus$Genus))

# Summarize values across all species in the genus. Take the most common character for character traits and take the mean value for numerical traits.
traits_incomplete_genus_summary <- lapply(traits_incomplete_genus_list, function(x){
  c(
  unlist(c(
    sapply(x[,factor], function(y){
      out_y <- table(y)
      ifelse(length(out_y) > 0, names(which.max(out_y)), NA)
    }))),    
  apply(x[,numerical], 2, mean, na.rm=T)
    )
})

# Bind rows into data.frame
complete_traits_incomplete_genus_summary <- data.frame(do.call("rbind", traits_incomplete_genus_summary))

## Check order of columns because they most likely shifted
str(complete_traits_incomplete_genus_summary)

# Make integer fields numeric
complete_traits_incomplete_genus_summary[,numerical] <- apply(complete_traits_incomplete_genus_summary[,numerical], 2, as.numeric)

# Edit fields
complete_traits_incomplete_genus_summary[,factor] <- lapply(complete_traits_incomplete_genus_summary[,factor], function(x) as.factor(tolower(trimws(x))))

# Remake Genus column
complete_traits_incomplete_genus_summary$Genus <- row.names(complete_traits_incomplete_genus_summary)

# Make new file just in case code doesn't work so you don't have to go back
complete_imputed <- complete_traits

complete_imputed$BodyShapeI.x[is.na(complete_imputed$BodyShapeI.x)] <- complete_traits_incomplete_genus_summary$BodyShapeI.x[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$BodyShapeI.x))]

complete_imputed$Fresh[is.na(complete_imputed$Fresh)] <- complete_traits_incomplete_genus_summary$Fresh[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$Fresh))]

complete_imputed$Brack[is.na(complete_imputed$Brack)] <- complete_traits_incomplete_genus_summary$Brack[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$Brack))]

complete_imputed$Saltwater[is.na(complete_imputed$Saltwater)] <- complete_traits_incomplete_genus_summary$Saltwater[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$Saltwater))]

complete_imputed$DemersPelag[is.na(complete_imputed$DemersPelag)] <- complete_traits_incomplete_genus_summary$DemersPelag[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$DemersPelag))]

complete_imputed$OperculumPresent[is.na(complete_imputed$OperculumPresent)] <- complete_traits_incomplete_genus_summary$OperculumPresent[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$OperculumPresent))]

complete_imputed$DorsalSpinesMin[is.na(complete_imputed$DorsalSpinesMin)] <- complete_traits_incomplete_genus_summary$DorsalSpinesMin[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$DorsalSpinesMin))]

complete_imputed$DorsalSpinesMax[is.na(complete_imputed$DorsalSpinesMax)] <- complete_traits_incomplete_genus_summary$DorsalSpinesMax[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$DorsalSpinesMax))]

complete_imputed$MaxLengthTL[is.na(complete_imputed$MaxLengthTL)] <- complete_traits_incomplete_genus_summary$MaxLengthTL[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$MaxLengthTL))]

complete_imputed$Troph[is.na(complete_imputed$Troph)] <- complete_traits_incomplete_genus_summary$Troph[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$Troph))]

complete_imputed$DepthMin[is.na(complete_imputed$DepthMin)] <- complete_traits_incomplete_genus_summary$DepthMin[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$DepthMin))]

complete_imputed$DepthMax[is.na(complete_imputed$DepthMax)] <- complete_traits_incomplete_genus_summary$DepthMax[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$DepthMax))]

complete_imputed$TempPrefMin[is.na(complete_imputed$TempPrefMin)] <- complete_traits_incomplete_genus_summary$TempPrefMin[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$TempPrefMin))]

complete_imputed$TempPrefMax[is.na(complete_imputed$TempPrefMax)] <- complete_traits_incomplete_genus_summary$TempPrefMax[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$TempPrefMax))]

complete_imputed$FeedingPath[is.na(complete_imputed$FeedingPath)] <- complete_traits_incomplete_genus_summary$FeedingPath[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$FeedingPath))]

complete_imputed$RepGuild1[is.na(complete_imputed$RepGuild1)] <- complete_traits_incomplete_genus_summary$RepGuild1[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$RepGuild1))]

complete_imputed$RepGuild2[is.na(complete_imputed$RepGuild2)] <- complete_traits_incomplete_genus_summary$RepGuild2[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$RepGuild2))]

complete_imputed$ParentalCare[is.na(complete_imputed$ParentalCare)] <- complete_traits_incomplete_genus_summary$ParentalCare[match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)][which(is.na(complete_imputed$ParentalCare))]

# Replace "x" with NA in specified numeric columns
complete_imputed <- complete_imputed %>%
  mutate_at(vars(numerical), ~ ifelse(. == "NaN", NA, .))

## Fill in NAs in trait data set "complete_imputed" with genus-level summaries
# complete_imputed <- complete_imputed %>%
#   mutate(across(everything(), ~ ifelse(is.na(.), complete_traits_incomplete_genus_summary[[cur_column()]][match(complete_imputed$Genus, complete_traits_incomplete_genus_summary$Genus)], .)))

str(complete_imputed)
```

## Trait data for incomplete species
- For species that were only identified to the genus taxonomic level we will use the average trait data for the genus to fill in the missing trait data
```{r,  message=FALSE}
#### Add genera from species not in fishbase
#### Use imcomplete_to_add file from fish_species_data.Rmd
### For species not included in fishbase, extract all species in the same genus
## Extract genus names
# Extract out the species that are not present on FishBase
# missing_species <- setdiff(existing_species, existent_species)
# missing_species
incomplete_to_add <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/incomplete_to_add.csv")

missing_species <- incomplete_to_add[,-1]

# Delete anything after the space to get the genus level for each species
missing_genus <- str_extract(missing_species, "[^ ]+")
missing_genus

# Remove duplicated genera
missing_genus <- missing_genus[!duplicated(missing_genus)]

# Extract all species of the genera in the vector
species_missing_genus_added <- unlist(lapply(missing_genus, function(x) species_list(Genus = x)))

### Download fishbase data from all species in those additional genera
## Extract traits from fishbase
# Main traits
trait_missing_genus_added <- species(species_missing_genus_added)

# Morphology
morphology_missing_genus_added <- morphology(species_missing_genus_added)
# Check if there are duplicated morphologies
duplicates_check <- which(duplicated(morphology_missing_genus_added$Species))
duplicates_check
# Deleted duplicates
morphology_missing_genus_added <- morphology_missing_genus_added %>%
  distinct(Species, .keep_all = TRUE)

# Estimates
estimate_missing_genus_added <- estimate(species_missing_genus_added)

## Ecology
ecology_missing_genus_added <- ecology(species_missing_genus_added)

# Reproduction
reproduction_missing_genus_added <- reproduction(species_missing_genus_added)

# Merge the five trait data.frames
traits_missing_genus_added <- full_join(trait_missing_genus_added, morphology_missing_genus_added, by = "Species")
traits_missing_genus_added <- full_join(traits_missing_genus_added, estimate_missing_genus_added, by = "Species")
traits_missing_genus_added <- full_join(traits_missing_genus_added, ecology_missing_genus_added, by = "Species")
traits_missing_genus_added <- full_join(traits_missing_genus_added, reproduction_missing_genus_added, by = "Species")

# Get rid of duplicated columns
traits_missing_genus_added <- traits_missing_genus_added[,!duplicated(colnames(traits_missing_genus_added))]

# # Calculate the estimated Weight with the following columns.
# traits_missing_genus_added$Weight <- traits_missing_genus_added$a * traits_missing_genus_added$MaxLengthTL^traits_missing_genus_added$b
# 
# # Calculate the estimated caudal fin length with the following columns.
# traits_missing_genus_added$CaudalFinLength <- traits_missing_genus_added$MaxLengthTL - traits_missing_genus_added$MaxLengthSL

# Make -1 in 1 by multiplication
traits_missing_genus_added$OperculumPresent <- traits_missing_genus_added$OperculumPresent*-1

## Filter the fields of interest
traits_missing_genus_added <- traits_missing_genus_added[names(complete_traits)[-ncol(complete_traits)]]

# Check columns to see data types
str(traits_missing_genus_added)

#### Calculate genus-level summary traits
### Split species by genera
traits_missing_genus_added_list <- split(traits_missing_genus_added, as.factor(traits_missing_genus_added$Genus))

### Summarize values across all species in the genus
traits_missing_genus_added_summary <- lapply(traits_missing_genus_added_list, function(x){
  c(
  unlist(c(
    sapply(x[,factor], function(y){
      out_y <- table(y)
      ifelse(length(out_y) > 0, names(which.max(out_y)), NA)
    }))),    
  apply(x[,numerical], 2, mean, na.rm=T)
    )
})

### Bind rows into data.frame
complete_traits_missing_genus_added_summary <- data.frame(do.call("rbind", traits_missing_genus_added_summary))

# Make integer fields numeric
complete_traits_missing_genus_added_summary[,numerical] <- sapply(complete_traits_missing_genus_added_summary[,numerical], as.numeric)

# Edit fields
complete_traits_missing_genus_added_summary[,factor] <- lapply(complete_traits_missing_genus_added_summary[,factor], function(x) as.factor(tolower(trimws(x))))

## Check order of columns because they most likely shifted
str(complete_traits_missing_genus_added_summary)

# Make Genus column
complete_traits_missing_genus_added_summary$Genus <- row.names(complete_traits_missing_genus_added_summary)
complete_traits_missing_genus_added_summary <- complete_traits_missing_genus_added_summary[!duplicated(complete_traits_missing_genus_added_summary[c('Genus')]), ]

# Duplicate the missing Genus rows that were deleted earlier
complete_traits_missing_genus_added_summary <- rbind(complete_traits_missing_genus_added_summary, complete_traits_missing_genus_added_summary[c("Callogobius", "Chromis", "Eviota", "Eviota", "Gobiodon", "Liopropoma", "Trimma", "Xenisthmus"), ])

# Re-order rows to organize duplicated rows
complete_traits_missing_genus_added_summary <- complete_traits_missing_genus_added_summary[order(row.names(complete_traits_missing_genus_added_summary)), ]

### Add species and genus names
complete_traits_missing_genus_added_summary$Species <- sort(missing_species)
complete_traits_missing_genus_added_summary$fromGenus <- TRUE

### Add final species to trait data.frame 
complete_imputed <- data.frame(rbind(complete_imputed, complete_traits_missing_genus_added_summary))

# Identify which rows have NAs and will be filled in with Genus level data
fromFamily <- apply(complete_imputed, 1, function(x){any(is.na(x))})
sum(fromFamily)

complete_imputed_genus <-  cbind(complete_imputed, fromFamily)
```

## Higher taxonomic classifications
- We need higher taxonomic information for each species to fill in missing data using family trait averages.
```{r, generate list of missing data species and acquire data for Family}
# Added from Family field
#complete_imputed_genus$fromFamily <- 0
genera <- complete_imputed_genus$Genus

# Download taxonomic information from FishBase
genera <- rfishbase::load_taxa() %>% 
  filter(Genus %in% genera) %>%
  collect()

# Choose which classification levels you want to keep
keep <- c("Genus", "Family", "Order", "Class")
genera <- genera[,keep]

# Merge dataframes together
complete_imputed_genus <- merge(genera, complete_imputed_genus, by = "Genus", all.x = F, all.y = F)

# Remove duplicates in dataframe
complete_imputed_genus <- complete_imputed_genus[!duplicated(complete_imputed_genus[c('Species')]), ]

# Move family column to the first column of the dataframe
complete_imputed_genus <- complete_imputed_genus %>% relocate(Family, .before = Genus) %>% 
  relocate(Order, .before = Family) %>% 
  relocate(Class, .before = Order)
```

## Find and fill in missing data using family averages
- Next we are going to replace an NA in species trait data information with a summary of trait from all species within that family. That way we can make sure we are not missing information from species. We do much of what we did above but this is now information only on species that are lacking information after the Genus level summary. So we download all the data for the species with that Family and then summarize it to give an estimate to the missing results.
```{r, message=FALSE}
# Run to check for duplicates
duplicates_check <- which(duplicated(complete_imputed_genus$Species))
duplicates_check

# Add genera of species with incomplete data
# Last time I went through this 724 species had NAs and needed data from the family level
species_incomplete_family <- complete_imputed_genus %>% dplyr::filter(!complete.cases(.)) %>% dplyr::select(Family)
species_incomplete_family <- unique(species_incomplete_family$Family)

## Download fishbase data from all species in those additional genera
# Generate the list of all species from those genera
species_incomplete_family <- unlist(lapply(species_incomplete_family, function(x) species_list(Family = x)))

# Get rid of everything after the space to get genera only
species_names_family <- str_extract(species_incomplete_family, "[^ ]+")

# Download taxonomic information from FishBase
species_names_family <- rfishbase::load_taxa() %>% 
  filter(Genus %in% species_names_family) %>%
  collect()

# Choose which classification levels you want to keep
keep <- c("Species", "Family")
species_names_family <- species_names_family[,keep]

## Extract traits from fishbase
# Main traits
trait_incomplete_family <- species(species_incomplete_family)

## Morphology
morphology_incomplete_family <- morphology(species_incomplete_family)
# Used to determine duplicates of Species in a column
duplicates_check <- which(duplicated(morphology_incomplete_family$Species))
duplicates_check 
# Removed duplicates
morphology_incomplete_family <- morphology_incomplete_family %>%
  distinct(Species, .keep_all = TRUE)

# Estimates
estimate_incomplete_family <- estimate(species_incomplete_family)

## Ecology
ecology_incomplete_family <- ecology(species_incomplete_family)
# Used to determine duplicates of Species in a column
duplicates_check <- which(duplicated(ecology_incomplete_family$Species))
duplicates_check 
# Removed duplicates
ecology_incomplete_family <- ecology_incomplete_family %>%
  distinct(Species, .keep_all = TRUE)

# Reproduction
reproduction_incomplete_family <- reproduction(species_incomplete_family)

# Merge the five data.frames by columns
traits_incomplete_family <- full_join(trait_incomplete_family, morphology_incomplete_family, by = "Species")
traits_incomplete_family <- full_join(traits_incomplete_family, estimate_incomplete_family, by = "Species")
traits_incomplete_family <- full_join(traits_incomplete_family, ecology_incomplete_family, by = "Species")
traits_incomplete_family <- full_join(traits_incomplete_family, reproduction_incomplete_family, by = "Species")

# Get rid of duplicated columns
traits_incomplete_family <- traits_incomplete_family[,!duplicated(colnames(traits_incomplete_family))]

# # Calculate the estimated Weight with the following columns.
# traits_incomplete_family$Weight <- traits_incomplete_family$a * traits_incomplete_family$MaxLengthTL^traits_incomplete_family$b
# 
# # Calculate the estimated caudal fin length with the following columns.
# traits_incomplete_family$CaudalFinLength <- traits_incomplete_family$MaxLengthTL - traits_incomplete_family$MaxLengthSL

# Make -1 in 1 by multiplication
traits_incomplete_family$OperculumPresent <- traits_incomplete_family$OperculumPresent*-1

# Filter the fields of interest
traits_incomplete_family <- traits_incomplete_family[names(complete_traits)[-ncol(complete_traits)]]

# Check columns to see data types
str(traits_incomplete_family)

# Add the family column into the dataframe
traits_incomplete_family <- merge(species_names_family, traits_incomplete_family, by = "Species", all.x = F)

## Calculate Family-level summary traits
# Split species by genera
traits_incomplete_family_list <- split(traits_incomplete_family, as.factor(traits_incomplete_family$Family))

# Summarize values across all species in the Family. Take the most common character for character traits and take the mean value for numerical traits.
traits_incomplete_family_summary <- lapply(traits_incomplete_family_list, function(x){
  c(
  unlist(c(
    sapply(x[,factor], function(y){
      out_y <- table(y)
      ifelse(length(out_y) > 0, names(which.max(out_y)), NA)
    }))),    
  apply(x[,numerical], 2, mean, na.rm=T)
    )
})

# Bind rows into data.frame
complete_traits_incomplete_family_summary <- data.frame(do.call("rbind", traits_incomplete_family_summary))

## Check order of columns because they most likely shifted
str(complete_traits_incomplete_family_summary)

# Make integer fields numeric
complete_traits_incomplete_family_summary[,numerical] <- apply(complete_traits_incomplete_family_summary[,numerical], 2, as.numeric)

# Edit fields
complete_traits_incomplete_family_summary[,factor] <- lapply(complete_traits_incomplete_family_summary[,factor], function(x) as.factor(tolower(trimws(x))))

# Remake Family column
complete_traits_incomplete_family_summary$Family <- row.names(complete_traits_incomplete_family_summary)

# Make new file just in case code doesn't work so you don't have to go back
complete_imputed_family <- complete_imputed_genus

complete_imputed_family$BodyShapeI.x[is.na(complete_imputed_family$BodyShapeI.x)] <- complete_traits_incomplete_family_summary$BodyShapeI.x[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$BodyShapeI.x))]

complete_imputed_family$Fresh[is.na(complete_imputed_family$Fresh)] <- complete_traits_incomplete_family_summary$Fresh[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$Fresh))]

complete_imputed_family$Brack[is.na(complete_imputed_family$Brack)] <- complete_traits_incomplete_family_summary$Brack[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$Brack))]

complete_imputed_family$Saltwater[is.na(complete_imputed_family$Saltwater)] <- complete_traits_incomplete_family_summary$Saltwater[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$Saltwater))]

complete_imputed_family$DemersPelag[is.na(complete_imputed_family$DemersPelag)] <- complete_traits_incomplete_family_summary$DemersPelag[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$DemersPelag))]

complete_imputed_family$OperculumPresent[is.na(complete_imputed_family$OperculumPresent)] <- complete_traits_incomplete_family_summary$OperculumPresent[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$OperculumPresent))]

complete_imputed_family$DorsalSpinesMin[is.na(complete_imputed_family$DorsalSpinesMin)] <- complete_traits_incomplete_family_summary$DorsalSpinesMin[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$DorsalSpinesMin))]

complete_imputed_family$DorsalSpinesMax[is.na(complete_imputed_family$DorsalSpinesMax)] <- complete_traits_incomplete_family_summary$DorsalSpinesMax[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$DorsalSpinesMax))]

complete_imputed_family$MaxLengthTL[is.na(complete_imputed_family$MaxLengthTL)] <- complete_traits_incomplete_family_summary$MaxLengthTL[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$MaxLengthTL))]

complete_imputed_family$Troph[is.na(complete_imputed_family$Troph)] <- complete_traits_incomplete_family_summary$Troph[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$Troph))]

complete_imputed_family$DepthMin[is.na(complete_imputed_family$DepthMin)] <- complete_traits_incomplete_family_summary$DepthMin[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$DepthMin))]

complete_imputed_family$DepthMax[is.na(complete_imputed_family$DepthMax)] <- complete_traits_incomplete_family_summary$DepthMax[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$DepthMax))]

complete_imputed_family$TempPrefMin[is.na(complete_imputed_family$TempPrefMin)] <- complete_traits_incomplete_family_summary$TempPrefMin[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$TempPrefMin))]

complete_imputed_family$TempPrefMax[is.na(complete_imputed_family$TempPrefMax)] <- complete_traits_incomplete_family_summary$TempPrefMax[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$TempPrefMax))]

complete_imputed_family$FeedingPath[is.na(complete_imputed_family$FeedingPath)] <- complete_traits_incomplete_family_summary$FeedingPath[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$FeedingPath))]

complete_imputed_family$RepGuild1[is.na(complete_imputed_family$RepGuild1)] <- complete_traits_incomplete_family_summary$RepGuild1[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$RepGuild1))]

complete_imputed_family$RepGuild2[is.na(complete_imputed_family$RepGuild2)] <- complete_traits_incomplete_family_summary$RepGuild2[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$RepGuild2))]

complete_imputed_family$ParentalCare[is.na(complete_imputed_family$ParentalCare)] <- complete_traits_incomplete_family_summary$ParentalCare[match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Family)][which(is.na(complete_imputed_family$ParentalCare))]

## Fill in NAs in trait data set "complete_imputed" with genus-level summaries
# complete_imputed_family <- complete_imputed_family %>%
#   mutate(across(everything(), ~ ifelse(is.na(.), complete_traits_incomplete_family_summary[[cur_column()]][match(complete_imputed_family$Family, complete_traits_incomplete_family_summary$Genus)], .)))
```

## Last file checks
- Edit rows and fields to make sure that final complete_imputed file has the necessary format for analyses. Make sure certain variables are numeric and others are factors. You can run the str command at the end to be sure everything is in the correct format.
```{r}
## Final editing steps
# Change species names that were altered by FishBase to the subspecies level because we only want species level information 
# complete$Species[complete$Species=="Platybelone argalus platyura"] <- "Platybelone argalus"
# complete$Species[complete$Species=="Tylosurus acus melanotus"] <- "Tylosurus melanotus"

complete <- complete_imputed_family

# Re-order rows
complete <- complete[order(complete$Species), ]

# Erase rownmaes since they need to be updated
rownames(complete) <- NULL

duplicates_check <- which(duplicated(complete$Species))
duplicates_check

### Rename rows
complete$Species <- gsub(" ", "_", complete$Species)
row.names(complete) <- complete$Species
# complete <- complete[,-1]

# Isolate the most complete traits (completeness > 99.9%), you will lose traits here
complete_best_traits <- complete[, which(completeness(complete) > 99.9)]

### Select all traits with more than 98% coverage, you will lose species and traits here
complete_best_species <- complete[, which(completeness(complete) >= 98)] %>% dplyr::filter(complete.cases(.))

# Check your final species count with the species list you started with
complete_species <- complete[,5]
existing_species <- gsub(" ", "_", existing_species)
diff_check <- setdiff(existing_species, complete_species)
diff_check
```

## Load out files
```{r, Fish traits file loadouts and session info}
## Load out files
# Traits for fish that are present in the marine lakes and ocean populations
# Has all or most species but loses traits. Species with 0 or very few NAs
write.csv(complete_best_traits,  "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/traits/complete_best_all_species_gnathostomata.csv")

# Has all or most traits but loses species. Traits with 0 or very few NAs
write.csv(complete_best_species,  "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/traits/complete_best_all_traits_gnathostomata.csv")

#Traits for fish that are present in Palau
# Traits with more NAs
# Edit species names, replace space with an underscore
write.csv(complete, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/traits/complete_traits_gnathostomata.csv")

sessionInfo()
```

## Post R processing
- Before proceeding to analyses with your trait file complete_traits_gnathostomata.csv you need to do some additional work with the dorsal spines and reproductive traits. Review the current traits NA values and use FishBase to see if the family page for the species contains any information you can use to fill in these NA values. Sometimes the R package rfishbase is not able to see some of the family information on the website. You find the family page for species by searching the species name (https://www.fishbase.us/) and then clicking the family name under the "Classification / Names" section on the page which should be the first section. After using FishBase to complete the dorsal spines and reproductive information your file will be named complete_traits_gnathostomata_with_FishBase_corrections.csv

- Then you use the book Fish Reproduction by Rocha et al. (2008) and use chapter 9, Reproductive Strategies of Fish by Patzner, R. A. to fill in some of the blanks of the RepGuild2 column. You will only use families where the RepGuild2 is the same for all species and fits with the categorization in our table already. Mainly the "open water/substratum egg scatterers". You will then name your file complete_traits_gnathostomata_with_Patzner_2008_data.csv

- After that you can determine the traits that you would like to include in your analyses which may be all or a subset. For this project we chose a subset that have quality definitions according to FishBase and are not repetitive traits. We named our final file final_traits_gnathostomata.csv which we use in all downstream analyses.
