---
title: "Phylogenetic diversity analyses"
output: github_document
editor_options: 
  chunk_output_type: console
---
### Load packages and files
```{r, Load packages and files}
# Load the knitr package if not already loaded
library(knitr)

# Source the R Markdown file
knit("/Users/bailey/Documents/research/fish_biodiversity/src/collection/load_collection_data.Rmd", output = "/Users/bailey/Documents/research/fish_biodiversity/src/collection/load_collection_data.md")

# Packages
library(picante)
library(vegan)
library(phytools)
library(ggplot2)
library(ggrepel)
library(caret)
library(BAT)
library(pairwiseAdonis)
library(emmeans)

# Site vectors
surveyed_sites <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOM", "OOO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_LCN <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOM", "OOO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_TLN_HLM <- c("BCM", "CLM", "FLK", "GLK", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOM", "OOO", "OTM", "RCA", "SLN", "ULN")

surveyed_sites_env <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OTM", "RCA", "SLN", "TLN", "ULN")

ocean_mixed_sites <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA", "FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

ocean_mixed_sites_env <- c("IBK", "NCN", "RCA", "FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

ocean_stratified_sites <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA", "BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

ocean_stratified_sites_env <- c("IBK", "NCN", "RCA", "BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

mixed_stratified_lakes <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "LLN", "MLN", "NLK", "NLN", "NLU", "OLO", "OTM", "SLN", "TLN", "ULN")

ocean_sites <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA")

ocean_sites_env <- c("IBK", "NCN", "RCA")

mixed_lakes <- c("FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

stratified_lakes <- c("BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

# Define your custom colors
custom_colors <- c("Reference" = "black", "Ocean" = "#EE6363", "Mixed" = "#87CEFA", "Stratified" = "#6E8B3D")

env_cont <- "orange"

set.seed(123)
```

### Edit input files
```{r, edit files}
env <- env[order(env$X),]
presabs_lake <- presabs_lake[order(row.names(presabs_lake)),]
surveyed_sites_lake <- surveyed_sites_lake[order(row.names(surveyed_sites_lake)),]

Site_type_group_ref <- env[,"Site_type"]
Site_type_group <- env[surveyed_sites,"Site_type"]
```

# Phylogenetic Diversity
## PD alpha Diversity
### PD alpha mntd, mpd, and pd calculations of phylogeny and files to be saved
```{r, PD alpha calculations, eval=F}
## Reference
# Dispersion
tree_disp <- dispersion(comm = presabs_lake, tree = tree, abund = F)
write.csv(tree_disp, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_disp.csv")

#Mean pairwise distances
tree_sesmpd <- ses.mpd(presabs_lake, cophenetic(tree), null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999, iterations = 1000)
write.csv(tree_sesmpd, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_sesmpd.csv")

#Mean nearest taxon distances
tree_sesmntd <- ses.mntd(presabs_lake, cophenetic(tree), null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999, iterations = 1000)
write.csv(tree_sesmntd, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_sesmntd.csv")

#Faith's PD
tree_sespd <- ses.pd(presabs_lake, tree, null.model = "taxa.labels", runs = 999, iterations = 1000, include.root = TRUE)
write.csv(tree_sespd, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_sespd.csv")


## Surveyed sites
# Dispersion
stree_disp <- dispersion(comm = surveyed_sites_lake, tree = stree, abund = F)
write.csv(stree_disp, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_disp.csv")

#Mean pairwise distances
stree_sesmpd <- ses.mpd(surveyed_sites_lake, cophenetic(stree), null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999, iterations = 1000)
write.csv(stree_sesmpd, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_sesmpd.csv")

#Mean nearest taxon distances
stree_sesmntd <- ses.mntd(surveyed_sites_lake, cophenetic(stree), null.model = "taxa.labels", abundance.weighted = FALSE, runs = 999, iterations = 1000)
write.csv(stree_sesmntd, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_sesmntd.csv")

#Faith's PD
stree_sespd <- ses.pd(surveyed_sites_lake, stree, null.model = "taxa.labels", runs = 999, iterations = 1000, include.root = TRUE)
write.csv(stree_sespd, "/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_sespd.csv")
```

### PD alpha files mntd, mpd, and pd to read into R 
- Read in Phylogenetic Diversity files and combine with env data frame
```{r, PD alpha file read-in}
## Reference
# Dispersion 
tree_disp <-read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_disp.csv")
tree_disp_env <- merge(tree_disp, env, by = "X", sort = F)
tree_disp_env$measure <- "tree_disp"
tree_disp_env$Site_type <- factor(tree_disp_env$Site_type, levels = c("Reference", "Ocean", "Mixed", "Stratified"))
row.names(tree_disp_env) <- tree_disp_env$X

# Mean pairwise distances
tree_sesmpd <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_sesmpd.csv")
tree_sesmpd_env <- merge(tree_sesmpd, env, by = "X", sort = F)
tree_sesmpd_env$measure <- "tree_sesmpd"
tree_sesmpd_env$Site_type <- factor(tree_sesmpd_env$Site_type, levels = c("Reference", "Ocean", "Mixed", "Stratified"))
row.names(tree_sesmpd_env) <- tree_sesmpd_env$X

# Mean nearest taxon distances
tree_sesmntd <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_sesmntd.csv")
tree_sesmntd_env <- merge(tree_sesmntd, env, by = "X", sort = F)
tree_sesmntd_env$measure <- "tree_sesmntd"
tree_sesmntd_env$Site_type <- factor(tree_sesmntd_env$Site_type, levels = c("Reference", "Ocean", "Mixed", "Stratified"))
row.names(tree_sesmntd_env) <- tree_sesmntd_env$X

#Faith's PD
tree_sespd <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/tree_sespd.csv")
tree_sespd_env <- merge(tree_sespd, env, by = "X", sort = F)
tree_sespd_env$measure <- "tree_sespd"
tree_sespd_env$Site_type <- factor(tree_sespd_env$Site_type, levels = c("Reference", "Ocean", "Mixed", "Stratified"))
stree_sespd_env <- merge(tree_sespd_env, tree_disp, by = "X", sort = F)
row.names(tree_sespd_env) <- tree_sespd_env$X


## Surveyed sites
# Dispersion 
stree_disp <-read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_disp.csv")
stree_disp_env <- merge(stree_disp, env, by = "X", sort = F)
stree_disp_env$measure <- "stree_disp"
stree_disp_env$Site_type <- factor(stree_disp_env$Site_type, levels = c("Ocean", "Mixed", "Stratified"))
row.names(stree_disp_env) <- stree_disp_env$X

# Mean pairwise distances
stree_sesmpd <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_sesmpd.csv")
stree_sesmpd_env <- merge(stree_sesmpd, env, by = "X", sort = F)
stree_sesmpd_env$measure <- "stree_sesmpd"
stree_sesmpd_env$Site_type <- factor(stree_sesmpd_env$Site_type, levels = c("Ocean", "Mixed", "Stratified"))
row.names(stree_sesmpd_env) <- stree_sesmpd_env$X

# Mean nearest taxon distances
stree_sesmntd <- read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_sesmntd.csv")
stree_sesmntd_env <- merge(stree_sesmntd, env, by = "X", sort = F)
stree_sesmntd_env$measure <- "stree_sesmntd"
stree_sesmntd_env$Site_type <- factor(stree_sesmntd_env$Site_type, levels = c("Ocean", "Mixed", "Stratified"))
row.names(stree_sesmntd_env) <- stree_sesmntd_env$X

#Faith's PD
stree_sespd <-read.csv("/Users/bailey/Documents/research/fish_biodiversity/data/analyses/PD/stree_sespd.csv")
stree_sespd_env <- merge(stree_sespd, env, by = "X", sort = F)
stree_sespd_env$measure <- "stree_sespd"
stree_sespd_env$Site_type <- factor(stree_sespd_env$Site_type, levels = c("Ocean", "Mixed", "Stratified"))
stree_sespd_env <- merge(stree_sespd_env, stree_disp, by = "X", sort = F)
row.names(stree_sespd_env) <- stree_sespd_env$X
```

### PD alpha outliers for each site type
```{r, PD alpha outliers}
outlier_PD_alpha <- stree_sespd_env %>%
  group_by(Site_type) %>%
  mutate(
    Q1 = quantile(pd.obs.z, 0.25),
    Q3 = quantile(pd.obs.z, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = pd.obs.z < lower_bound | pd.obs.z > upper_bound
  )

outlier_PD_alpha <- as.data.frame(outlier_PD_alpha)
row.names(outlier_PD_alpha) <- outlier_PD_alpha$X

# Create the plot
(outlier_PD_alpha_plot <- ggplot(outlier_PD_alpha, aes(x = Site_type, y = pd.obs.z, fill = Site_type)) +
  geom_violin(alpha = 0.9, draw_quantiles = c(0.25, 0.5, 0.75), aes(fill = Site_type)) +
  geom_jitter(aes(color = is_outlier), width = 0.1, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = outlier_PD_alpha, label = outlier_PD_alpha$X, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme(text = element_text(size = 16),
        legend.position = "right",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  scale_y_continuous(expand = c(0,0.5)) +
  labs(y = "PRic-z", x = "Site type", color = "Outlier:", tag = "b") +
  guides(fill = "none"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/outlier_PD_alpha.jpg", outlier_PD_alpha_plot, width = 6.26, height = 6, units = "in")
```

### Identify VIF of environmental and geographic variables
```{r, env and geo VIFs}
library(MASS)

# Determine which env variables have variance
environment <- c("temperature_median", "salinity_median", "oxygen_median", "pH_median")
nzv <- nearZeroVar(env[,environment])
nzv
# All env variables have variance

mod <-  lm(pd.obs ~ temperature_median + salinity_median + oxygen_median + pH_median, stree_sespd_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)

mod <-  lm(pd.obs ~ temperature_median + salinity_median + oxygen_median, stree_sespd_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)
environment <- c("temperature_median", "salinity_median", "oxygen_median")

mod <- aov(temperature_median ~ Site_type, stree_sespd_env)
summary(mod)
mod <- aov(pd.obs ~ temperature_median + Site_type, stree_sespd_env)
car::vif(mod)

mod <- aov(salinity_median ~ Site_type, stree_sespd_env)
summary(mod)
mod <- aov(pd.obs ~ salinity_median + Site_type, stree_sespd_env)
car::vif(mod)

mod <- aov(oxygen_median ~ Site_type, stree_sespd_env)
summary(mod)
mod <- aov(pd.obs ~ oxygen_median + Site_type, stree_sespd_env)
car::vif(mod)

SR_ss_env <- stree_sespd_env[surveyed_sites_env,]
SR_ss_env$Site_type <- factor(SR_ss_env$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Run linear discriminant analysis of environmental variables
LDA <- lda(SR_ss_env[,environment], SR_ss_env$Site_type)
spe.class <- predict(LDA)$class
spe.post <- predict(LDA)$posterior
(spe.table <- table(SR_ss_env$Site_type, spe.class))
diag(prop.table(spe.table, 1))


# Determine which geo variables have variance
geography <- c("volume_m3_w_chemocline", "volume_m3", "surface_area_m2", "distance_to_ocean_min_m", "distance_to_ocean_mean_m", "distance_to_ocean_median_m", "tidal_lag_time_minutes", "tidal_efficiency", "perimeter_fromSat", "max_depth", "logArea")
nzv <- nearZeroVar(env[,geography])
nzv
# All geo variables have variance

mod <-  lm(pd.obs ~ volume_m3_w_chemocline + volume_m3 + surface_area_m2 + distance_to_ocean_min_m + distance_to_ocean_mean_m + distance_to_ocean_median_m + tidal_lag_time_minutes + tidal_efficiency + perimeter_fromSat + max_depth + logArea, stree_sespd_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)

mod <-  lm(pd.obs ~ distance_to_ocean_min_m + max_depth + logArea, stree_sespd_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)
geography <- c("distance_to_ocean_min_m", "max_depth", "logArea")

mod <- aov(distance_to_ocean_min_m ~ Site_type, stree_sespd_env)
summary(mod)
mod <- aov(pd.obs ~ distance_to_ocean_min_m + Site_type, stree_sespd_env)
car::vif(mod)

mod <- aov(max_depth ~ Site_type, stree_sespd_env)
summary(mod)
mod <- aov(pd.obs ~ max_depth + Site_type, stree_sespd_env)
car::vif(mod)

mod <- aov(logArea ~ Site_type, stree_sespd_env)
summary(mod)
mod <- aov(pd.obs ~ logArea + Site_type, stree_sespd_env)
car::vif(mod)

SR_ss_geo <- stree_sespd_env[surveyed_sites,]
SR_ss_geo$Site_type <- factor(SR_ss_geo$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Run linear discriminant analysis of geographical variables
LDA <- lda(SR_ss_geo[,geography], SR_ss_geo$Site_type)
spe.class <- predict(LDA)$class
spe.post <- predict(LDA)$posterior
(spe.table <- table(SR_ss_geo$Site_type, spe.class))
diag(prop.table(spe.table, 1))


# Combine environment and geography
mod <-  lm(pd.obs ~ salinity_median + oxygen_median + temperature_median + distance_to_ocean_min_m + max_depth + logArea, stree_sespd_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)

mod <-  lm(pd.obs ~ oxygen_median + temperature_median + distance_to_ocean_min_m + logArea, stree_sespd_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)
envgeo <- c("oxygen_median", "temperature_median", "distance_to_ocean_min_m", "logArea")


# Run linear discriminant analysis of geographical variables
LDA <- lda(SR_ss_env[,envgeo], SR_ss_env$Site_type)
spe.class <- predict(LDA)$class
spe.post <- predict(LDA)$posterior
(spe.table <- table(SR_ss_env$Site_type, spe.class))
diag(prop.table(spe.table, 1))
```

### PD alpha & site type
```{r, PD alpha & site type}
### Site_type
# Run ANOVA on the original data
anova_PRic_result <- aov(pd.obs ~ Site_type, data = stree_sespd_env[surveyed_sites,])
summary(anova_PRic_result)

# Perform Tukey's HSD test
tukey_PRic_result <- TukeyHSD(anova_PRic_result)
print(tukey_PRic_result)

# Sites
N <- length(anova_PRic_result$residuals)
# Categories
k <- length(anova_PRic_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_PRic_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Site_type <- tukey_PRic_result$Site_type

# q-values
(OM_q <- (abs(Site_type[1,1]))/USE)
(MS_q <- (abs(Site_type[3,1]))/BSE)
(SO_q <- (abs(Site_type[2,1]))/USE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx

# Run ANOVA on the original data
anova_PRicz_result <- aov(pd.obs.z ~ Site_type, data = stree_sespd_env[surveyed_sites,])
summary(anova_PRicz_result)

# Perform Tukey's HSD test
tukey_PRicz_result <- TukeyHSD(anova_PRicz_result)
print(tukey_PRicz_result)

# Sites
N <- length(anova_PRicz_result$residuals)
# Categories
k <- length(anova_PRicz_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_PRicz_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Site_type <- tukey_PRicz_result$Site_type

# q-values
(OM_q <- (abs(Site_type[1,1]))/USE)
(MS_q <- (abs(Site_type[3,1]))/BSE)
(SO_q <- (abs(Site_type[2,1]))/USE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx

# Run ANOVA on the original data
anova_PDisp_result <- aov(Dispersion ~ Site_type, data = stree_sespd_env[surveyed_sites,])
summary(anova_PDisp_result)

# Perform Tukey's HSD test
tukey_PDisp_result <- TukeyHSD(anova_PDisp_result)
print(tukey_PDisp_result)

# Sites
N <- length(anova_PDisp_result$residuals)
# Categories
k <- length(anova_PDisp_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_PDisp_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Site_type <- tukey_PDisp_result$Site_type

# q-values
(OM_q <- (abs(Site_type[1,1]))/USE)
(MS_q <- (abs(Site_type[3,1]))/BSE)
(SO_q <- (abs(Site_type[2,1]))/USE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx
```

### PD alpha env and geo plots
```{r, PD alpha lm plots}
# Temperature
PD_alpha_T_plot <- ggplot(data = stree_sespd_env[surveyed_sites_env,], mapping = aes(y = pd.obs.z, x = temperature_median, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = stree_sespd_env[surveyed_sites_env,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Temperature (ºC)", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "a")
(PD_alpha_T_plot <- PD_alpha_T_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_T_plot.jpg", plot = PD_alpha_T_plot, width = 6.26, height = 6, units = "in")

# Salinity
PD_alpha_S_plot <- ggplot(data = stree_sespd_env[surveyed_sites_env,], mapping = aes(y = pd.obs.z, x = salinity_median, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = stree_sespd_env[surveyed_sites_env,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Salinity (ppt)", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "b")
(PD_alpha_S_plot <- PD_alpha_S_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_S_plot.jpg", plot = PD_alpha_S_plot, width = 6.26, height = 6, units = "in")

# Oxygen
PD_alpha_O_plot <- ggplot(data = stree_sespd_env[surveyed_sites_env,], mapping = aes(y = pd.obs.z, x = oxygen_median, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = stree_sespd_env[surveyed_sites_env,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Oxygen (mg/L)", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "c")
(PD_alpha_O_plot <- PD_alpha_O_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_O_plot.jpg", plot = PD_alpha_O_plot, width = 6.26, height = 6, units = "in")

# Distance from the ocean mean
PD_alpha_D_plot <- ggplot(data = stree_sespd_env[surveyed_sites,], mapping = aes(y = pd.obs.z, x = distance_to_ocean_min_m, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = stree_sespd_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Isolation (m)", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "a")
(PD_alpha_D_plot <- PD_alpha_D_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_D_plot.jpg", plot = PD_alpha_D_plot, width = 6.26, height = 6, units = "in")

# Max depth
PD_alpha_MD_plot <- ggplot(data = stree_sespd_env[surveyed_sites,], mapping = aes(y = pd.obs.z, x = max_depth, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = stree_sespd_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Age (m)", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "b")
(PD_alpha_MD_plot <- PD_alpha_MD_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_MD_plot.jpg", plot = PD_alpha_MD_plot, width = 6.26, height = 6, units = "in")

# Log Area
PD_alpha_LA_plot <- ggplot(data = stree_sespd_env[surveyed_sites,], mapping = aes(y = pd.obs.z, x = logArea, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = stree_sespd_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Log Area (m"^"2"~")", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "c")
(PD_alpha_LA_plot <- PD_alpha_LA_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_LA_plot.jpg", plot = PD_alpha_LA_plot, width = 6.26, height = 6, units = "in")
```

### PD alpha with env and geo linear models & ANOVAs
```{r, PD alpha lm and ANOVA}
par(mfrow=c(2,2)) 
##### PRicz ANOVAs
#### Environmental
### Surveyed sites
## Temperature
# Interaction
PD_PRicz_am_temp <- aov(pd.obs.z ~ temperature_median * Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_temp) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_temp)
# Plot residuals
plot(PD_PRicz_am_temp)
# Summarize the ANOVA results
summary(PD_PRicz_am_temp)
# p-values
temp_p_values <- summary(PD_PRicz_am_temp)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
# ANCOVA
PD_PRicz_am_temp <- aov(pd.obs.z ~ temperature_median + Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_temp) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_temp)
# Plot residuals
plot(PD_PRicz_am_temp)

### Ocean & Mixed sites
PD_PRicz_OM_am_temp <- aov(pd.obs.z ~ temperature_median * Site_type, data = stree_sespd_env[ocean_mixed_sites_env,])
summary(PD_PRicz_OM_am_temp)
PD_PRicz_OM_am_temp <- aov(pd.obs.z ~ temperature_median + Site_type, data = stree_sespd_env[ocean_mixed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OM_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OM_am_temp) ~ stree_sespd_env[ocean_mixed_sites_env,"Site_type"])

### Ocean & Stratified sites
PD_PRicz_OS_am_temp <- aov(pd.obs.z ~ temperature_median * Site_type, data = stree_sespd_env[ocean_stratified_sites_env,])
summary(PD_PRicz_OS_am_temp)
PD_PRicz_OS_am_temp <- aov(pd.obs.z ~ temperature_median + Site_type, data = stree_sespd_env[ocean_stratified_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OS_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OS_am_temp) ~ stree_sespd_env[ocean_stratified_sites_env,"Site_type"])

### Mixed & Stratified lakes
PD_PRicz_MS_am_temp <- aov(pd.obs.z ~ temperature_median * Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
summary(PD_PRicz_MS_am_temp)
PD_PRicz_MS_am_temp <- aov(pd.obs.z ~ temperature_median + Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_MS_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_MS_am_temp) ~ stree_sespd_env[mixed_stratified_lakes,"Site_type"])

### Mixed lakes
PD_PRicz_M_am_temp <- aov(pd.obs.z ~ temperature_median, data = stree_sespd_env[mixed_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_M_am_temp))

### Stratified lakes
PD_PRicz_S_am_temp <- aov(pd.obs.z ~ temperature_median, data = stree_sespd_env[stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_S_am_temp))

## Salinity
# Interaction
PD_PRicz_am_sal <- aov(pd.obs.z ~ salinity_median * Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_sal) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_sal)
# Plot residuals
plot(PD_PRicz_am_sal)
# Summarize the ANOVA results
summary(PD_PRicz_am_sal)
# p-values
sal_p_values <- summary(PD_PRicz_am_sal)[[1]][, "Pr(>F)"]
p_values <- c(sal_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
# ANCOVA
PD_PRicz_am_sal <- aov(pd.obs.z ~ salinity_median + Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_sal) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_sal)
# Plot residuals
plot(PD_PRicz_am_sal)

### Ocean & Mixed sites
PD_PRicz_OM_am_sal <- aov(pd.obs.z ~ salinity_median * Site_type, data = stree_sespd_env[ocean_mixed_sites_env,])
summary(PD_PRicz_OM_am_sal)
PD_PRicz_OM_am_sal <- aov(pd.obs.z ~ salinity_median + Site_type, data = stree_sespd_env[ocean_mixed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OM_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OM_am_sal) ~ stree_sespd_env[ocean_mixed_sites_env,"Site_type"])

### Ocean & Stratified sites
PD_PRicz_OS_am_sal <- aov(pd.obs.z ~ salinity_median * Site_type, data = stree_sespd_env[ocean_stratified_sites_env,])
summary(PD_PRicz_OS_am_sal)
PD_PRicz_OS_am_sal <- aov(pd.obs.z ~ salinity_median + Site_type, data = stree_sespd_env[ocean_stratified_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OS_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OS_am_sal) ~ stree_sespd_env[ocean_stratified_sites_env,"Site_type"])

### Mixed & Stratified lakes
PD_PRicz_MS_am_sal <- aov(pd.obs.z ~ salinity_median * Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
summary(PD_PRicz_MS_am_sal)
PD_PRicz_MS_am_sal <- aov(pd.obs.z ~ salinity_median + Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_MS_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_MS_am_sal) ~ stree_sespd_env[mixed_stratified_lakes,"Site_type"])

### Mixed lakes
PD_PRicz_M_am_sal <- aov(pd.obs.z ~ salinity_median, data = stree_sespd_env[mixed_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_M_am_sal))

### Stratified lakes
PD_PRicz_S_am_sal <- aov(pd.obs.z ~ salinity_median, data = stree_sespd_env[stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_S_am_sal))

## Oxygen
# Interaction
PD_PRicz_am_oxy <- aov(pd.obs.z ~ oxygen_median * Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_oxy) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_oxy)
# Plot residuals
plot(PD_PRicz_am_oxy)
# Summarize the ANOVA results
summary(PD_PRicz_am_oxy)
# p-values
oxy_p_values <- summary(PD_PRicz_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
# ANCOVA
PD_PRicz_am_oxy <- aov(pd.obs.z ~ oxygen_median + Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_oxy) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_oxy)
# Plot residuals
plot(PD_PRicz_am_oxy)

### Ocean & Mixed sites
PD_PRicz_OM_am_oxy <- aov(pd.obs.z ~ oxygen_median * Site_type, data = stree_sespd_env[ocean_mixed_sites_env,])
summary(PD_PRicz_OM_am_oxy)
PD_PRicz_OM_am_oxy <- aov(pd.obs.z ~ oxygen_median + Site_type, data = stree_sespd_env[ocean_mixed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OM_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OM_am_oxy) ~ stree_sespd_env[ocean_mixed_sites_env,"Site_type"])

### Ocean & Stratified sites
PD_PRicz_OS_am_oxy <- aov(pd.obs.z ~ oxygen_median * Site_type, data = stree_sespd_env[ocean_stratified_sites_env,])
summary(PD_PRicz_OS_am_oxy)
PD_PRicz_OS_am_oxy <- aov(pd.obs.z ~ oxygen_median + Site_type, data = stree_sespd_env[ocean_stratified_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OS_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OS_am_oxy) ~ stree_sespd_env[ocean_stratified_sites_env,"Site_type"])

### Mixed & Stratified lakes
PD_PRicz_MS_am_oxy <- aov(pd.obs.z ~ oxygen_median * Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
summary(PD_PRicz_MS_am_oxy)
PD_PRicz_MS_am_oxy <- aov(pd.obs.z ~ oxygen_median + Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_MS_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_MS_am_oxy) ~ stree_sespd_env[mixed_stratified_lakes,"Site_type"])

### Mixed lakes
PD_PRicz_M_am_oxy <- aov(pd.obs.z ~ oxygen_median, data = stree_sespd_env[mixed_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_M_am_oxy))

### Stratified lakes
PD_PRicz_S_am_oxy <- aov(pd.obs.z ~ oxygen_median, data = stree_sespd_env[stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_S_am_oxy))

# Summarize ANOVA results and calculate pairwise comparisons
summary(PD_PRicz_am_temp)
PD_PRicz_amp_temp <- emmeans(PD_PRicz_am_temp, pairwise ~ Site_type, adjust = "bonferroni")
PD_PRicz_amp_temp$contrasts
summary(PD_PRicz_am_sal)
PD_PRicz_amp_sal <- emmeans(PD_PRicz_am_sal, pairwise ~ Site_type, adjust = "bonferroni")
PD_PRicz_amp_sal$contrasts
summary(PD_PRicz_am_oxy)
PD_PRicz_amp_oxy <- emmeans(PD_PRicz_am_oxy, pairwise ~ Site_type, adjust = "bonferroni")
PD_PRicz_amp_oxy$contrasts
# p-values
temp_p_values <- summary(PD_PRicz_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PRicz_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PRicz_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OM lm results
summary(PD_PRicz_OM_am_temp)
summary(PD_PRicz_OM_am_sal)
summary(PD_PRicz_OM_am_oxy)
# p-values
temp_p_values <- summary(PD_PRicz_OM_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PRicz_OM_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PRicz_OM_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OS lm results
summary(PD_PRicz_OS_am_temp)
summary(PD_PRicz_OS_am_sal)
summary(PD_PRicz_OS_am_oxy)
# p-values
temp_p_values <- summary(PD_PRicz_OS_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PRicz_OS_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PRicz_OS_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize MS lm results
summary(PD_PRicz_MS_am_temp)
summary(PD_PRicz_MS_am_sal)
summary(PD_PRicz_MS_am_oxy)
# p-values
temp_p_values <- summary(PD_PRicz_MS_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PRicz_MS_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PRicz_MS_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize M lm results
summary(PD_PRicz_M_am_temp)
summary(PD_PRicz_M_am_sal)
summary(PD_PRicz_M_am_oxy)
# p-values
temp_p_values <- summary(PD_PRicz_M_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PRicz_M_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PRicz_M_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize S lm results
summary(PD_PRicz_S_am_temp)
summary(PD_PRicz_S_am_sal)
summary(PD_PRicz_S_am_oxy)
# p-values
temp_p_values <- summary(PD_PRicz_S_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PRicz_S_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PRicz_S_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

#### Geographical
### Surveyed sites
## Distance
# Interaction
PD_PRicz_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m * Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_dist) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_dist)
# Plot residuals
plot(PD_PRicz_am_dist)
# Summarize the ANOVA results
summary(PD_PRicz_am_dist)
# p-values
dist_p_values <- summary(PD_PRicz_am_dist)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
# ANCOVA
PD_PRicz_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m + Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_dist) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_dist)
# Plot residuals
plot(PD_PRicz_am_dist)

### Ocean & Mixed sites
PD_PRicz_OM_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m * Site_type, data = stree_sespd_env[ocean_mixed_sites,])
summary(PD_PRicz_OM_am_dist)
PD_PRicz_OM_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m + Site_type, data = stree_sespd_env[ocean_mixed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OM_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OM_am_dist) ~ stree_sespd_env[ocean_mixed_sites,"Site_type"])

### Ocean & Stratified sites
PD_PRicz_OS_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m * Site_type, data = stree_sespd_env[ocean_stratified_sites,])
summary(PD_PRicz_OS_am_dist)
PD_PRicz_OS_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m + Site_type, data = stree_sespd_env[ocean_stratified_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OS_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OS_am_dist) ~ stree_sespd_env[ocean_stratified_sites,"Site_type"])

### Mixed & Stratified lakes
PD_PRicz_MS_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m * Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
summary(PD_PRicz_MS_am_dist)
PD_PRicz_MS_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m + Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_MS_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_MS_am_dist) ~ stree_sespd_env[mixed_stratified_lakes,"Site_type"])

### Mixed lakes
PD_PRicz_M_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m, data = stree_sespd_env[mixed_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_M_am_dist))

### Stratified lakes
PD_PRicz_S_am_dist <- aov(pd.obs.z ~ distance_to_ocean_min_m, data = stree_sespd_env[stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_S_am_dist))

## Max Depth
# Interaction
PD_PRicz_am_mxd <- aov(pd.obs.z ~ max_depth * Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_mxd) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_mxd)
# Plot residuals
plot(PD_PRicz_am_mxd)
# Summarize the ANOVA results
summary(PD_PRicz_am_mxd)
# p-values
mxd_p_values <- summary(PD_PRicz_am_mxd)[[1]][, "Pr(>F)"]
p_values <- c(mxd_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Ocean & Mixed sites
PD_PRicz_OM_am_mxd <- aov(pd.obs.z ~ max_depth * Site_type, data = stree_sespd_env[ocean_mixed_sites,])
summary(PD_PRicz_OM_am_mxd)
PD_PRicz_OM_am_mxd <- aov(pd.obs.z ~ max_depth + Site_type, data = stree_sespd_env[ocean_mixed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OM_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OM_am_mxd) ~ stree_sespd_env[ocean_mixed_sites,"Site_type"])

### Ocean & Stratified sites
PD_PRicz_OS_am_mxd <- aov(pd.obs.z ~ max_depth * Site_type, data = stree_sespd_env[ocean_stratified_sites,])
summary(PD_PRicz_OS_am_mxd)
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OS_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OS_am_mxd) ~ stree_sespd_env[ocean_stratified_sites,"Site_type"])

### Mixed & Stratified lakes
PD_PRicz_MS_am_mxd <- aov(pd.obs.z ~ max_depth * Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
summary(PD_PRicz_MS_am_mxd)
PD_PRicz_MS_am_mxd <- aov(pd.obs.z ~ max_depth + Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_MS_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_MS_am_mxd) ~ stree_sespd_env[mixed_stratified_lakes,"Site_type"])

### Mixed lakes
PD_PRicz_M_am_mxd <- aov(pd.obs.z ~ max_depth, data = stree_sespd_env[mixed_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_M_am_mxd))

### Stratified lakes
PD_PRicz_S_am_mxd <- aov(pd.obs.z ~ max_depth, data = stree_sespd_env[stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_S_am_mxd))

## Log Area
# Interaction
PD_PRicz_am_lga <- aov(pd.obs.z ~ logArea * Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_lga) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_lga)
# Plot residuals
plot(PD_PRicz_am_lga)
# Summarize the ANOVA results
summary(PD_PRicz_am_lga)
# p-values
lga_p_values <- summary(PD_PRicz_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
# ANCOVA
PD_PRicz_am_lga <- aov(pd.obs.z ~ logArea + Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_am_lga) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PRicz_am_lga)
# Plot residuals
plot(PD_PRicz_am_lga)

### Ocean & Mixed sites
PD_PRicz_OM_am_lga <- aov(pd.obs.z ~ logArea * Site_type, data = stree_sespd_env[ocean_mixed_sites,])
summary(PD_PRicz_OM_am_lga)
PD_PRicz_OM_am_lga <- aov(pd.obs.z ~ logArea + Site_type, data = stree_sespd_env[ocean_mixed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OM_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OM_am_lga) ~ stree_sespd_env[ocean_mixed_sites,"Site_type"])

### Ocean & Stratified sites
PD_PRicz_OS_am_lga <- aov(pd.obs.z ~ logArea * Site_type, data = stree_sespd_env[ocean_stratified_sites,])
summary(PD_PRicz_OS_am_lga)
PD_PRicz_OS_am_lga <- aov(pd.obs.z ~ logArea + Site_type, data = stree_sespd_env[ocean_stratified_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_OS_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_OS_am_lga) ~ stree_sespd_env[ocean_stratified_sites,"Site_type"])

### Mixed & Stratified lakes
PD_PRicz_MS_am_lga <- aov(pd.obs.z ~ logArea * Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
summary(PD_PRicz_MS_am_lga)
PD_PRicz_MS_am_lga <- aov(pd.obs.z ~ logArea + Site_type, data = stree_sespd_env[mixed_stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_MS_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PRicz_MS_am_lga) ~ stree_sespd_env[mixed_stratified_lakes,"Site_type"])

### Mixed lakes
PD_PRicz_M_am_lga <- aov(pd.obs.z ~ logArea, data = stree_sespd_env[mixed_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_M_am_lga))

### Stratified lakes
PD_PRicz_S_am_lga <- aov(pd.obs.z ~ logArea, data = stree_sespd_env[stratified_lakes,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PRicz_S_am_lga))

# Summarize ANOVA results and calculate pairwise comparisons
summary(PD_PRicz_am_dist)
PD_PRicz_amp_dist <- emmeans(PD_PRicz_am_dist, pairwise ~ Site_type, adjust = "bonferroni")
PD_PRicz_amp_dist$contrasts
summary(PD_PRicz_am_mxd)
PD_PRicz_amp_mxd <- emmeans(PD_PRicz_am_mxd, pairwise ~ Site_type, adjust = "bonferroni")
PD_PRicz_amp_mxd$contrasts
summary(PD_PRicz_am_lga)
PD_PRicz_amp_lga <- emmeans(PD_PRicz_am_lga, pairwise ~ Site_type, adjust = "bonferroni")
PD_PRicz_amp_lga$contrasts
# p-values
dist_p_values <- summary(PD_PRicz_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PRicz_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PRicz_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OM lm results
summary(PD_PRicz_OM_am_dist)
summary(PD_PRicz_OM_am_mxd)
summary(PD_PRicz_OM_am_lga)
# p-values
dist_p_values <- summary(PD_PRicz_OM_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PRicz_OM_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PRicz_OM_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OS lm results
summary(PD_PRicz_OS_am_dist)
summary(PD_PRicz_OS_am_mxd)
summary(PD_PRicz_OS_am_lga)
# p-values
dist_p_values <- summary(PD_PRicz_OS_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PRicz_OS_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PRicz_OS_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize MS lm results
summary(PD_PRicz_MS_am_dist)
summary(PD_PRicz_MS_am_mxd)
summary(PD_PRicz_MS_am_lga)
# p-values
dist_p_values <- summary(PD_PRicz_MS_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PRicz_MS_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PRicz_MS_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize M lm results
summary(PD_PRicz_M_am_dist)
summary(PD_PRicz_M_am_mxd)
summary(PD_PRicz_M_am_lga)
# p-values
dist_p_values <- summary(PD_PRicz_M_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PRicz_M_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PRicz_M_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize S lm results
summary(PD_PRicz_S_am_dist)
summary(PD_PRicz_S_am_mxd)
summary(PD_PRicz_S_am_lga)
# p-values
dist_p_values <- summary(PD_PRicz_S_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PRicz_S_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PRicz_S_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


##### PDisp ANOVAs
#### Environmental
### Surveyed sites
## Temperature
# Interaction
PD_PDisp_am_temp <- aov(Dispersion ~ temperature_median * Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_temp) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_temp)
# Plot residuals
plot(PD_PDisp_am_temp)
# Summarize the ANOVA results
summary(PD_PDisp_am_temp)
# ANCOVA
PD_PDisp_am_temp <- aov(Dispersion ~ temperature_median + Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_temp) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_temp)
# Plot residuals
plot(PD_PDisp_am_temp)

## Salinity
# Interaction
PD_PDisp_am_sal <- aov(Dispersion ~ salinity_median * Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_sal) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_sal)
# Plot residuals
plot(PD_PDisp_am_sal)
# Summarize the ANOVA results
summary(PD_PDisp_am_sal)
# ANCOVA
PD_PDisp_am_sal <- aov(Dispersion ~ salinity_median + Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_sal) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_sal)
# Plot residuals
plot(PD_PDisp_am_sal)

## Oxygen
# Interaction
PD_PDisp_am_oxy <- aov(Dispersion ~ oxygen_median * Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_oxy) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_oxy)
# Plot residuals
plot(PD_PDisp_am_oxy)
# Summarize the ANOVA results
summary(PD_PDisp_am_oxy)
# ANCOVA
PD_PDisp_am_oxy <- aov(Dispersion ~ oxygen_median + Site_type, data = stree_sespd_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_oxy) ~ stree_sespd_env[surveyed_sites_env,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_oxy)
# Plot residuals
plot(PD_PDisp_am_oxy)

# Summarize ANOVA results and calculate pairwise comparisons
summary(PD_PDisp_am_temp)
PD_PDisp_amp_temp <- emmeans(PD_PDisp_am_temp, pairwise ~ Site_type, adjust = "bonferroni")
PD_PDisp_amp_temp$contrasts
summary(PD_PDisp_am_sal)
PD_PDisp_amp_sal <- emmeans(PD_PDisp_am_sal, pairwise ~ Site_type, adjust = "bonferroni")
PD_PDisp_amp_sal$contrasts
summary(PD_PDisp_am_oxy)
PD_PDisp_amp_oxy <- emmeans(PD_PDisp_am_oxy, pairwise ~ Site_type, adjust = "bonferroni")
PD_PDisp_amp_oxy$contrasts
# p-values
temp_p_values <- summary(PD_PDisp_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(PD_PDisp_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(PD_PDisp_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Mixed lakes
PD_PDisp_env_M_lm <- lm(Dispersion ~ salinity_median + oxygen_median + temperature_median, stree_sespd_env[mixed_lakes,])
summary(PD_PDisp_env_M_lm)
Anova(PD_PDisp_env_M_lm, type = 3)
p_values <- summary(PD_PDisp_env_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Stratified lakes
PD_PDisp_env_S_lm <- lm(Dispersion ~ salinity_median + oxygen_median + temperature_median, stree_sespd_env[stratified_lakes,])
summary(PD_PDisp_env_S_lm)
Anova(PD_PDisp_env_S_lm, type = 3)
p_values <- summary(PD_PDisp_env_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

#### Geographical
### Surveyed sites
## Distance
# Interaction
PD_PDisp_am_dist <- aov(Dispersion ~ distance_to_ocean_min_m * Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_dist) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_dist)
# Plot residuals
plot(PD_PDisp_am_dist)
# Summarize the ANOVA results
summary(PD_PDisp_am_dist)
# ANCOVA
PD_PDisp_am_dist <- aov(Dispersion ~ distance_to_ocean_min_m + Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_dist) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_dist)
# Plot residuals
plot(PD_PDisp_am_dist)

## Max Depth
# Interaction
PD_PDisp_am_mxd <- aov(Dispersion ~ max_depth * Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_mxd) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_mxd)
# Plot residuals
plot(PD_PDisp_am_mxd)
# Summarize the ANOVA results
summary(PD_PDisp_am_mxd)
# ANCOVA
PD_PDisp_am_mxd <- aov(Dispersion ~ max_depth + Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_mxd) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_mxd)
# Plot residuals
plot(PD_PDisp_am_mxd)

## Log Area
# Interaction
PD_PDisp_am_lga <- aov(Dispersion ~ logArea * Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_lga) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_lga)
# Plot residuals
plot(PD_PDisp_am_lga)
# Summarize the ANOVA results
summary(PD_PDisp_am_lga)
# ANCOVA
PD_PDisp_am_lga <- aov(Dispersion ~ logArea + Site_type, data = stree_sespd_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(PD_PDisp_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(PD_PDisp_am_lga) ~ stree_sespd_env[surveyed_sites,"Site_type"])
# Check for outliers
outlierTest(PD_PDisp_am_lga)
# Plot residuals
plot(PD_PDisp_am_lga)

# Summarize ANOVA results and calculate pairwise comparisons
summary(PD_PDisp_am_dist)
PD_PDisp_amp_dist <- emmeans(PD_PDisp_am_dist, pairwise ~ Site_type, adjust = "bonferroni")
PD_PDisp_amp_dist$contrasts
summary(PD_PDisp_am_mxd)
PD_PDisp_amp_mxd <- emmeans(PD_PDisp_am_mxd, pairwise ~ Site_type, adjust = "bonferroni")
PD_PDisp_amp_mxd$contrasts
summary(PD_PDisp_am_lga)
PD_PDisp_amp_lga <- emmeans(PD_PDisp_am_lga, pairwise ~ Site_type, adjust = "bonferroni")
PD_PDisp_amp_lga$contrasts
# p-values
dist_p_values <- summary(PD_PDisp_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(PD_PDisp_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(PD_PDisp_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Mixed lakes
PD_PDisp_geo_M_lm <- lm(Dispersion ~ distance_to_ocean_min_m + max_depth + logArea, stree_sespd_env[mixed_lakes,])
summary(PD_PDisp_geo_M_lm)
Anova(PD_PDisp_geo_M_lm, type = 3)
p_values <- summary(PD_PDisp_geo_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Stratified lakes
PD_PDisp_geo_S_lm <- lm(Dispersion ~ distance_to_ocean_min_m + max_depth + logArea, stree_sespd_env[stratified_lakes,])
summary(PD_PDisp_geo_S_lm)
Anova(PD_PDisp_geo_S_lm, type = 3)
p_values <- summary(PD_PDisp_geo_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
par(mfrow=c(1,1))
```

### PD alpha z-scores & dispersion
```{r, z-scores & dispersion}
PD_alpha_plot <- ggplot(data = stree_sespd_env, mapping = aes(y = pd.obs.z, x = Dispersion, color = Site_type, fill = Site_type)) +
  geom_point(stat = 'identity',
    size = 4,
    alpha = 1) + 
  geom_text_repel(label = stree_sespd_env[,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 14),
    legend.position = "bottom", 
    legend.title = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    axis.text = element_text(size = 14, color = "black"),
    axis.line = element_line(color = "black"),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(x="PDisp", y="PRic-z", colour = "Site type:", fill = "Site type:", tag = "a")
(PD_alpha_plot <- PD_alpha_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_alpha_plot.jpg", plot = PD_alpha_plot, width = 4.88, height = 6, units = "in")
```

## PD beta diversity
### PD beta diversity distance calculations plus dispersion and PERMANOVA
```{r, PD beta distances}
### Regular
# Reference
PD_beta_ref_dist <- BAT::beta(presabs_lake, tree, abund = F)
(PD_beta_ref_BD <- betadisper(PD_beta_ref_dist$Btotal, Site_type_group_ref))
(PD_beta_ref_AOV <- anova(PD_beta_ref_BD))
(PD_beta_ref_THSD <- TukeyHSD(PD_beta_ref_BD))
(PD_beta_ref_PM <- adonis2(PD_beta_ref_dist$Btotal ~ env[,"Site_type"], permutations = 999))
(PD_beta_ref_PM_pair <- pairwise.adonis(PD_beta_ref_dist$Btotal, env[,"Site_type"], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites total
PD_beta_dist <- BAT::beta(surveyed_sites_lake, stree, abund = F)
(PD_beta_BD <- betadisper(PD_beta_dist$Btotal, Site_type_group))
(PD_beta_AOV <- anova(PD_beta_BD))
(PD_beta_THSD <- TukeyHSD(PD_beta_BD))
(PD_beta_PM <- adonis2(PD_beta_dist$Btotal ~ env[surveyed_sites,"Site_type"], permutations = 999))
(PD_beta_PM_pair <- pairwise.adonis(PD_beta_dist$Btotal, env[surveyed_sites,"Site_type"], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites replacement
(PD_beta_BD <- betadisper(PD_beta_dist$Brepl, Site_type_group))
(PD_beta_AOV <- anova(PD_beta_BD))
(PD_beta_THSD <- TukeyHSD(PD_beta_BD))
(PD_beta_PM <- adonis2(PD_beta_dist$Brepl ~ env[surveyed_sites,"Site_type"], permutations = 999))
(PD_beta_PM_pair <- pairwise.adonis(PD_beta_dist$Brepl, env[surveyed_sites,"Site_type"], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites richness
(PD_beta_BD <- betadisper(PD_beta_dist$Brich, Site_type_group))
(PD_beta_AOV <- anova(PD_beta_BD))
(PD_beta_THSD <- TukeyHSD(PD_beta_BD))
(PD_beta_PM <- adonis2(PD_beta_dist$Brich ~ env[surveyed_sites,"Site_type"], permutations = 999))
(PD_beta_PM_pair <- pairwise.adonis(PD_beta_dist$Brich, env[surveyed_sites,"Site_type"], p.adjust.m = "bonferroni", perm = 999))
# Without LCN
PD_beta_wo_LCN_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_LCN,], stree, abund = F)
(PD_beta_wo_LCN_BD <- betadisper(PD_beta_wo_LCN_dist$Btotal, Site_type_group[-8]))
(PD_beta_wo_LCN_AOV <- anova(PD_beta_wo_LCN_BD))
(PD_beta_wo_LCN_THSD <- TukeyHSD(PD_beta_wo_LCN_BD))
(PD_beta_wo_LCN_PM <- adonis2(PD_beta_wo_LCN_dist$Btotal ~ env[surveyed_sites_wo_LCN,"Site_type"], permutations = 999))
(PD_beta_wo_LCN_PM_pair <- pairwise.adonis(PD_beta_wo_LCN_dist$Btotal, env[surveyed_sites_wo_LCN,"Site_type"], p.adjust.m = "bonferroni", perm = 999))
# Without TLN and HLM
PD_beta_wo_TLN_HLM_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_TLN_HLM,], stree, abund = F)
(PD_beta_wo_TLN_HLM_BD <- betadisper(PD_beta_wo_TLN_HLM_dist$Btotal, Site_type_group[-c(5,21)]))
(PD_beta_wo_TLN_HLM_AOV <- anova(PD_beta_wo_TLN_HLM_BD))
(PD_beta_wo_TLN_HLM_THSD <- TukeyHSD(PD_beta_wo_TLN_HLM_BD))
(PD_beta_wo_TLN_HLM_PM <- adonis2(PD_beta_wo_TLN_HLM_dist$Btotal ~ env[surveyed_sites_wo_TLN_HLM,"Site_type"], permutations = 999))
(PD_beta_wo_TLN_HLM_PM_pair <- pairwise.adonis(PD_beta_wo_TLN_HLM_dist$Btotal, env[surveyed_sites_wo_TLN_HLM,"Site_type"], p.adjust.m = "bonferroni", perm = 999))
# Mixed and stratified lakes
PD_beta_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], stree, abund = F)
# Ocean sites and mixed lakes
PD_beta_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites,], stree, abund = F)
# Stratified lakes and ocean sites
PD_beta_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites,], stree, abund = F)
# Mixed lakes
PD_beta_M_dist<- BAT::beta(surveyed_sites_lake[mixed_lakes,], stree, abund = F)
# Stratified lakes
PD_beta_S_dist<- BAT::beta(surveyed_sites_lake[stratified_lakes,], stree, abund = F)


### Environmental
# Surveyed sites
PD_beta_env_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_env,], stree, abund = F)
# Mixed and stratified lakes
PD_beta_env_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], stree, abund = F)
# Ocean sites and mixed lakes
PD_beta_env_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites_env,], stree, abund = F)
# Stratified lakes and ocean sites
PD_beta_env_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites_env,], stree, abund = F)


### Geographic
# Surveyed sites
PD_beta_geo_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites,], stree, abund = F)
# Mixed and stratified lakes
PD_beta_geo_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], stree, abund = F)
# Ocean sites and mixed lakes
PD_beta_geo_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites,], stree, abund = F)
# Stratified lakes and ocean sites
PD_beta_geo_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites,], stree, abund = F)
```

### PD beta q values of dispersion statistics
```{r, PD beta q-values}
# Sites
N <- 22
# Categories
k <- 3
# Sum of squares
SS <- PD_beta_AOV$`Sum Sq`[2]
# Mean of squares
MS <- PD_beta_AOV$`Mean Sq`[2]
# q value
q.value <- qtukey(p = 0.95, nmeans = k, df = N - k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

group <- PD_beta_THSD$group

(OM_q <- (abs(group[1,1]))/USE)
(SO_q <- (abs(group[2,1]))/USE)
(MS_q <- (abs(group[3,1]))/BSE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx
```

### PD beta phylogenetic dispersions
- The dissimilarity between sites of the same site type
```{r, PD beta dispersion plot}
# Dispersion within-groups
PD_beta_BD_dist <- PD_beta_BD$distances
PD_beta_BD_dist <- as.data.frame(PD_beta_BD_dist)
PD_beta_BD_dist$X <- row.names(PD_beta_BD_dist)
PD_beta_BD_dist_env <- merge(PD_beta_BD_dist, env[surveyed_sites,], by = "X", sort = F)

# Add levels to Site_type column
PD_beta_BD_dist_env$Site_type <- factor(PD_beta_BD_dist_env$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Determine outliers based on dispersion of sites by site type
outlier_PD_beta_BD_dist_env <- PD_beta_BD_dist_env %>%
  group_by(Site_type) %>%
  mutate(
    Q1 = quantile(PD_beta_BD_dist, 0.25),
    Q3 = quantile(PD_beta_BD_dist, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = PD_beta_BD_dist < lower_bound | PD_beta_BD_dist > upper_bound
  )

outlier_PD_beta_BD_dist_env$is_outlier

# Plot dispersion
(PD_beta_BD_dist_env_plot <- ggplot(PD_beta_BD_dist_env, aes(x = Site_type, y = PD_beta_BD_dist, color = Site_type, fill = Site_type)) +
  geom_violin(alpha = 0.6, draw_quantiles = c(0.25, 0.5, 0.75), aes(color = Site_type, fill = Site_type)) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  guides(fill = "none", color = "none") +
  geom_jitter(shape = 21,
            size = 4,
            alpha = 0.8,
            width = 0.1) +
    geom_text_repel(data = PD_beta_BD_dist_env, label = PD_beta_BD_dist_env$X, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme_bw() +
  theme(text = element_text(size = 17), legend.text = element_text(size = 17),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(), axis.text = element_text(size = 17, color = "black"),
    axis.line = element_line(color = "black")) +
  scale_y_continuous(expand = c(0,0.05)) +
  xlab("Site type") +
  ylab("Distance to Centroid") +
  labs(color = "Site_type", tag = "a"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_BD_dist_plot.jpg", PD_beta_BD_dist_env_plot, width = 5.4, height = 6, units = "in")
```

### PD beta NMDS
```{r, PD beta NMDS, results=FALSE}
### Regular
# Reference
PD_beta_ref_NMDS <- metaMDS(PD_beta_ref_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Surveyed sites 
PD_beta_NMDS <- metaMDS(PD_beta_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
round(PD_beta_NMDS$stress, digits = 2)
PD_beta_rep_NMDS <- metaMDS(PD_beta_dist$Brepl, try = 1000, parallel = 4, trymax = 500, maxit = 500)
PD_beta_ric_NMDS <- metaMDS(PD_beta_dist$Brich, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Without LCN
PD_beta_wo_LCN_NMDS <- metaMDS(PD_beta_wo_LCN_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Without TLN and HLM
PD_beta_wo_TLN_HLM_NMDS <- metaMDS(PD_beta_wo_TLN_HLM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
PD_beta_MS_NMDS <- metaMDS(PD_beta_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
PD_beta_OM_NMDS <- metaMDS(PD_beta_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
PD_beta_SO_NMDS <- metaMDS(PD_beta_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed lakes
PD_beta_M_NMDS <- metaMDS(PD_beta_M_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes
PD_beta_S_NMDS <- metaMDS(PD_beta_S_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)


### Environmental
# Surveyed sites 
PD_beta_env_NMDS <- metaMDS(PD_beta_env_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
PD_beta_env_MS_NMDS <- metaMDS(PD_beta_env_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
PD_beta_env_OM_NMDS <- metaMDS(PD_beta_env_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
PD_beta_env_SO_NMDS <- metaMDS(PD_beta_env_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)


### Geographic
# Surveyed sites 
PD_beta_geo_NMDS <- metaMDS(PD_beta_geo_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
PD_beta_geo_MS_NMDS <- metaMDS(PD_beta_geo_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
PD_beta_geo_OM_NMDS <- metaMDS(PD_beta_geo_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
PD_beta_geo_SO_NMDS <- metaMDS(PD_beta_geo_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
```

### Load an envfit analysis function for adjusting p-values
- Need to run this function for downstream analyses in this Markdown file
```{r, envfit p-value function}
# Function p.adjust.envfit
# Calculates adjusted P values for results stored in envfit object,
# created using envfit function from vegan, which fits supplementary variables
# onto axes of unconstrained ordination. 
# Arguments: 
# x - envfit object
# method - method for correction of multiple testing issue (default = 'bonferroni',
#          see ''?p.adjust' for more options)
# n - optional, number of tests for which to correct; if not given, the number is
#          taken as the number of tests conducted by envfit function (for both vectors and factors).
# Author: David Zeleny
p.adjust.envfit <- function (x, method = 'bonferroni', n)
{
  x.new <- x
  if (!is.null (x$vectors)) pval.vectors <- x$vectors$pvals else pval.vectors <- NULL
  if (!is.null (x$factors)) pval.factors <- x$factors$pvals else pval.factors <- NULL
  if (missing (n)) n <- length (pval.vectors) + length (pval.factors)
  if (!is.null (x$vectors)) x.new$vectors$pvals <- p.adjust (x$vectors$pvals, method = method, n = n)
  if (!is.null (x$factors)) x.new$factors$pvals <- p.adjust (x$factors$pvals, method = method, n = n)
  cat ('Adjustment of significance by', method, 'method')
  return (x.new)
}
```

### PD beta varitation partitioning of env and geo variables
```{r}
env_var <- env[surveyed_sites_env,environment]
geo_var <- env[surveyed_sites_env,geography]
PD_beta_varpart <- varpart(PD_beta_env_dist$Btotal, env_var, geo_var)
PD_beta_varpart$part

# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_varpart.jpg", width = 4.5, height = 4.5, units = "in", res = 300, type = "cairo")
# Plot the variation partitioning results
par(mar = c(3, 3, 3, 3) + 1)  # Increase bottom margin if needed
plot(PD_beta_varpart,
     Xnames = c("Env", "Geo"),
     bg = c("mediumpurple", "orange"), alpha = 80,
     digits = 1,
     asp = 1)
# Close the jpg device
dev.off()

plot(PD_beta_varpart,
     Xnames = c("Env", "Geo"),
     bg = c("mediumpurple", "orange"), alpha = 80,
     digits = 1,
     asp = 1)
```

### PD beta env and geo correlated variables using envfit
```{r, PD beta envfit}
### Environmental
# Surveyed sites 
# For figure
(PD_beta_env_ef <- envfit(PD_beta_env_NMDS, env[surveyed_sites_env,"S"], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Site_type"]))
(PD_beta_env_efp <- p.adjust.envfit(PD_beta_env_ef))
# Surveyed sites by Site_type
(PD_beta_env_A_ef <- envfit(PD_beta_env_NMDS, env[surveyed_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Site_type"]))
(PD_beta_env_A_efp <- p.adjust.envfit(PD_beta_env_A_ef))
# Mixed and stratified lakes
(PD_beta_env_MS_ef <- envfit(PD_beta_env_MS_NMDS, env[mixed_stratified_lakes,environment], permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
(PD_beta_env_MS_efp <- p.adjust.envfit(PD_beta_env_MS_ef))
# Ocean sites and mixed lakes
(PD_beta_env_OM_ef <- envfit(PD_beta_env_OM_NMDS, env[ocean_mixed_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Site_type"]))
(PD_beta_env_OM_efp <- p.adjust.envfit(PD_beta_env_OM_ef))
# Stratified lakes and ocean sites
(PD_beta_env_SO_ef <- envfit(PD_beta_env_SO_NMDS, env[ocean_stratified_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Site_type"]))
(PD_beta_env_SO_efp <- p.adjust.envfit(PD_beta_env_SO_ef))


# Mixed lakes
(PD_beta_M_ef <- envfit(PD_beta_M_NMDS, env[mixed_lakes,c("temperature_median","salinity_median","oxygen_median","distance_to_ocean_min_m","max_depth","logArea")], permutations = 999, na.rm = TRUE))
(PD_beta_M_efp <- p.adjust.envfit(PD_beta_M_ef))


# Stratified lakes
(PD_beta_S_ef <- envfit(PD_beta_S_NMDS, env[stratified_lakes,c("temperature_median","salinity_median","oxygen_median","distance_to_ocean_min_m","max_depth","logArea")], permutations = 999, na.rm = TRUE))
(PD_beta_S_efp <- p.adjust.envfit(PD_beta_S_ef))


### Geographic
# Surveyed sites
# For figure
(PD_beta_geo_ef <- envfit(PD_beta_geo_NMDS, env[surveyed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Site_type"]))
(PD_beta_geo_efp <- p.adjust.envfit(PD_beta_geo_ef))
# Surveyed sites by Site_type
(PD_beta_geo_A_ef <- envfit(PD_beta_geo_NMDS, env[surveyed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Site_type"]))
(PD_beta_geo_A_efp <- p.adjust.envfit(PD_beta_geo_A_ef))
# Mixed and stratified lakes
(PD_beta_geo_MS_ef <- envfit(PD_beta_geo_MS_NMDS, env[mixed_stratified_lakes,geography], permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
(PD_beta_geo_MS_efp <- p.adjust.envfit(PD_beta_geo_MS_ef))
# Ocean sites and mixed lakes
(PD_beta_geo_OM_ef <- envfit(PD_beta_geo_OM_NMDS, env[ocean_mixed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Site_type"]))
(PD_beta_geo_OM_efp <- p.adjust.envfit(PD_beta_geo_OM_ef))
# Stratified lakes and ocean sites
(PD_beta_geo_SO_ef <- envfit(PD_beta_geo_SO_NMDS, env[ocean_stratified_sites,geography], permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Site_type"]))
(PD_beta_geo_SO_efp <- p.adjust.envfit(PD_beta_geo_SO_ef))
```

### PD beta Mantel correlation tests
```{r, PD beta Mantel tests}
### Environmental
# Surveyed sites
env_dist_t <- dist(scaled_env[surveyed_sites_env,"temperature_median"], method = "euclidean")
(PD_beta_env_mant_t <- mantel(PD_beta_env_dist$Btotal, env_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Site_type"]))
env_dist_s <- dist(scaled_env[surveyed_sites_env,"salinity_median"], method = "euclidean")
(PD_beta_env_mant_s <- mantel(PD_beta_env_dist$Btotal, env_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Site_type"]))
env_dist_o <- dist(scaled_env[surveyed_sites_env,"oxygen_median"], method = "euclidean")
(PD_beta_env_mant_o <- mantel(PD_beta_env_dist$Btotal, env_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Site_type"]))
# Adjust p-values
PD_beta_env_mant_pv <- rbind(PD_beta_env_mant_t$signif, PD_beta_env_mant_s$signif, PD_beta_env_mant_o$signif)
PD_beta_env_mant_pv <- PD_beta_env_mant_pv[,1]
(PD_beta_env_mant_pv <- p.adjust(PD_beta_env_mant_pv, method = "bonferroni"))

# Mixed and stratified lakes
env_MS_dist_t <- dist(scaled_env[mixed_stratified_lakes,"temperature_median"], method = "euclidean")
(PD_beta_env_MS_mant_t <- mantel(PD_beta_env_MS_dist$Btotal, env_MS_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
env_MS_dist_s <- dist(scaled_env[mixed_stratified_lakes,"salinity_median"], method = "euclidean")
(PD_beta_env_MS_mant_s <- mantel(PD_beta_env_MS_dist$Btotal, env_MS_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
env_MS_dist_o <- dist(scaled_env[mixed_stratified_lakes,"oxygen_median"], method = "euclidean")
(PD_beta_env_MS_mant_o <- mantel(PD_beta_env_MS_dist$Btotal, env_MS_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
# Adjust p-values
PD_beta_env_MS_mant_pv <- rbind(PD_beta_env_MS_mant_t$signif, PD_beta_env_MS_mant_s$signif, PD_beta_env_MS_mant_o$signif)
PD_beta_env_MS_mant_pv <- PD_beta_env_MS_mant_pv[,1]
(PD_beta_env_MS_mant_pv <- p.adjust(PD_beta_env_MS_mant_pv, method = "bonferroni"))

# Ocean sites and mixed lakes
env_OM_dist_t <- dist(scaled_env[ocean_mixed_sites_env,"temperature_median"], method = "euclidean")
(PD_beta_env_OM_mant_t <- mantel(PD_beta_env_OM_dist$Btotal, env_OM_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Site_type"]))
env_OM_dist_s <- dist(scaled_env[ocean_mixed_sites_env,"salinity_median"], method = "euclidean")
(PD_beta_env_OM_mant_s <- mantel(PD_beta_env_OM_dist$Btotal, env_OM_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Site_type"]))
env_OM_dist_o <- dist(scaled_env[ocean_mixed_sites_env,"oxygen_median"], method = "euclidean")
(PD_beta_env_OM_mant_o <- mantel(PD_beta_env_OM_dist$Btotal, env_OM_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Site_type"]))
# Adjust p-values
PD_beta_env_OM_mant_pv <- rbind(PD_beta_env_OM_mant_t$signif, PD_beta_env_OM_mant_s$signif, PD_beta_env_OM_mant_o$signif)
PD_beta_env_OM_mant_pv <- PD_beta_env_OM_mant_pv[,1]
(PD_beta_env_OM_mant_pv <- p.adjust(PD_beta_env_OM_mant_pv, method = "bonferroni"))

# Stratified lakes and ocean sites
env_SO_dist_t <- dist(scaled_env[ocean_stratified_sites_env,"temperature_median"], method = "euclidean")
(PD_beta_env_SO_mant_t <- mantel(PD_beta_env_SO_dist$Btotal, env_SO_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Site_type"]))
env_SO_dist_s <- dist(scaled_env[ocean_stratified_sites_env,"salinity_median"], method = "euclidean")
(PD_beta_env_SO_mant_s <- mantel(PD_beta_env_SO_dist$Btotal, env_SO_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Site_type"]))
env_SO_dist_o <- dist(scaled_env[ocean_stratified_sites_env,"oxygen_median"], method = "euclidean")
(PD_beta_env_SO_mant_o <- mantel(PD_beta_env_SO_dist$Btotal, env_SO_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Site_type"]))
# Adjust p-values
PD_beta_env_SO_mant_pv <- rbind(PD_beta_env_SO_mant_t$signif, PD_beta_env_SO_mant_s$signif, PD_beta_env_SO_mant_o$signif)
PD_beta_env_SO_mant_pv <- PD_beta_env_SO_mant_pv[,1]
(PD_beta_env_SO_mant_pv <- p.adjust(PD_beta_env_SO_mant_pv, method = "bonferroni"))


# Mixed lakes
env_M_dist_t <- dist(scaled_env[mixed_lakes,"temperature_median"], method = "euclidean")
(PD_beta_M_mant_t <- mantel(PD_beta_M_dist$Btotal, env_M_dist_t, method = "spearman", permutations = 999, na.rm = TRUE))
env_M_dist_s <- dist(scaled_env[mixed_lakes,"salinity_median"], method = "euclidean")
(PD_beta_M_mant_s <- mantel(PD_beta_M_dist$Btotal, env_M_dist_s, method = "spearman", permutations = 999, na.rm = TRUE))
env_M_dist_o <- dist(scaled_env[mixed_lakes,"oxygen_median"], method = "euclidean")
(PD_beta_M_mant_o <- mantel(PD_beta_M_dist$Btotal, env_M_dist_o, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_dm <- dist(scaled_env[mixed_lakes,"distance_to_ocean_min_m"], method = "euclidean")
(PD_beta_M_mant_dm <- mantel(PD_beta_M_dist$Btotal, geo_M_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_md <- dist(scaled_env[mixed_lakes,"max_depth"], method = "euclidean")
(PD_beta_M_mant_md <- mantel(PD_beta_M_dist$Btotal, geo_M_dist_md, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_la <- dist(scaled_env[mixed_lakes,"logArea"], method = "euclidean")
(PD_beta_M_mant_la <- mantel(PD_beta_M_dist$Btotal, geo_M_dist_la, method = "spearman", permutations = 999, na.rm = TRUE))
# Adjust p-values
PD_beta_M_mant_pv <- rbind(PD_beta_M_mant_t$signif, PD_beta_M_mant_s$signif, PD_beta_M_mant_o$signif, PD_beta_M_mant_dm$signif, PD_beta_M_mant_md$signif, PD_beta_M_mant_la$signif)
PD_beta_M_mant_pv <- PD_beta_M_mant_pv[,1]
(PD_beta_M_mant_pv <- p.adjust(PD_beta_M_mant_pv, method = "bonferroni"))


# Stratified lakes
env_S_dist_t <- dist(scaled_env[stratified_lakes,"temperature_median"], method = "euclidean")
(PD_beta_S_mant_t <- mantel(PD_beta_S_dist$Btotal, env_S_dist_t, method = "spearman", permutations = 999, na.rm = TRUE))
env_S_dist_s <- dist(scaled_env[stratified_lakes,"salinity_median"], method = "euclidean")
(PD_beta_S_mant_s <- mantel(PD_beta_S_dist$Btotal, env_S_dist_s, method = "spearman", permutations = 999, na.rm = TRUE))
env_S_dist_o <- dist(scaled_env[stratified_lakes,"oxygen_median"], method = "euclidean")
(PD_beta_S_mant_o <- mantel(PD_beta_S_dist$Btotal, env_S_dist_o, method = "spearman", permutations = 999, na.rm = TRUE))
geo_S_dist_dm <- dist(scaled_env[stratified_lakes,"distance_to_ocean_min_m"], method = "euclidean")
(PD_beta_S_mant_dm <- mantel(PD_beta_S_dist$Btotal, geo_S_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE))
geo_S_dist_md <- dist(scaled_env[stratified_lakes,"max_depth"], method = "euclidean")
(PD_beta_S_mant_md <- mantel(PD_beta_S_dist$Btotal, geo_S_dist_md, method = "spearman", permutations = 999, na.rm = TRUE))
geo_S_dist_la <- dist(scaled_env[stratified_lakes,"logArea"], method = "euclidean")
(PD_beta_S_mant_la <- mantel(PD_beta_S_dist$Btotal, geo_S_dist_la, method = "spearman", permutations = 999, na.rm = TRUE))
# Adjust p-values
PD_beta_S_mant_pv <- rbind(PD_beta_S_mant_t$signif, PD_beta_S_mant_s$signif, PD_beta_S_mant_o$signif, PD_beta_S_mant_dm$signif, PD_beta_S_mant_md$signif, PD_beta_S_mant_la$signif)
PD_beta_S_mant_pv <- PD_beta_S_mant_pv[,1]
(PD_beta_S_mant_pv <- p.adjust(PD_beta_S_mant_pv, method = "bonferroni"))


### Geographic
# Surveyed sites
geo_dist_dm <- dist(scaled_env[surveyed_sites,"distance_to_ocean_min_m"], method = "euclidean")
(PD_beta_geo_mant_dm <- mantel(PD_beta_geo_dist$Btotal, geo_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Site_type"]))
geo_dist_md <- dist(scaled_env[surveyed_sites,"max_depth"], method = "euclidean")
(PD_beta_geo_mant_md <- mantel(PD_beta_geo_dist$Btotal, geo_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Site_type"]))
geo_dist_la <- dist(scaled_env[surveyed_sites,"logArea"], method = "euclidean")
(PD_beta_geo_mant_la <- mantel(PD_beta_geo_dist$Btotal, geo_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Site_type"]))
# Adjust p-values
PD_beta_geo_mant_pv <- rbind(PD_beta_geo_mant_dm$signif, PD_beta_geo_mant_md$signif, PD_beta_geo_mant_la$signif)
PD_beta_geo_mant_pv <- PD_beta_geo_mant_pv[,1]
(PD_beta_geo_mant_pv <- p.adjust(PD_beta_geo_mant_pv, method = "bonferroni"))

# Mixed and stratified lakes 
geo_MS_dist_dm <- dist(scaled_env[mixed_stratified_lakes,"distance_to_ocean_min_m"], method = "euclidean")
(PD_beta_geo_MS_mant_dm <- mantel(PD_beta_geo_MS_dist$Btotal, geo_MS_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
geo_MS_dist_md <- dist(scaled_env[mixed_stratified_lakes,"max_depth"], method = "euclidean")
(PD_beta_geo_MS_mant_md <- mantel(PD_beta_geo_MS_dist$Btotal, geo_MS_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
geo_MS_dist_la <- dist(scaled_env[mixed_stratified_lakes,"logArea"], method = "euclidean")
(PD_beta_geo_MS_mant_la <- mantel(PD_beta_geo_MS_dist$Btotal, geo_MS_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Site_type"]))
# Adjust p-values
PD_beta_geo_MS_mant_pv <- rbind(PD_beta_geo_MS_mant_dm$signif, PD_beta_geo_MS_mant_md$signif, PD_beta_geo_MS_mant_la$signif)
PD_beta_geo_MS_mant_pv <- PD_beta_geo_MS_mant_pv[,1]
(PD_beta_geo_MS_mant_pv <- p.adjust(PD_beta_geo_MS_mant_pv, method = "bonferroni"))

# Ocean sites and mixed lakes
geo_OM_dist_dm <- dist(scaled_env[ocean_mixed_sites,"distance_to_ocean_min_m"], method = "euclidean")
(PD_beta_geo_OM_mant_dm <- mantel(PD_beta_geo_OM_dist$Btotal, geo_OM_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Site_type"]))
geo_OM_dist_md <- dist(scaled_env[ocean_mixed_sites,"max_depth"], method = "euclidean")
(PD_beta_geo_OM_mant_md <- mantel(PD_beta_geo_OM_dist$Btotal, geo_OM_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Site_type"]))
geo_OM_dist_la <- dist(scaled_env[ocean_mixed_sites,"logArea"], method = "euclidean")
(PD_beta_geo_OM_mant_la <- mantel(PD_beta_geo_OM_dist$Btotal, geo_OM_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Site_type"]))
# Adjust p-values
PD_beta_geo_OM_mant_pv <- rbind( PD_beta_geo_OM_mant_dm$signif, PD_beta_geo_OM_mant_md$signif, PD_beta_geo_OM_mant_la$signif)
PD_beta_geo_OM_mant_pv <- PD_beta_geo_OM_mant_pv[,1]
(PD_beta_geo_OM_mant_pv <- p.adjust(PD_beta_geo_OM_mant_pv, method = "bonferroni"))

# Stratified lakes and ocean sites
geo_SO_dist_dm <- dist(scaled_env[ocean_stratified_sites,"distance_to_ocean_min_m"], method = "euclidean")
(PD_beta_geo_SO_mant_dm <- mantel(PD_beta_geo_SO_dist$Btotal, geo_SO_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Site_type"]))
geo_SO_dist_md <- dist(scaled_env[ocean_stratified_sites,"max_depth"], method = "euclidean")
(PD_beta_geo_SO_mant_md <- mantel(PD_beta_geo_SO_dist$Btotal, geo_SO_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Site_type"]))
geo_SO_dist_la <- dist(scaled_env[ocean_stratified_sites,"logArea"], method = "euclidean")
(PD_beta_geo_SO_mant_la <- mantel(PD_beta_geo_SO_dist$Btotal, geo_SO_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Site_type"]))
# Adjust p-values
PD_beta_geo_SO_mant_pv <- rbind( PD_beta_geo_SO_mant_dm$signif, PD_beta_geo_SO_mant_md$signif, PD_beta_geo_SO_mant_la$signif)
PD_beta_geo_SO_mant_pv <- PD_beta_geo_SO_mant_pv[,1]
(PD_beta_geo_SO_mant_pv <- p.adjust(PD_beta_geo_SO_mant_pv, method = "bonferroni"))
```

### PD beta NMDS ordination plots
```{r, PD beta NMDS plots}
# PD beta total NMDS scores
PD_beta_NMDS_data.scores <- as.data.frame(scores(PD_beta_NMDS))
PD_beta_NMDS_data.scores$Site_type <- env[surveyed_sites,"Site_type"]
PD_beta_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
PD_beta_NMDS_data.scores$Site_type <- factor(PD_beta_NMDS_data.scores$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Get significantly correlated environmental variables
PD_beta_env_ef_coord_cont <- as.data.frame(scores(PD_beta_env_ef, "vectors")) * ordiArrowMul(PD_beta_env_ef)
PD_beta_env_row_names <- c("S")
# Assign the new row names to the data frame
PD_beta_env_ef_coord_cont <- data.frame(row.names = PD_beta_env_row_names, PD_beta_env_ef_coord_cont)

# PD_beta_geo_ef_coord_cont <- as.data.frame(scores(PD_beta_geo_ef, "vectors")) * ordiArrowMul(PD_beta_geo_ef)
# PD_beta_geo_row_names <- c("minD")
# # Assign the new row names to the data frame
# PD_beta_geo_ef_coord_cont <- data.frame(row.names = PD_beta_geo_row_names, PD_beta_geo_ef_coord_cont)

# Plot NMDS ordination of PD beta total with CI = 0.95
PD_beta_ef_plot <- ggplot(data = PD_beta_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.95) +
  geom_point(data = PD_beta_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               data = PD_beta_env_ef_coord_cont, linewidth =1, alpha = 0.2, color = env_cont) +
  geom_text(data = PD_beta_env_ef_coord_cont, aes(x = NMDS1, y = NMDS2),
            color = env_cont, label = row.names(PD_beta_env_ef_coord_cont), size = 5) +
  geom_text_repel(data = PD_beta_NMDS_data.scores, label = PD_beta_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 16), 
        legend.position = "bottom", 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 16), 
        legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(PD_beta_ef_plot <- PD_beta_ef_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_NMDS.jpg", PD_beta_ef_plot, width = 6, height = 6, units = "in")


# Plot NMDS ordination of PD beta total with CI = 0.90
PD_beta_ef_plot_CI90 <- ggplot(data = PD_beta_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.90) +
  geom_point(data = PD_beta_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               data = PD_beta_env_ef_coord_cont, linewidth =1, alpha = 0.2, color = env_cont) +
  geom_text(data = PD_beta_env_ef_coord_cont, aes(x = NMDS1, y = NMDS2),
            color = env_cont, label = row.names(PD_beta_env_ef_coord_cont), size = 5) +
  geom_text_repel(data = PD_beta_NMDS_data.scores, label = PD_beta_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 17), 
        legend.position = "bottom", 
        legend.title = element_text(size = 17), 
        legend.text = element_text(size = 17), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 17), 
        legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(PD_beta_ef_plot_CI90 <- PD_beta_ef_plot_CI90 + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_NMDS_CI90.jpg", PD_beta_ef_plot_CI90, width = 5.4, height = 6, units = "in")


# PD beta total NMDS scores without LCN
PD_beta_wo_LCN_NMDS_data.scores <- as.data.frame(scores(PD_beta_wo_LCN_NMDS))
PD_beta_wo_LCN_NMDS_data.scores$Site_type <- env[surveyed_sites_wo_LCN,"Site_type"]
PD_beta_wo_LCN_NMDS_data.scores$Lakes <- env[surveyed_sites_wo_LCN,1]
PD_beta_wo_LCN_NMDS_data.scores$Site_type <- factor(PD_beta_wo_LCN_NMDS_data.scores$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of PD beta total without LCN with CI = 0.95
PD_beta_wo_LCN_plot <- ggplot(data = PD_beta_wo_LCN_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.95) +
  geom_point(data = PD_beta_wo_LCN_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = PD_beta_wo_LCN_NMDS_data.scores, label = PD_beta_wo_LCN_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 14), 
        legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_wo_LCN_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(PD_beta_wo_LCN_plot <- PD_beta_wo_LCN_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_NMDS_wo_LCN.jpg", PD_beta_wo_LCN_plot, width = 4.88, height = 6, units = "in")


# PD beta total NMDS scores without TLN and HLM
PD_beta_wo_TLN_HLM_NMDS_data.scores <- as.data.frame(scores(PD_beta_wo_TLN_HLM_NMDS))
PD_beta_wo_TLN_HLM_NMDS_data.scores$Site_type <- env[surveyed_sites_wo_TLN_HLM,"Site_type"]
PD_beta_wo_TLN_HLM_NMDS_data.scores$Lakes <- env[surveyed_sites_wo_TLN_HLM,1]
PD_beta_wo_TLN_HLM_NMDS_data.scores$Site_type <- factor(PD_beta_wo_TLN_HLM_NMDS_data.scores$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of PD beta total without TLN and HLM with CI = 0.95
PD_beta_wo_TLN_HLM_plot <- ggplot(data = PD_beta_wo_TLN_HLM_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.95) +
  geom_point(data = PD_beta_wo_TLN_HLM_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = PD_beta_wo_TLN_HLM_NMDS_data.scores, label = PD_beta_wo_TLN_HLM_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 14), 
        legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_wo_TLN_HLM_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(PD_beta_wo_TLN_HLM_plot <- PD_beta_wo_TLN_HLM_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_NMDS_wo_TLN_HLM.jpg", PD_beta_wo_TLN_HLM_plot, width = 4.88, height = 6, units = "in")


# PD beta replacement NMDS scores
PD_beta_rep_NMDS_data.scores <- as.data.frame(scores(PD_beta_rep_NMDS))
PD_beta_rep_NMDS_data.scores$Site_type <- env[surveyed_sites,"Site_type"]
PD_beta_rep_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
PD_beta_rep_NMDS_data.scores$Site_type <- factor(PD_beta_rep_NMDS_data.scores$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of PD beta replacement with CI = 0.95
PD_beta_rep_plot <- ggplot(data = PD_beta_rep_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.95) +
  geom_point(data = PD_beta_rep_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = PD_beta_rep_NMDS_data.scores, label = PD_beta_rep_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 14), 
        legend.key = element_blank()) +
  annotate("text", x = -0.8, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_rep_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(PD_beta_rep_plot <- PD_beta_rep_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_rep_NMDS.jpg", PD_beta_rep_plot, width = 4.88, height = 6, units = "in")


# PD beta richness NMDS scores
PD_beta_ric_NMDS_data.scores <- as.data.frame(scores(PD_beta_ric_NMDS))
PD_beta_ric_NMDS_data.scores$Site_type <- env[surveyed_sites,"Site_type"]
PD_beta_ric_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
PD_beta_ric_NMDS_data.scores$Site_type <- factor(PD_beta_ric_NMDS_data.scores$Site_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of PD beta richness with CI = 0.95
PD_beta_ric_plot <- ggplot(data = PD_beta_ric_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.95) +
  geom_point(data = PD_beta_ric_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = PD_beta_ric_NMDS_data.scores, label = PD_beta_ric_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 14), 
        legend.key = element_blank()) +
  annotate("text", x = -0.8, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_ric_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(PD_beta_ric_plot <- PD_beta_ric_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_ric_NMDS.jpg", PD_beta_ric_plot, width = 4.88, height = 6, units = "in")


# PD beta ref NMDS scores
PD_beta_ref_NMDS_data.scores <- as.data.frame(scores(PD_beta_ref_NMDS))
PD_beta_ref_NMDS_data.scores$Site_type <- env[,"Site_type"]
PD_beta_ref_NMDS_data.scores$Lakes <- env[,1]
PD_beta_ref_NMDS_data.scores$Site_type <- factor(PD_beta_ref_NMDS_data.scores$Site_type, levels = c("Reference", "Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of PD beta ref with CI = 0.95
PD_beta_ref_plot <- ggplot(data = PD_beta_ref_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Site_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Site_type), type = "t", level = 0.95) +
  geom_point(data = PD_beta_ref_NMDS_data.scores, aes(color = Site_type, fill = Site_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = PD_beta_ref_NMDS_data.scores, label = PD_beta_ref_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "right", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 14), 
        legend.key = element_blank()) +
  annotate("text", x = -0.8, y = 0.5, size = 5,
           label = paste("Stress: ", round(PD_beta_ref_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(PD_beta_ref_plot <- PD_beta_ref_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_ref_NMDS.jpg", PD_beta_ref_plot, width = 6.26, height = 6, units = "in")
```

### PD beta dendrogram
```{r, PD beta dendrogram}
# Cluster communities using average-linkage algorithm
PD_beta_dist_clust <- hclust(PD_beta_dist$Btotal, method = "average")

# The following will create a dendrogram of PD_beta putting things in one dimensional space
# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_dendrogram.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

# Create your plot using the plot() function
plot(PD_beta_dist_clust, 
     xlab = "Sites",
     ylab = "PD beta",
     main = "",
     sub = "",
     pch = 20,
     col= "black")

# Close the jpg device
dev.off()


# The following will graph and identify the best supported number of clusters of the dendrogram
# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_dg_silhouette.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

Si <- numeric(nrow(presabs_lake[surveyed_sites,]))
for (k in 2:(nrow(presabs_lake[surveyed_sites,])-1))
{
  sil<-silhouette(cutree(PD_beta_dist_clust, k=k), PD_beta_dist$Btotal)
  Si[k] <- summary(sil)$avg.width
}
k.best<-which.max(Si)
plot(1:nrow(presabs_lake[surveyed_sites,]), Si, type = "h", main = "Silhouette", xlab = "K", ylab = "Width")
axis(1, k.best, paste("optimum", k.best, sep="\n"), col = "red", font = 2, col.axis = "red")

# Close the jpg device
dev.off()


# Partitioning around mediods uses distances matrix to found how points according to silhouette are grouped around a centroid. The greater the number, the greater the distance from the centroid.
# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/PD_beta_dg_partitioning.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

PD_beta_pam <- pam(PD_beta_dist$Btotal,2)
plot(PD_beta_pam)

# Close the jpg device
dev.off()
```

### PD beta outliers
```{r, PD beta outliers}
# Determine NMDS1 outliers
ordered_PD_beta_NMDS1_data.scores <- PD_beta_NMDS_data.scores[order(PD_beta_NMDS_data.scores$NMDS1), ]

outlier_PD_beta_NMDS1 <- ordered_PD_beta_NMDS1_data.scores %>%
  group_by(Site_type) %>%
  mutate(
    Q1 = quantile(NMDS1, 0.25),
    Q3 = quantile(NMDS1, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = NMDS1 < lower_bound | NMDS1 > upper_bound
  )

outlier_PD_beta_NMDS1 <- as.data.frame(outlier_PD_beta_NMDS1)

row.names(outlier_PD_beta_NMDS1) <- outlier_PD_beta_NMDS1$Lakes

# Create the plot
(outlier_PD_beta_NMDS1_plot <- ggplot(outlier_PD_beta_NMDS1, aes(x = Site_type, y = NMDS1, fill = Site_type)) +
  geom_violin(alpha = 0.9, draw_quantiles = c(0.25, 0.5, 0.75), aes(fill = Site_type)) +
  geom_jitter(aes(color = is_outlier), width = 0.1, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = outlier_PD_beta_NMDS1, label = outlier_PD_beta_NMDS1$Lakes, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme(text = element_text(size = 16),
        legend.position = "bottom",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  scale_y_continuous(expand = c(0,0.05)) +
  guides(fill = "none") + 
  labs(y = "NMDS1 Distances", x = "Site type", color = "Outlier:", tag = "a"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/outlier_PD_beta_NMDS1.jpg", outlier_PD_beta_NMDS1_plot, width = 4.88, height = 6, units = "in")


# Determine NMDS2 outliers
ordered_PD_beta_NMDS2_data.scores <- PD_beta_NMDS_data.scores[order(PD_beta_NMDS_data.scores$NMDS2), ]

outlier_PD_beta_NMDS2 <- ordered_PD_beta_NMDS2_data.scores %>%
  group_by(Site_type) %>%
  mutate(
    Q1 = quantile(NMDS2, 0.25),
    Q3 = quantile(NMDS2, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = NMDS2 < lower_bound | NMDS2 > upper_bound
  )

outlier_PD_beta_NMDS2 <- as.data.frame(outlier_PD_beta_NMDS2)

row.names(outlier_PD_beta_NMDS2) <- outlier_PD_beta_NMDS2$Lakes

# Create the plot
(outlier_PD_beta_NMDS2_plot <- ggplot(outlier_PD_beta_NMDS2, aes(x = Site_type, y = NMDS2, fill = Site_type)) +
  geom_violin(alpha = 0.9, draw_quantiles = c(0.25, 0.5, 0.75), aes(fill = Site_type)) +
  geom_jitter(aes(color = is_outlier), width = 0.1, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = outlier_PD_beta_NMDS2, label = outlier_PD_beta_NMDS2$Lakes, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme(text = element_text(size = 16),
        legend.position = "bottom",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  scale_y_continuous(expand = c(0,0.05)) +
  guides(fill = "none") + 
  labs(y = "NMDS2 Distances", x = "Site type", color = "Outlier:", tag = "b"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/outlier_PD_beta_NMDS2.jpg", outlier_PD_beta_NMDS2_plot, width = 4.88, height = 6, units = "in")
```

### Package and version info
```{r, session info}
sessionInfo()
```

#### stree Z score vs p value plots
```{r}
stree_sesmpd_plot <- ggplot(data = stree_sesmpd_env, mapping = aes(y = mpd.obs.p, x = mpd.obs.z, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
    size = 4,
    alpha = 1) + 
  geom_text_repel(label = stree_sesmpd_env[,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(y= "p-value", x= "z-score", colour = "Site type:", fill = "Site type:", tag = "D") +
  ylim(c(0,1)) +
  scale_x_continuous(breaks = c(-6,-4, -2, 0, 2), limits = c(-7,3)) +
  annotate('rect', ymin=0, ymax=1, xmin=-2, xmax=2, alpha = 0.3, fill='grey')
stree_sesmpd_plot <- stree_sesmpd_plot + guides(color = guide_legend(override.aes = list(label = "")))
stree_sesmpd_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/stree_sesmpd_plot_z.jpg", stree_sesmpd_plot, width = 5, height = 4, units = "in")

stree_sesmntd_plot <- ggplot(data = stree_sesmntd_env, mapping = aes(y = mntd.obs.p, x = mntd.obs.z, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
    size = 5,
    alpha = 1) + 
  geom_text_repel(label = stree_sesmntd_env[,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(y= "p-value", x= "z-score", colour = "Site type:", fill = "Site type:", tag = "E") +
  ylim(c(0,1)) +
  scale_x_continuous(breaks = c(-6,-4, -2, 0, 2), limits = c(-7,3)) +
  annotate('rect', ymin=0, ymax=1, xmin=-2, xmax=2, alpha = 0.3, fill='grey')
stree_sesmntd_plot <- stree_sesmntd_plot + guides(color = guide_legend(override.aes = list(label = "")))
stree_sesmntd_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/stree_sesmntd_plot_z.jpg", stree_sesmntd_plot, width = 5, height = 4, units = "in")

stree_sespd_plot <- ggplot(data = stree_sespd_env, mapping = aes(y = pd.obs.p, x = pd.obs.z, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
    size = 5,
    alpha = 1) + 
  geom_text_repel(label = stree_sespd_env[,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(y= "p-value", x= "z-score", colour = "Site type:", fill = "Site type:", tag = "F") +
  ylim(c(0,1)) +
  scale_x_continuous(breaks = c(-6,-4, -2, 0, 2), limits = c(-7,3)) +
  annotate('rect', ymin=0, ymax=1, xmin=-2, xmax=2, alpha = 0.3, fill='grey')
stree_sespd_plot <- stree_sespd_plot + guides(color = guide_legend(override.aes = list(label = "")))
stree_sespd_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/stree_sespd_plot_z.jpg", stree_sespd_plot, width = 5, height = 4, units = "in")
```

#### stree Z score vs obs value plots
```{r}
stree_sesmpd_plot <- ggplot(data = stree_sesmpd_env, mapping = aes(y = mpd.obs, x = mpd.obs.z, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
    size = 5,
    alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(y= "Observed MPD", x= "z-score", colour = "Site type:", fill = "Site type:") +
  ylim(c(0,1)) +
  scale_x_continuous(breaks = c(-6,-4, -2, 0, 2), limits = c(-7,3)) +
  annotate('rect', ymin=0, ymax=1, xmin=-2, xmax=2, alpha = 0.3, fill='grey')
stree_sesmpd_plot <- stree_sesmpd_plot + guides(color = guide_legend(override.aes = list(label = "")))
stree_sesmpd_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/stree_sesmpd_plot_obs+z.jpg", stree_sesmpd_plot, width = 5, height = 4, units = "in")

stree_sesmntd_plot <- ggplot(data = stree_sesmntd_env, mapping = aes(y = mntd.obs, x = mntd.obs.z, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
    size = 5,
    alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(y= "Observed MNTD", x= "z-score", colour = "Site type:", fill = "Site type:") +
  ylim(c(0,1)) +
  scale_x_continuous(breaks = c(-6,-4, -2, 0, 2), limits = c(-7,3)) +
  annotate('rect', ymin=0, ymax=1, xmin=-2, xmax=2, alpha = 0.3, fill='grey')
stree_sesmntd_plot <- stree_sesmntd_plot + guides(color = guide_legend(override.aes = list(label = "")))
stree_sesmntd_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/stree_sesmntd_plot_obs+z.jpg", stree_sesmntd_plot, width = 5, height = 4, units = "in")

stree_sespd_plot <- ggplot(data = stree_sespd_env, mapping = aes(y = pd.obs, x = pd.obs.z, color = Site_type, fill = Site_type)) + 
  geom_point(stat = 'identity',
    size = 5,
    alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(y= "Observed PD", x= "z-score", colour = "Site type:", fill = "Site type:") +
  ylim(c(0,6)) +
  scale_x_continuous(breaks = c(-6,-4, -2, 0, 2), limits = c(-7,3)) +
  annotate('rect', ymin=0, ymax=6, xmin=-2, xmax=2, alpha = 0.3, fill='grey')
stree_sespd_plot <- stree_sespd_plot + guides(color = guide_legend(override.aes = list(label = "")))
stree_sespd_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/PD/stree_sespd_plot_obs+z.jpg", stree_sespd_plot, width = 5, height = 4, units = "in")
```

### Fourth corner approach
```{r, eval=F}
library(mvabund)
library(lattice)
rows_with_na <- which(rowSums(is.na(stree_vif)) > 0)
rows_with_na

ft1=traitglm(surveyed_sites_lake[surveyed_sites_env,], env[surveyed_sites_env,environment], stree_vif[,c(1:13)])
ft1$fourth

a = max(abs(ft1$fourth.corner))
colort = colorRampPalette(c("blue","purple","red")) 
plot.4th = levelplot(t(as.matrix(ft1$fourth.corner)), xlab="Environmental Variables",
                     ylab="Species traits", col.regions=colort(100), at=seq(-a, a, length=100),
                     scales = list(x = list(rot = 45)))
print(plot.4th)

ft <- fourthcorner(env[surveyed_sites_env,environment], surveyed_sites_lake[surveyed_sites_env,], stree_vif[,c(1:5,7:10,12:13)], tr01 = T,modeltype=6)

stree_comp <- stree_vif[complete.cases(stree_vif),]
ssl_comp <- surveyed_sites_lake[,complete.cases(stree_vif)]

ft <- fourthcorner(env[surveyed_sites_env,environment], ssl_comp[surveyed_sites_env,], stree_comp, tr01 = T,modeltype=6)

ft <- fourthcorner(env[surveyed_sites_env,environment], surveyed_sites_lake[surveyed_sites_env,], stree_vif[,c(1:5,7:10,12:13)], tr01 = T,modeltype=6)

stree_comp <- stree_vif[complete.cases(stree_vif),]
ssl_comp <- surveyed_sites_lake[,complete.cases(stree_vif)]

ft <- fourthcorner(env[surveyed_sites_env,environment], ssl_comp[surveyed_sites_env,], stree_comp, tr01 = T,modeltype=6)
```
