---
title: "Species diversity analyses"
output: github_document
editor_options: 
  chunk_output_type: console
---
### Load packages and files
```{r, load packages and files}
# Load the knitr package if not already loaded
library(knitr)

# Source the R Markdown file
knit("/Users/bailey/Documents/research/Fish_Biodiversity/src/collection/load_collection_data.Rmd", output = "/Users/bailey/Documents/research/Fish_Biodiversity/src/collection/load_collection_data.md")

# Packages
library(picante)
library(vegan)
library(phytools)
library(ggplot2)
library(ggrepel)
library(caret)
library(BAT)
library(pairwiseAdonis)
library(ggvenn)
library(dplyr)
library(emmeans)
library(car)

# Site vectors
surveyed_sites <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OCM", "OCO", "OLO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_LCN <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OCM", "OCO", "OLO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_OCO <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OCM", "OLO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_TLN_HLM <- c("BCM", "CLM", "FLK", "GLK", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OCM", "OCO", "OLO", "OTM", "RCA", "SLN", "ULN")

surveyed_sites_env <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OTM", "RCA", "SLN", "TLN", "ULN")

ocean_mixed_sites <- c("FLK", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLN", "NLU", "OCM", "OCO", "OLO", "RCA", "ULN")

ocean_mixed_sites_env <- c("FLK", "HLO", "IBK", "LLN", "MLN", "NCN", "NLN", "NLU", "OLO", "RCA", "ULN")

ocean_stratified_sites <- c("BCM", "CLM", "GLK", "HLM", "IBK", "LCN", "NCN", "NLK", "OCM", "OCO", "OTM", "RCA", "SLN", "TLN")

ocean_stratified_sites_env <- c("BCM", "CLM", "GLK", "HLM", "IBK", "NCN", "NLK", "OTM", "RCA", "SLN", "TLN")

mixed_stratified_lakes <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "LLN", "MLN", "NLK", "NLN", "NLU", "OLO", "OTM", "SLN", "TLN", "ULN")

ocean_sites <- c("IBK", "LCN", "NCN", "OCM", "OCO", "RCA")

ocean_sites_env <- c("IBK", "NCN", "RCA")

mixed_lakes <- c("FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

stratified_lakes <- c("BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

# Define your custom colors
custom_colors <- c("Reference" = "black", "Ocean" = "#EE6363", "Mixed" = "#87CEFA", "Stratified" = "#6E8B3D")

env_cont <- "#FFB90F"

set.seed(123)
```

### Edit input files
```{r, edit files}
strat_fish <- stratified_fish[,-c(1,10)]
mix_fish <- mixed_fish[,-c(1,10)]
oc_fish <- ocean_fish[,-c(1,8)]

env <- env[order(env$X),]
presabs_lake <- presabs_lake[order(row.names(presabs_lake)),]
surveyed_sites_lake <- surveyed_sites_lake[order(row.names(surveyed_sites_lake)),]

Location_type_group_ref <- env[,"Location_type"]
Location_type_group <- env[surveyed_sites,"Location_type"]
```

# Species Diversity
## SD alpha diversity
### SD alpha bar plot
```{r, SD alpha bar plot}
#Plot species richness with a bar plot
SR_env$Location_type <- factor(SR_env$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

(SR_plot <- ggplot(data = SR_env[-20,], mapping = aes(x = reorder(X, row_sum, decreasing = T), y = row_sum, color = Location_type, fill = Location_type, )) + 
  geom_bar(stat = 'identity',
  alpha = 1) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 14), 
  legend.title = element_text(size = 14), 
  legend.text = element_text(size = 14),
  axis.text = element_text(size = 14, color = "black"),
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
  axis.line = element_line(color = "black"),
  plot.background = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Site", y="Species Richness", colour = "Location type:", fill = "Location type:"))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SR_plot.jpg", plot = SR_plot, width = 6, height = 6, units = "in")
```

### SD alpha venn diagram plot
```{r, SD alpha venn diagram plot}
Location_type_pres <- strat_presabs_lake[c(24:26),]
Location_type_pres <- t(Location_type_pres)
Location_type_pres <- Location_type_pres[which(rowSums(Location_type_pres) > 0),]

# Convert numerical values to logical
Location_type_pres_logical <- as.data.frame(Location_type_pres > 0)

# Check the structure of the transformed data
str(Location_type_pres_logical)

# Create the Venn diagram
(venn_plot <- ggvenn(Location_type_pres_logical,
        show_percentage = F,
        fill_color = c("#EE6363", "#87CEFA", "#6E8B3D"),
        fill_alpha = 0.7,
        stroke_alpha = 0,
        stroke_size = 0.5, 
        set_name_size = 5,
        text_size = 5))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/venn_plot.jpg", plot = venn_plot, width = 4.25, height = 4, units = "in")
```

### SD alpha subsample venn diagram plots
```{r, SD alpha venn diagram plots}
# Randomly sample six columns from stratified_fish and mixed_fish
sampled_stratified <- sample(names(strat_fish)[-1], 6)
sampled_mixed <- sample(names(mix_fish)[-1], 6)

# Select sampled columns from stratified_fish and mixed_fish
stratified_subset <- select(stratified_fish, all_of(sampled_stratified))
mixed_subset <- select(mixed_fish, all_of(sampled_mixed))

# Get rid of species not present in the 6 remaining sites
stratified_subset <- stratified_subset[which(rowSums(stratified_subset[, -1]) > 0),]
mixed_subset <- mixed_subset[which(rowSums(mixed_subset[, -1]) > 0),]

# Make Location type columns in each dataframe
stratified_subset[,7] <- rowSums(stratified_subset[,c(1:6)])
colnames(stratified_subset)[colnames(stratified_subset) == "V7"] <- "Stratified lakes"

mixed_subset[,7] <- rowSums(mixed_subset[,c(1:6)])
colnames(mixed_subset)[colnames(mixed_subset) == "V7"] <- "Mixed lakes"

oc_fish[,7] <- rowSums(oc_fish[,c(1:6)])
colnames(oc_fish)[colnames(oc_fish) == "V7"] <- "Ocean sites"

# Merge the two data frames, replacing missing values with 0
oc_mix_merge <- merge(oc_fish, mixed_subset, by = 0, all = TRUE)
row.names(oc_mix_merge) <- oc_mix_merge$Row.names
oc_mix_merge <- oc_mix_merge[,-1]
oc_mix_strat_merge <- merge(oc_mix_merge, stratified_subset, by = 0, all = TRUE)
row.names(oc_mix_strat_merge) <- oc_mix_strat_merge$Row.names
oc_mix_strat_merge <- oc_mix_strat_merge[,-1]

# Keep only Location type
keep <- c("Ocean sites", "Mixed lakes", "Stratified lakes")
Location_type_pres_sample <- oc_mix_strat_merge[,keep]

# Replace NA with 0
Location_type_pres_sample[is.na(Location_type_pres_sample)] <- 0

## Remove all species not found in any locations
# Identifies which rows are greater than 0
Location_type_pres_sample <- Location_type_pres_sample[which(rowSums(Location_type_pres_sample) > 0),]

Location_type_pres_sample_logical <- as.data.frame(Location_type_pres_sample > 0)

venn_plot <- ggvenn(Location_type_pres_sample_logical,
        show_percentage = F,
        fill_color = c("#EE6363", "#87CEFA", "#6E8B3D"),
        fill_alpha = 0.7,
        stroke_alpha = 0,
        stroke_size = 0.5, 
        set_name_size = 5,
        text_size = 5)
venn_plot
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/venn_plot_s10.jpg", plot = venn_plot, width = 4.25, height = 4, units = "in")
```

### SD alpha outliers for each Location type
```{r, SD alpha outliers}
outlier_SD_alpha <- SR_env[surveyed_sites,] %>%
  group_by(Location_type) %>%
  mutate(
  Q1 = quantile(row_sum, 0.25),
  Q3 = quantile(row_sum, 0.75),
  IQR = Q3 - Q1,
  lower_bound = Q1 - 1.5 * IQR,
  upper_bound = Q3 + 1.5 * IQR,
  is_outlier = row_sum < lower_bound | row_sum > upper_bound
  )

outlier_SD_alpha <- as.data.frame(outlier_SD_alpha)
row.names(outlier_SD_alpha) <- outlier_SD_alpha$X

# Create the plot
(outlier_SD_alpha_plot <- ggplot(outlier_SD_alpha, aes(x = Location_type, y = row_sum, fill = Location_type)) +
  geom_violin(alpha = 0.9, draw_quantiles = c(0.25, 0.5, 0.75), aes(fill = Location_type)) +
  geom_jitter(aes(color = is_outlier), width = 0.1, size = 2, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = outlier_SD_alpha, label = outlier_SD_alpha$X, size = 3, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme(text = element_text(size = 12),
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(color = "black", size = 12),
    axis.line = element_line(color = "black"),
    axis.title.y = element_text(margin = margin(t = 0)),  # reduce top margin (space from text)
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) +
  scale_y_continuous(expand = c(0,2)) +
  guides(color = "none", fill = "none") + 
  labs(y = "SRic", x = "Location type", color = "Outlier:", tag = "a"))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/outlier_SD_alpha.jpg", outlier_SD_alpha_plot, width = 3.25, height = 3.32, units = "in")
```

### Identify VIF of environmental and geographical variables
```{r, env and geo VIFs}
library(MASS)

# Determine which env variables have variance
environment <- c("temperature_median", "salinity_median", "oxygen_median", "pH_median")
nzv <- nearZeroVar(env[,environment])
nzv
# All env variables have variance

mod <-  lm(row_sum ~ temperature_median + salinity_median + oxygen_median + pH_median, SR_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)

mod <-  lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)
environment <- c("temperature_median", "salinity_median", "oxygen_median")


mod <- aov(temperature_median ~ Location_type, SR_env)
summary(mod)
mod <- aov(row_sum ~ oxygen_median + Location_type, SR_env)
car::vif(mod)

mod <- aov(salinity_median ~ Location_type, SR_env)
summary(mod)
mod <- aov(row_sum ~ salinity_median + Location_type, SR_env)
car::vif(mod)

mod <- aov(oxygen_median ~ Location_type, SR_env)
summary(mod)
mod <- aov(row_sum ~ oxygen_median + Location_type, SR_env)
car::vif(mod)

SR_ss_env <- SR_env[surveyed_sites_env,]
SR_ss_env$Location_type <- factor(SR_ss_env$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Run linear discriminant analysis of environmental variables
LDA <- lda(SR_ss_env[,environment], SR_ss_env$Location_type)
spe.class <- predict(LDA)$class
spe.post <- predict(LDA)$posterior
(spe.table <- table(SR_ss_env$Location_type, spe.class))
diag(prop.table(spe.table, 1))


# Determine which geo variables have variance
geography <- c("volume_m3_w_chemocline", "volume_m3", "surface_area_m2", "distance_to_ocean_min_m", "distance_to_ocean_mean_m", "distance_to_ocean_median_m", "tidal_lag_time_minutes", "tidal_efficiency", "perimeter_fromSat", "max_depth", "logArea")
nzv <- nearZeroVar(env[,geography])
nzv
# All geo variables have variance

mod <-  lm(row_sum ~ volume_m3_w_chemocline + volume_m3 + surface_area_m2 + distance_to_ocean_min_m + distance_to_ocean_mean_m + distance_to_ocean_median_m + tidal_lag_time_minutes + tidal_efficiency + perimeter_fromSat + max_depth + logArea, SR_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)

mod <-  lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)
geography <- c("distance_to_ocean_min_m", "max_depth", "logArea")

mod <- aov(distance_to_ocean_min_m ~ Location_type, SR_env)
summary(mod)
mod <- aov(row_sum ~ distance_to_ocean_min_m + Location_type, SR_env)
car::vif(mod)

mod <- aov(max_depth ~ Location_type, SR_env)
summary(mod)
mod <- aov(row_sum ~ max_depth + Location_type, SR_env)
car::vif(mod)

mod <- aov(logArea ~ Location_type, SR_env)
summary(mod)
mod <- aov(row_sum ~ logArea + Location_type, SR_env)
car::vif(mod)

SR_ss_geo <- SR_env[surveyed_sites,]
SR_ss_geo$Location_type <- factor(SR_ss_geo$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Run linear discriminant analysis of geographical variables
LDA <- lda(SR_ss_geo[,geography], SR_ss_geo$Location_type)
spe.class <- predict(LDA)$class
spe.post <- predict(LDA)$posterior
(spe.table <- table(SR_ss_geo$Location_type, spe.class))
diag(prop.table(spe.table, 1))


# Combine environment and geography
mod <-  lm(row_sum ~ salinity_median + oxygen_median + temperature_median + distance_to_ocean_min_m + max_depth + logArea, SR_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)

mod <-  lm(row_sum ~ oxygen_median + temperature_median + distance_to_ocean_min_m + logArea, SR_env)
summary(mod)
Anova(mod, type = 3)
rms::vif(mod)
car::vif(mod)
envgeo <- c("oxygen_median", "temperature_median", "distance_to_ocean_min_m", "logArea")


# Run linear discriminant analysis of geographical variables
LDA <- lda(SR_ss_env[,envgeo], SR_ss_env$Location_type)
spe.class <- predict(LDA)$class
spe.post <- predict(LDA)$posterior
(spe.table <- table(SR_ss_env$Location_type, spe.class))
diag(prop.table(spe.table, 1))
```

### SD alpha & Location type
```{r, SD alpha & Location type}
### Run negative binomial regression
# All location types
SD_logSRic <- MASS::glm.nb(row_sum ~ Location_type, data = SR_env[surveyed_sites,])
summary(SD_logSRic)$deviance / summary(SD_logSRic)$df.residual
# Check for outliers
outlierTest(SD_logSRic)
# Plot residuals
plot(SD_logSRic)
# Summarize the Interaction results
summary(SD_logSRic)
# p-values
logSRic_drop1_results <- drop1(SD_logSRic, test = "Chisq")
logSRic_drop1_results
SD_logSRic_p <- emmeans(SD_logSRic, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_p

# Ocean vs. Mixed
SD_logSRic_om <- MASS::glm.nb(row_sum ~ Location_type, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_om)$deviance / summary(SD_logSRic_om)$df.residual
# Check for outliers
outlierTest(SD_logSRic_om)
# Plot residuals
plot(SD_logSRic_om)
# Summarize the Interaction results
summary(SD_logSRic_om)
# p-values
logSRic_om_drop1_results <- drop1(SD_logSRic_om, test = "Chisq")
logSRic_om_drop1_results

# Ocean vs. Stratified
SD_logSRic_os <- MASS::glm.nb(row_sum ~ Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_os)$deviance / summary(SD_logSRic_os)$df.residual
# Check for outliers
outlierTest(SD_logSRic_os)
# Plot residuals
plot(SD_logSRic_os)
# Summarize the Interaction results
summary(SD_logSRic_os)
# p-values
logSRic_os_drop1_results <- drop1(SD_logSRic_os, test = "Chisq")
logSRic_os_drop1_results

# Mixed vs. Stratified
SD_logSRic_ms <- MASS::glm.nb(row_sum ~ Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_ms)$deviance / summary(SD_logSRic_ms)$df.residual
# Check for outliers
outlierTest(SD_logSRic_ms)
# Plot residuals
plot(SD_logSRic_ms)
# Summarize the Interaction results
summary(SD_logSRic_ms)
# p-values
logSRic_ms_drop1_results <- drop1(SD_logSRic_ms, test = "Chisq")
logSRic_ms_drop1_results

### Run ANOVA
anova_logSRic_result <- aov(log(row_sum) ~ Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(anova_logSRic_result))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(anova_logSRic_result) ~ SR_env[surveyed_sites,"Location_type"])
plot(anova_logSRic_result)

summary(anova_logSRic_result)

# Perform Tukey's HSD test
tukey_logSRic_result <- TukeyHSD(anova_logSRic_result)
print(tukey_logSRic_result)

# Sites
N <- length(anova_logSRic_result$residuals)
# Categories
k <- length(anova_logSRic_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_logSRic_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Location_type <- tukey_logSRic_result$Location_type

# q-values
(OM_q <- (abs(Location_type[1,1]))/USE)
(MS_q <- (abs(Location_type[3,1]))/BSE)
(SO_q <- (abs(Location_type[2,1]))/USE)


# Run ANOVA
anova_SRic_result <- aov(row_sum ~ Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(anova_SRic_result))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(anova_SRic_result) ~ SR_env[surveyed_sites,"Location_type"])
plot(anova_logSRic_result)

summary(anova_SRic_result)

# Perform Tukey's HSD test
tukey_SRic_result <- TukeyHSD(anova_SRic_result)
print(tukey_SRic_result)

# Sites
N <- length(anova_SRic_result$residuals)
# Categories
k <- length(anova_SRic_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_SRic_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Location_type <- tukey_SRic_result$Location_type

# q-values
(OM_q <- (abs(Location_type[1,1]))/USE)
(MS_q <- (abs(Location_type[3,1]))/BSE)
(SO_q <- (abs(Location_type[2,1]))/USE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx
```

### SD alpha env and geo variance partitioning
```{r}
envgeo_var <- env[surveyed_sites_env,c(environment, geography)]
surveyed_sites_lake_env <- surveyed_sites_lake[surveyed_sites_env,]
cca_var_all <- cca(surveyed_sites_lake_env ~ ., envgeo_var)
anova(cca_var_all)  # 0.001 *** - it is significant
adjR2_cca <- RsquareAdj(cca_var_all)$adj.r.squared # 0.10 - adjusted R2 explained by the 6 variables is 10%
cca_var_0 <- cca(surveyed_sites_lake_env ~ 1, data = envgeo_var)

# Run forward selection
SD_alpha_os_R2 <- ordiR2step(cca_var_0, scope = formula (cca_var_all), R2scope = adjR2_cca, direction = 'forward', permutations = 999)
SD_alpha_os_R2
SD_alpha_os_R2$anova
SD_alpha_os_R2_adj <- SD_alpha_os_R2
SD_alpha_os_R2_adj$anova$`Pr(>F)` <- p.adjust(SD_alpha_os_R2$anova$`Pr(>F)`, method = 'holm', n = ncol(envgeo_var))
SD_alpha_os_R2_adj$anova
```

### SD alpha env and geo plots
```{r, SD alpha lm plots}
# Temperature
SD_alpha_T_plot <- ggplot(data = SR_env[surveyed_sites_env,], mapping = aes(y = log(row_sum), x = temperature_median, color = Location_type, fill = Location_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = SR_env[surveyed_sites_env,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Temperature (ºC)", y="Log-SRic", colour = "Location type:", fill = "Location type:", tag = "a")
(SD_alpha_T_plot <- SD_alpha_T_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_alpha_T_plot.jpg", plot = SD_alpha_T_plot, width = 6.26, height = 6, units = "in")

# Salinity
SD_alpha_S_plot <- ggplot(data = SR_env[surveyed_sites_env,], mapping = aes(y = log(row_sum), x = salinity_median, color = Location_type, fill = Location_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = SR_env[surveyed_sites_env,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Salinity (ppt)", y="Log-SRic", colour = "Location type:", fill = "Location type:", tag = "b")
(SD_alpha_S_plot <- SD_alpha_S_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_alpha_S_plot.jpg", plot = SD_alpha_S_plot, width = 6.26, height = 6, units = "in")

# Oxygen
SD_alpha_O_plot <- ggplot(data = SR_env[surveyed_sites_env,], mapping = aes(y = log(row_sum), x = oxygen_median, color = Location_type, fill = Location_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = SR_env[surveyed_sites_env,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Oxygen (mg/L)", y="Log-SRic", colour = "Location type:", fill = "Location type:", tag = "c")
(SD_alpha_O_plot <- SD_alpha_O_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_alpha_O_plot.jpg", plot = SD_alpha_O_plot, width = 6.26, height = 6, units = "in")

# Distance from the ocean mean
SD_alpha_D_plot <- ggplot(data = SR_env[surveyed_sites,], mapping = aes(y = log(row_sum), x = distance_to_ocean_min_m, color = Location_type, fill = Location_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = SR_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Isolation (m)", y="Log-SRic", colour = "Location type:", fill = "Location type:", tag = "a")
(SD_alpha_D_plot <- SD_alpha_D_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_alpha_D_plot.jpg", plot = SD_alpha_D_plot, width = 6.26, height = 6, units = "in")

# Max depth
SD_alpha_MD_plot <- ggplot(data = SR_env[surveyed_sites,], mapping = aes(y = log(row_sum), x = max_depth, color = Location_type, fill = Location_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = SR_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Age (m)", y="Log-SRic", colour = "Location type:", fill = "Location type:", tag = "b")
(SD_alpha_MD_plot <- SD_alpha_MD_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_alpha_MD_plot.jpg", plot = SD_alpha_MD_plot, width = 6.26, height = 6, units = "in")

# Log Area
SD_alpha_LA_plot <- ggplot(data = SR_env[surveyed_sites,], mapping = aes(y = log(row_sum), x = logArea, color = Location_type, fill = Location_type)) + 
  geom_point(stat = 'identity',
  size = 4,
  alpha = 1) + 
  geom_smooth(method = "lm", se = FALSE) +
  geom_text_repel(label = SR_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
  axis.text = element_text(size = 16, color = "black"),
  axis.line = element_line(color = "black"),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank()) + 
  labs(x="Log Area (m"^"2"~")", y="Log-SRic", colour = "Location type:", fill = "Location type:", tag = "c")
(SD_alpha_LA_plot <- SD_alpha_LA_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_alpha_LA_plot.jpg", plot = SD_alpha_LA_plot, width = 6.26, height = 6, units = "in")

# 1. Create a linear model object
ocean_model <- lm(log(row_sum) ~ logArea, data = SR_env[ocean_sites,])
# 2. Extract the coefficients
ocean_coefficients <- coef(ocean_model)
# 3. Get the slope value (the second element)
ocean_slope <- ocean_coefficients[2]
# Print the slope
print(ocean_slope)
# 1. Create a linear model object
mixed_model <- lm(log(row_sum) ~ logArea, data = SR_env[mixed_lakes,])
# 2. Extract the coefficients
mixed_coefficients <- coef(mixed_model)
# 3. Get the slope value (the second element)
mixed_slope <- mixed_coefficients[2]
# Print the slope
print(mixed_slope)
# 1. Create a linear model object
stratified_model <- lm(log(row_sum) ~ logArea, data = SR_env[stratified_lakes,])
# 2. Extract the coefficients
stratified_coefficients <- coef(stratified_model)
# 3. Get the slope value (the second element)
stratified_slope <- stratified_coefficients[2]
# Print the slope
print(stratified_slope)
```

### SD alpha with env and geo linear models & ANCOVAs
```{r, SD alpha lm & ANCOVA}
par(mfrow=c(2,2)) 
####### logSRic GLMs
###### Environmental
##### Surveyed sites
#### Temperature
# Interaction
SD_logSRic_am_temp <- MASS::glm.nb(row_sum ~ temperature_median * Location_type, data = SR_env[surveyed_sites_env,])
summary(SD_logSRic_am_temp)$deviance / summary(SD_logSRic_am_temp)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_temp)
# Plot residuals
plot(SD_logSRic_am_temp)
# Summarize the Interaction results
summary(SD_logSRic_am_temp)
# p-values
temp_drop1_results <- drop1(SD_logSRic_am_temp, test = "Chisq")
temp_drop1_results
# Additive model
SD_logSRic_am_temp <- MASS::glm.nb(row_sum ~ temperature_median + Location_type, data = SR_env[surveyed_sites_env,])
summary(SD_logSRic_am_temp)$deviance / summary(SD_logSRic_am_temp)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_temp)
# Plot residuals
plot(SD_logSRic_am_temp)

### Ocean & Mixed sites
# Interaction
SD_logSRic_OM_am_temp <- MASS::glm.nb(row_sum ~ temperature_median * Location_type, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_OM_am_temp)$deviance / summary(SD_logSRic_OM_am_temp)$df.residual
# Plot residuals
plot(SD_logSRic_OM_am_temp)
# p-values
temp_drop1_results <- drop1(SD_logSRic_OM_am_temp, test = "Chisq")
temp_drop1_results
# Additive model
SD_logSRic_OM_am_temp <- MASS::glm.nb(row_sum ~ temperature_median + Location_type, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_OM_am_temp)$deviance / summary(SD_logSRic_OM_am_temp)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OM_am_temp)
# Plot residuals
plot(SD_logSRic_OM_am_temp)

### Ocean & Stratified sites
# Interaction
SD_logSRic_OS_am_temp <- MASS::glm.nb(row_sum ~ temperature_median * Location_type, data = SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_OS_am_temp)$deviance / summary(SD_logSRic_OS_am_temp)$df.residual
# Plot residuals
plot(SD_logSRic_OS_am_temp)
# p-values
temp_drop1_results <- drop1(SD_logSRic_OS_am_temp, test = "Chisq")
temp_drop1_results
# Additive model
SD_logSRic_OS_am_temp <- MASS::glm.nb(row_sum ~ temperature_median + Location_type, data = SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_OS_am_temp)$deviance / summary(SD_logSRic_OS_am_temp)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OS_am_temp)
# Plot residuals
plot(SD_logSRic_OS_am_temp)

### Mixed & Stratified lakes
SD_logSRic_MS_am_temp <- MASS::glm.nb(row_sum ~ temperature_median * Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_temp)$deviance / summary(SD_logSRic_MS_am_temp)$df.residual
# Plot residuals
plot(SD_logSRic_MS_am_temp)
# p-values
temp_drop1_results <- drop1(SD_logSRic_MS_am_temp, test = "Chisq")
temp_drop1_results
# Additive model
SD_logSRic_MS_am_temp <- MASS::glm.nb(row_sum ~ temperature_median + Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_temp)$deviance / summary(SD_logSRic_MS_am_temp)$df.residual
# Check for outliers
outlierTest(SD_logSRic_MS_am_temp)
# Plot residuals
plot(SD_logSRic_MS_am_temp)

### Mixed lakes
SD_logSRic_M_am_temp <- MASS::glm.nb(row_sum ~ temperature_median, data = SR_env[mixed_lakes,])

### Stratified lakes
SD_logSRic_S_am_temp <- MASS::glm.nb(row_sum ~ temperature_median, data = SR_env[stratified_lakes,])


#### Salinity
# Interaction
SD_logSRic_am_sal <- MASS::glm.nb(row_sum ~ salinity_median * Location_type, data = SR_env[surveyed_sites_env,])
summary(SD_logSRic_am_sal)$deviance / summary(SD_logSRic_am_sal)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_sal)
# Plot residuals
plot(SD_logSRic_am_sal)
# p-values
sal_drop1_results <- drop1(SD_logSRic_am_sal, test = "Chisq")
sal_drop1_results
# Additive model
SD_logSRic_am_sal <- MASS::glm.nb(row_sum ~ salinity_median + Location_type, data = SR_env[surveyed_sites_env,])
summary(SD_logSRic_am_sal)$deviance / summary(SD_logSRic_am_sal)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_sal)
# Plot residuals
plot(SD_logSRic_am_sal)

### Ocean & Mixed sites
# Interaction
SD_logSRic_OM_am_sal <- MASS::glm.nb(row_sum ~ salinity_median * Location_type, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_OM_am_sal)$deviance / summary(SD_logSRic_OM_am_sal)$df.residual
# Plot residuals
plot(SD_logSRic_OM_am_sal)
# p-values
sal_drop1_results <- drop1(SD_logSRic_OM_am_sal, test = "Chisq")
sal_drop1_results
# Additive model
SD_logSRic_OM_am_sal <- MASS::glm.nb(row_sum ~ salinity_median + Location_type, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_OM_am_sal)$deviance / summary(SD_logSRic_OM_am_sal)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OM_am_sal)
# Plot residuals
plot(SD_logSRic_OM_am_sal)

### Ocean & Stratified sites
# Interaction
SD_logSRic_OS_am_sal <- MASS::glm.nb(row_sum ~ salinity_median * Location_type, data = SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_OS_am_sal)$deviance / summary(SD_logSRic_OS_am_sal)$df.residual
# Plot residuals
plot(SD_logSRic_OS_am_sal)
# p-values
sal_drop1_results <- drop1(SD_logSRic_OS_am_sal, test = "Chisq")
sal_drop1_results
# Additive model
SD_logSRic_OS_am_sal <- MASS::glm.nb(row_sum ~ salinity_median + Location_type, data = SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_OS_am_sal)$deviance / summary(SD_logSRic_OS_am_sal)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OS_am_sal)
# Plot residuals
plot(SD_logSRic_OS_am_sal)

### Mixed & Stratified lakes
SD_logSRic_MS_am_sal <- MASS::glm.nb(row_sum ~ salinity_median * Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_sal)$deviance / summary(SD_logSRic_MS_am_sal)$df.residual
# Plot residuals
plot(SD_logSRic_MS_am_sal)
# p-values
sal_drop1_results <- drop1(SD_logSRic_MS_am_sal, test = "Chisq")
sal_drop1_results
# Additive model
SD_logSRic_MS_am_sal <- MASS::glm.nb(row_sum ~ salinity_median + Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_sal)$deviance / summary(SD_logSRic_MS_am_sal)$df.residual
# Check for outliers
outlierTest(SD_logSRic_MS_am_sal)
# Plot residuals
plot(SD_logSRic_MS_am_sal)

### Mixed lakes
SD_logSRic_M_am_sal <- MASS::glm.nb(row_sum ~ salinity_median, data = SR_env[mixed_lakes,])

### Stratified lakes
SD_logSRic_S_am_sal <- MASS::glm.nb(row_sum ~ salinity_median, data = SR_env[stratified_lakes,])

## Oxygen
# Interaction
SD_logSRic_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median * Location_type, data = SR_env[surveyed_sites_env,])
summary(SD_logSRic_am_oxy)$deviance / summary(SD_logSRic_am_oxy)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_oxy)
# Plot residuals
plot(SD_logSRic_am_oxy)
# p-values
oxy_drop1_results <- drop1(SD_logSRic_am_oxy, test = "Chisq")
oxy_drop1_results
# Get the p-value for Temperature specifically for EACH group
slopes <- emtrends(SD_logSRic_am_oxy, ~ Location_type, var = "oxygen_median")
# view the p-values
test(slopes)

### Ocean & Mixed sites
# Interaction
SD_logSRic_OM_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median * Location_type, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_OM_am_oxy)$deviance / summary(SD_logSRic_OM_am_oxy)$df.residual
# Plot residuals
plot(SD_logSRic_OM_am_oxy)
# p-values
oxy_drop1_results <- drop1(SD_logSRic_OM_am_oxy, test = "Chisq")
oxy_drop1_results
# Additive model
SD_logSRic_OM_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median + Location_type, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_OM_am_oxy)$deviance / summary(SD_logSRic_OM_am_oxy)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OM_am_oxy)
# Plot residuals
plot(SD_logSRic_OM_am_oxy)

### Ocean & Stratified sites
# Interaction
SD_logSRic_OS_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median * Location_type, data = SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_OS_am_oxy)$deviance / summary(SD_logSRic_OS_am_oxy)$df.residual
# Plot residuals
plot(SD_logSRic_OS_am_oxy)
# p-values
oxy_drop1_results <- drop1(SD_logSRic_OS_am_oxy, test = "Chisq")
oxy_drop1_results
# Get the p-value for Temperature specifically for EACH group
slopes <- emtrends(SD_logSRic_OS_am_oxy, ~ Location_type, var = "oxygen_median")
# view the p-values
test(slopes)

### Mixed & Stratified lakes
SD_logSRic_MS_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median * Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_oxy)$deviance / summary(SD_logSRic_MS_am_oxy)$df.residual
# Plot residuals
plot(SD_logSRic_MS_am_oxy)
# p-values
oxy_drop1_results <- drop1(SD_logSRic_MS_am_oxy, test = "Chisq")
oxy_drop1_results
# Get the p-value for Temperature specifically for EACH group
slopes <- emtrends(SD_logSRic_MS_am_oxy, ~ Location_type, var = "oxygen_median")
# view the p-values
test(slopes)

### Mixed lakes
SD_logSRic_M_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median, data = SR_env[mixed_lakes,])

### Stratified lakes
SD_logSRic_S_am_oxy <- MASS::glm.nb(row_sum ~ oxygen_median, data = SR_env[stratified_lakes,])

# Summarize GLM results and calculate pairwise comparisons
summary(SD_logSRic_am_temp)
SD_logSRic_amp_temp <- emmeans(SD_logSRic_am_temp, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_amp_temp$contrasts
summary(SD_logSRic_am_sal)
SD_logSRic_amp_sal <- emmeans(SD_logSRic_am_sal, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_amp_sal$contrasts
summary(SD_logSRic_am_oxy)
SD_logSRic_amp_oxy <- emmeans(SD_logSRic_am_oxy, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_amp_oxy$contrasts
# p-values
temp_drop1_results <- drop1(SD_logSRic_am_temp, test = "Chisq")
temp_drop1_results
temp_p_values <- temp_drop1_results[["Pr(>Chi)"]]
sal_drop1_results <- drop1(SD_logSRic_am_sal, test = "Chisq")
sal_drop1_results
sal_p_values <- sal_drop1_results[["Pr(>Chi)"]]
oxy_drop1_results <- drop1(SD_logSRic_am_oxy, test = "Chisq")
oxy_drop1_results
oxy_p_values <- oxy_drop1_results[["Pr(>Chi)"]]
p_values <- c(temp_p_values[2:3], sal_p_values[2:3], oxy_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OM lm results
temp_drop1_OM_results <- drop1(SD_logSRic_OM_am_temp, test = "Chisq")
temp_drop1_OM_results
summary(SD_logSRic_OM_am_temp)
sal_drop1_OM_results <- drop1(SD_logSRic_OM_am_sal, test = "Chisq")
sal_drop1_OM_results
summary(SD_logSRic_OM_am_sal)
oxy_drop1_OM_results <- drop1(SD_logSRic_OM_am_oxy, test = "Chisq")
oxy_drop1_OM_results
summary(SD_logSRic_OM_am_oxy)
# p-values
temp_p_values <- temp_drop1_OM_results[["Pr(>Chi)"]]
sal_p_values <- sal_drop1_OM_results[["Pr(>Chi)"]]
oxy_p_values <- oxy_drop1_OM_results[["Pr(>Chi)"]]
p_values <- c(temp_p_values[2:3], sal_p_values[2:3], oxy_p_values[2:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OS lm results
temp_drop1_OS_results <- drop1(SD_logSRic_OS_am_temp, test = "Chisq")
temp_drop1_OS_results
summary(SD_logSRic_OS_am_temp)
sal_drop1_OS_results <- drop1(SD_logSRic_OS_am_sal, test = "Chisq")
sal_drop1_OS_results
summary(SD_logSRic_OS_am_sal)
oxy_drop1_OS_results <- drop1(SD_logSRic_OS_am_oxy, test = "Chisq")
oxy_drop1_OS_results
summary(SD_logSRic_OS_am_oxy)
# p-values
temp_p_values <- temp_drop1_OS_results[["Pr(>Chi)"]]
sal_p_values <- sal_drop1_OS_results[["Pr(>Chi)"]]
oxy_p_values <- oxy_drop1_OS_results[["Pr(>Chi)"]]
p_values <- c(temp_p_values[2:3], sal_p_values[2:3], oxy_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize MS lm results
temp_drop1_MS_results <- drop1(SD_logSRic_MS_am_temp, test = "Chisq")
temp_drop1_MS_results
summary(SD_logSRic_MS_am_temp)
sal_drop1_MS_results <- drop1(SD_logSRic_MS_am_sal, test = "Chisq")
sal_drop1_MS_results
summary(SD_logSRic_MS_am_sal)
oxy_drop1_MS_results <- drop1(SD_logSRic_MS_am_oxy, test = "Chisq")
oxy_drop1_MS_results
summary(SD_logSRic_MS_am_oxy)
# p-values
temp_p_values <- temp_drop1_MS_results[["Pr(>Chi)"]]
sal_p_values <- sal_drop1_MS_results[["Pr(>Chi)"]]
oxy_p_values <- oxy_drop1_MS_results[["Pr(>Chi)"]]
p_values <- c(temp_p_values[2:3], sal_p_values[2:3], oxy_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize M lm results
temp_drop1_M_results <- drop1(SD_logSRic_M_am_temp, test = "Chisq")
temp_drop1_M_results
summary(SD_logSRic_M_am_temp)
sal_drop1_M_results <- drop1(SD_logSRic_M_am_sal, test = "Chisq")
sal_drop1_M_results
summary(SD_logSRic_M_am_sal)
oxy_drop1_M_results <- drop1(SD_logSRic_M_am_oxy, test = "Chisq")
oxy_drop1_M_results
summary(SD_logSRic_M_am_oxy)
# p-values
temp_p_values <- temp_drop1_M_results[["Pr(>Chi)"]]
sal_p_values <- sal_drop1_M_results[["Pr(>Chi)"]]
oxy_p_values <- oxy_drop1_M_results[["Pr(>Chi)"]]
p_values <- c(temp_p_values[2], sal_p_values[2], oxy_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize S lm results
temp_drop1_S_results <- drop1(SD_logSRic_S_am_temp, test = "Chisq")
temp_drop1_S_results
summary(SD_logSRic_S_am_temp)
sal_drop1_S_results <- drop1(SD_logSRic_S_am_sal, test = "Chisq")
sal_drop1_S_results
summary(SD_logSRic_S_am_sal)
oxy_drop1_S_results <- drop1(SD_logSRic_S_am_oxy, test = "Chisq")
oxy_drop1_S_results
summary(SD_logSRic_S_am_oxy)
# p-values
temp_p_values <- temp_drop1_S_results[["Pr(>Chi)"]]
sal_p_values <- sal_drop1_S_results[["Pr(>Chi)"]]
oxy_p_values <- oxy_drop1_S_results[["Pr(>Chi)"]]
p_values <- c(temp_p_values[2], sal_p_values[2], oxy_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


###### Geographical
##### Surveyed sites
#### Distance
# Interaction
SD_logSRic_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m * Location_type, data = SR_env[surveyed_sites,])
summary(SD_logSRic_am_dist)$deviance / summary(SD_logSRic_am_dist)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_dist)
# Plot residuals
plot(SD_logSRic_am_dist)
# Summarize the Interaction results
summary(SD_logSRic_am_dist)
# p-values
dist_drop1_results <- drop1(SD_logSRic_am_dist, test = "Chisq")
dist_drop1_results
# Additive model
SD_logSRic_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m + Location_type, data = SR_env[surveyed_sites,])
summary(SD_logSRic_am_dist)$deviance / summary(SD_logSRic_am_dist)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_dist)
# Plot residuals
plot(SD_logSRic_am_dist)

### Ocean & Mixed sites
# Interaction
SD_logSRic_OM_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m * Location_type, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_OM_am_dist)$deviance / summary(SD_logSRic_OM_am_dist)$df.residual
# Plot residuals
plot(SD_logSRic_OM_am_dist)
# p-values
dist_drop1_results <- drop1(SD_logSRic_OM_am_dist, test = "Chisq")
dist_drop1_results
# Additive model
SD_logSRic_OM_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m + Location_type, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_OM_am_dist)$deviance / summary(SD_logSRic_OM_am_dist)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OM_am_dist)
# Plot residuals
plot(SD_logSRic_OM_am_dist)

### Ocean & Stratified sites
# Interaction
SD_logSRic_OS_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m * Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_OS_am_dist)$deviance / summary(SD_logSRic_OS_am_dist)$df.residual
# Plot residuals
plot(SD_logSRic_OS_am_dist)
# p-values
dist_drop1_results <- drop1(SD_logSRic_OS_am_dist, test = "Chisq")
dist_drop1_results
# Additive model
SD_logSRic_OS_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m + Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_OS_am_dist)$deviance / summary(SD_logSRic_OS_am_dist)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OS_am_dist)
# Plot residuals
plot(SD_logSRic_OS_am_dist)

### Mixed & Stratified lakes
SD_logSRic_MS_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m * Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_dist)$deviance / summary(SD_logSRic_MS_am_dist)$df.residual
# Plot residuals
plot(SD_logSRic_MS_am_dist)
# p-values
dist_drop1_results <- drop1(SD_logSRic_MS_am_dist, test = "Chisq")
dist_drop1_results
# Additive model
SD_logSRic_MS_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m + Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_dist)$deviance / summary(SD_logSRic_MS_am_dist)$df.residual
# Check for outliers
outlierTest(SD_logSRic_MS_am_dist)
# Plot residuals
plot(SD_logSRic_MS_am_dist)

### Mixed lakes
SD_logSRic_M_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m, data = SR_env[mixed_lakes,])

### Stratified lakes
SD_logSRic_S_am_dist <- MASS::glm.nb(row_sum ~ distance_to_ocean_min_m, data = SR_env[stratified_lakes,])


#### Max depth
# Interaction
SD_logSRic_am_mxd <- MASS::glm.nb(row_sum ~ max_depth * Location_type, data = SR_env[surveyed_sites,])
summary(SD_logSRic_am_mxd)$deviance / summary(SD_logSRic_am_mxd)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_mxd)
# Plot residuals
plot(SD_logSRic_am_mxd)
# p-values
mxd_drop1_results <- drop1(SD_logSRic_am_mxd, test = "Chisq")
mxd_drop1_results
# Additive model
SD_logSRic_am_mxd <- MASS::glm.nb(row_sum ~ max_depth + Location_type, data = SR_env[surveyed_sites,])
summary(SD_logSRic_am_mxd)$deviance / summary(SD_logSRic_am_mxd)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_mxd)
# Plot residuals
plot(SD_logSRic_am_mxd)

### Ocean & Mixed sites
# Interaction
SD_logSRic_OM_am_mxd <- MASS::glm.nb(row_sum ~ max_depth * Location_type, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_OM_am_mxd)$deviance / summary(SD_logSRic_OM_am_mxd)$df.residual
# Plot residuals
plot(SD_logSRic_OM_am_mxd)
# p-values
mxd_drop1_results <- drop1(SD_logSRic_OM_am_mxd, test = "Chisq")
mxd_drop1_results
# Additive model
SD_logSRic_OM_am_mxd <- MASS::glm.nb(row_sum ~ max_depth + Location_type, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_OM_am_mxd)$deviance / summary(SD_logSRic_OM_am_mxd)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OM_am_mxd)
# Plot residuals
plot(SD_logSRic_OM_am_mxd)

### Ocean & Stratified sites
# Interaction
SD_logSRic_OS_am_mxd <- MASS::glm.nb(row_sum ~ max_depth * Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_OS_am_mxd)$deviance / summary(SD_logSRic_OS_am_mxd)$df.residual
# Plot residuals
plot(SD_logSRic_OS_am_mxd)
# p-values
mxd_drop1_results <- drop1(SD_logSRic_OS_am_mxd, test = "Chisq")
mxd_drop1_results
# Additive model
SD_logSRic_OS_am_mxd <- MASS::glm.nb(row_sum ~ max_depth + Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_OS_am_mxd)$deviance / summary(SD_logSRic_OS_am_mxd)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OS_am_mxd)
# Plot residuals
plot(SD_logSRic_OS_am_mxd)

### Mixed & Stratified lakes
SD_logSRic_MS_am_mxd <- MASS::glm.nb(row_sum ~ max_depth * Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_mxd)$deviance / summary(SD_logSRic_MS_am_mxd)$df.residual
# Plot residuals
plot(SD_logSRic_MS_am_mxd)
# p-values
mxd_drop1_results <- drop1(SD_logSRic_MS_am_mxd, test = "Chisq")
mxd_drop1_results
# Additive model
SD_logSRic_MS_am_mxd <- MASS::glm.nb(row_sum ~ max_depth + Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_mxd)$deviance / summary(SD_logSRic_MS_am_mxd)$df.residual
# Check for outliers
outlierTest(SD_logSRic_MS_am_mxd)
# Plot residuals
plot(SD_logSRic_MS_am_mxd)

### Mixed lakes
SD_logSRic_M_am_mxd <- MASS::glm.nb(row_sum ~ max_depth, data = SR_env[mixed_lakes,])

### Stratified lakes
SD_logSRic_S_am_mxd <- MASS::glm.nb(row_sum ~ max_depth, data = SR_env[stratified_lakes,])

## Log Area
# Interaction
SD_logSRic_am_lga <- MASS::glm.nb(row_sum ~ logArea * Location_type, data = SR_env[surveyed_sites,])
summary(SD_logSRic_am_lga)$deviance / summary(SD_logSRic_am_lga)$df.residual
# Check for outliers
outlierTest(SD_logSRic_am_lga)
# Plot residuals
plot(SD_logSRic_am_lga)
# p-values
lga_drop1_results <- drop1(SD_logSRic_am_lga, test = "Chisq")
lga_drop1_results
summary(SD_logSRic_am_lga)
# Get the p-value for Log Area specifically for EACH group
slopes <- emtrends(SD_logSRic_am_lga, ~ Location_type, var = "logArea")
# view the p-values
test(slopes)

### Ocean & Mixed sites
# Interaction
SD_logSRic_OM_am_lga <- MASS::glm.nb(row_sum ~ logArea * Location_type, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_OM_am_lga)$deviance / summary(SD_logSRic_OM_am_lga)$df.residual
# Plot residuals
plot(SD_logSRic_OM_am_lga)
# p-values
lga_drop1_results <- drop1(SD_logSRic_OM_am_lga, test = "Chisq")
lga_drop1_results
summary(SD_logSRic_OM_am_lga)
# Get the p-value for Log Area specifically for EACH group
slopes <- emtrends(SD_logSRic_OM_am_lga, ~ Location_type, var = "logArea")
# view the p-values
test(slopes)

### Ocean & Stratified sites
# Interaction
SD_logSRic_OS_am_lga <- MASS::glm.nb(row_sum ~ logArea * Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_OS_am_lga)$deviance / summary(SD_logSRic_OS_am_lga)$df.residual
# Plot residuals
plot(SD_logSRic_OS_am_lga)
# p-values
lga_drop1_results <- drop1(SD_logSRic_OS_am_lga, test = "Chisq")
lga_drop1_results
# Additive model
SD_logSRic_OS_am_lga <- MASS::glm.nb(row_sum ~ logArea + Location_type, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_OS_am_lga)$deviance / summary(SD_logSRic_OS_am_lga)$df.residual
# Check for outliers
outlierTest(SD_logSRic_OS_am_lga)
# Plot residuals
plot(SD_logSRic_OS_am_lga)

### Mixed & Stratified lakes
SD_logSRic_MS_am_lga <- MASS::glm.nb(row_sum ~ logArea * Location_type, data = SR_env[mixed_stratified_lakes,])
# Plot residuals
plot(SD_logSRic_MS_am_lga)
# p-values
lga_drop1_results <- drop1(SD_logSRic_MS_am_lga, test = "Chisq")
lga_drop1_results
# Additive model
SD_logSRic_MS_am_lga <- MASS::glm.nb(row_sum ~ logArea + Location_type, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_MS_am_lga)$deviance / summary(SD_logSRic_MS_am_lga)$df.residual
# Check for outliers
outlierTest(SD_logSRic_MS_am_lga)
# Plot residuals
plot(SD_logSRic_MS_am_lga)

### Mixed lakes
SD_logSRic_M_am_lga <- MASS::glm.nb(row_sum ~ logArea, data = SR_env[mixed_lakes,])

### Stratified lakes
SD_logSRic_S_am_lga <- MASS::glm.nb(row_sum ~ logArea, data = SR_env[stratified_lakes,])

# Summarize GLM results and calculate pairwise comparisons
summary(SD_logSRic_am_dist)
SD_logSRic_amp_dist <- emmeans(SD_logSRic_am_dist, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_amp_dist$contrasts
summary(SD_logSRic_am_mxd)
SD_logSRic_amp_mxd <- emmeans(SD_logSRic_am_mxd, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_amp_mxd$contrasts
summary(SD_logSRic_am_lga)
SD_logSRic_amp_lga <- emmeans(SD_logSRic_am_lga, pairwise ~ Location_type, adjust = "bonferroni")
SD_logSRic_amp_lga$contrasts
# p-values
dist_drop1_results <- drop1(SD_logSRic_am_dist, test = "Chisq")
dist_drop1_results
dist_p_values <- dist_drop1_results[["Pr(>Chi)"]]
mxd_drop1_results <- drop1(SD_logSRic_am_mxd, test = "Chisq")
mxd_drop1_results
mxd_p_values <- mxd_drop1_results[["Pr(>Chi)"]]
lga_drop1_results <- drop1(SD_logSRic_am_lga, test = "Chisq")
lga_drop1_results
lga_p_values <- lga_drop1_results[["Pr(>Chi)"]]
p_values <- c(dist_p_values[2:3], mxd_p_values[2:3], lga_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OM lm results
dist_drop1_OM_results <- drop1(SD_logSRic_OM_am_dist, test = "Chisq")
dist_drop1_OM_results
summary(SD_logSRic_OM_am_dist)
mxd_drop1_OM_results <- drop1(SD_logSRic_OM_am_mxd, test = "Chisq")
mxd_drop1_OM_results
summary(SD_logSRic_OM_am_mxd)
lga_drop1_OM_results <- drop1(SD_logSRic_OM_am_lga, test = "Chisq")
lga_drop1_OM_results
summary(SD_logSRic_OM_am_lga)
# p-values
dist_p_values <- dist_drop1_OM_results[["Pr(>Chi)"]]
mxd_p_values <- mxd_drop1_OM_results[["Pr(>Chi)"]]
lga_p_values <- lga_drop1_OM_results[["Pr(>Chi)"]]
p_values <- c(dist_p_values[2:3], mxd_p_values[2:3], lga_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize OS lm results
dist_drop1_OS_results <- drop1(SD_logSRic_OS_am_dist, test = "Chisq")
dist_drop1_OS_results
summary(SD_logSRic_OS_am_dist)
mxd_drop1_OS_results <- drop1(SD_logSRic_OS_am_mxd, test = "Chisq")
mxd_drop1_OS_results
summary(SD_logSRic_OS_am_mxd)
lga_drop1_OS_results <- drop1(SD_logSRic_OS_am_lga, test = "Chisq")
lga_drop1_OS_results
summary(SD_logSRic_OS_am_lga)
# p-values
dist_p_values <- dist_drop1_OS_results[["Pr(>Chi)"]]
mxd_p_values <- mxd_drop1_OS_results[["Pr(>Chi)"]]
lga_p_values <- lga_drop1_OS_results[["Pr(>Chi)"]]
p_values <- c(dist_p_values[2:3], mxd_p_values[2:3], lga_p_values[2:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize MS lm results
dist_drop1_MS_results <- drop1(SD_logSRic_MS_am_dist, test = "Chisq")
dist_drop1_MS_results
summary(SD_logSRic_MS_am_dist)
mxd_drop1_MS_results <- drop1(SD_logSRic_MS_am_mxd, test = "Chisq")
mxd_drop1_MS_results
summary(SD_logSRic_MS_am_mxd)
lga_drop1_MS_results <- drop1(SD_logSRic_MS_am_lga, test = "Chisq")
lga_drop1_MS_results
summary(SD_logSRic_MS_am_lga)
# p-values
dist_p_values <- dist_drop1_MS_results[["Pr(>Chi)"]]
mxd_p_values <- mxd_drop1_MS_results[["Pr(>Chi)"]]
lga_p_values <- lga_drop1_MS_results[["Pr(>Chi)"]]
p_values <- c(dist_p_values[2:3], mxd_p_values[2:3], lga_p_values[2:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize M lm results
dist_drop1_M_results <- drop1(SD_logSRic_M_am_dist, test = "Chisq")
dist_drop1_M_results
summary(SD_logSRic_M_am_dist)
mxd_drop1_M_results <- drop1(SD_logSRic_M_am_mxd, test = "Chisq")
mxd_drop1_M_results
summary(SD_logSRic_M_am_mxd)
lga_drop1_M_results <- drop1(SD_logSRic_M_am_lga, test = "Chisq")
lga_drop1_M_results
summary(SD_logSRic_M_am_lga)
# p-values
dist_p_values <- dist_drop1_M_results[["Pr(>Chi)"]]
mxd_p_values <- mxd_drop1_M_results[["Pr(>Chi)"]]
lga_p_values <- lga_drop1_M_results[["Pr(>Chi)"]]
p_values <- c(dist_p_values[2], mxd_p_values[2], lga_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Summarize S lm results
dist_drop1_S_results <- drop1(SD_logSRic_S_am_dist, test = "Chisq")
dist_drop1_S_results
summary(SD_logSRic_S_am_dist)
mxd_drop1_S_results <- drop1(SD_logSRic_S_am_mxd, test = "Chisq")
mxd_drop1_S_results
summary(SD_logSRic_S_am_mxd)
lga_drop1_S_results <- drop1(SD_logSRic_S_am_lga, test = "Chisq")
lga_drop1_S_results
summary(SD_logSRic_S_am_lga)
# p-values
dist_p_values <- dist_drop1_S_results[["Pr(>Chi)"]]
mxd_p_values <- mxd_drop1_S_results[["Pr(>Chi)"]]
lga_p_values <- lga_drop1_S_results[["Pr(>Chi)"]]
p_values <- c(dist_p_values[2], mxd_p_values[2], lga_p_values[2])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


##### SRic ANCOVAs
#### Environmental
### Surveyed sites
## Temperature
# Interaction
SD_SRic_am_temp <- aov(row_sum ~ temperature_median * Location_type, data = SR_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_temp) ~ SR_env[surveyed_sites_env,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_temp)
# Plot residuals
plot(SD_SRic_am_temp)
# Summarize the ANCOVA results
summary(SD_SRic_am_temp)
# ANCOVA
SD_SRic_am_temp <- aov(row_sum ~ temperature_median + Location_type, data = SR_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_temp))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_temp) ~ SR_env[surveyed_sites_env,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_temp)
# Plot residuals
plot(SD_SRic_am_temp)

## Salinity
# Interaction
SD_SRic_am_sal <- aov(row_sum ~ salinity_median * Location_type, data = SR_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_sal) ~ SR_env[surveyed_sites_env,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_sal)
# Plot residuals
plot(SD_SRic_am_sal)
# Summarize the ANCOVA results
summary(SD_SRic_am_sal)
# ANCOVA
SD_SRic_am_sal <- aov(row_sum ~ salinity_median + Location_type, data = SR_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_sal))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_sal) ~ SR_env[surveyed_sites_env,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_sal)
# Plot residuals
plot(SD_SRic_am_sal)

## Oxygen
# Interaction
SD_SRic_am_oxy <- aov(row_sum ~ oxygen_median * Location_type, data = SR_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_oxy) ~ SR_env[surveyed_sites_env,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_oxy)
# Plot residuals
plot(SD_SRic_am_oxy)
# Summarize the ANCOVA results
summary(SD_SRic_am_oxy)
# ANCOVA
SD_SRic_am_oxy <- aov(row_sum ~ oxygen_median + Location_type, data = SR_env[surveyed_sites_env,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_oxy))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_oxy) ~ SR_env[surveyed_sites_env,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_oxy)
# Plot residuals
plot(SD_SRic_am_oxy)

# Summarize ANCOVA results and calculate pairwise comparisons
summary(SD_SRic_am_temp)
SD_SRic_amp_temp <- emmeans(SD_SRic_am_temp, pairwise ~ Location_type, adjust = "bonferroni")
SD_SRic_amp_temp$contrasts
summary(SD_SRic_am_sal)
SD_SRic_amp_sal <- emmeans(SD_SRic_am_sal, pairwise ~ Location_type, adjust = "bonferroni")
SD_SRic_amp_sal$contrasts
summary(SD_SRic_am_oxy)
SD_SRic_amp_oxy <- emmeans(SD_SRic_am_oxy, pairwise ~ Location_type, adjust = "bonferroni")
SD_SRic_amp_oxy$contrasts
# p-values
temp_p_values <- summary(SD_SRic_am_temp)[[1]][, "Pr(>F)"]
sal_p_values <- summary(SD_SRic_am_sal)[[1]][, "Pr(>F)"]
oxy_p_values <- summary(SD_SRic_am_oxy)[[1]][, "Pr(>F)"]
p_values <- c(temp_p_values[1:3], sal_p_values[1:3], oxy_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Mixed lakes
SD_SRic_env_M_lm <- lm(row_sum ~ salinity_median + oxygen_median + temperature_median, SR_env[mixed_lakes,])
shapiro.test(residuals(SD_SRic_env_M_lm))
summary(SD_SRic_env_M_lm)
car::Anova(SD_SRic_env_M_lm, type = 3)
p_values <- summary(SD_SRic_env_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Stratified lakes
SD_SRic_env_S_lm <- lm(row_sum ~ salinity_median + oxygen_median + temperature_median, SR_env[stratified_lakes,])
shapiro.test(residuals(SD_SRic_env_S_lm))
summary(SD_SRic_env_S_lm)
car::Anova(SD_SRic_env_S_lm, type = 3)
p_values <- summary(SD_SRic_env_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

#### Geographical
### Surveyed sites
## Distance
# Interaction
SD_SRic_am_dist <- aov(row_sum ~ distance_to_ocean_min_m * Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_dist) ~ SR_env[surveyed_sites,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_dist)
# Plot residuals
plot(SD_SRic_am_dist)
# Summarize the ANCOVA results
summary(SD_SRic_am_dist)
# ANCOVA
SD_SRic_am_dist <- aov(row_sum ~ distance_to_ocean_min_m + Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_dist))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_dist) ~ SR_env[surveyed_sites,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_dist)
# Plot residuals
plot(SD_SRic_am_dist)

## Max Depth
# Interaction
SD_SRic_am_mxd <- aov(row_sum ~ max_depth * Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_mxd) ~ SR_env[surveyed_sites,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_mxd)
# Plot residuals
plot(SD_SRic_am_mxd)
# Summarize the ANCOVA results
summary(SD_SRic_am_mxd)
# ANCOVA
SD_SRic_am_mxd <- aov(row_sum ~ max_depth + Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_mxd))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_mxd) ~ SR_env[surveyed_sites,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_mxd)
# Plot residuals
plot(SD_SRic_am_mxd)

## Log Area
# Interaction
SD_SRic_am_lga <- aov(row_sum ~ logArea * Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_lga) ~ SR_env[surveyed_sites,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_lga)
# Plot residuals
plot(SD_SRic_am_lga)
# Summarize the ANCOVA results
summary(SD_SRic_am_lga)
# ANCOVA
SD_SRic_am_lga <- aov(row_sum ~ logArea + Location_type, data = SR_env[surveyed_sites,])
# Shapiro-Wilk test on residuals
shapiro.test(residuals(SD_SRic_am_lga))
# Levene’s test for homogeneity of variance
car::leveneTest(residuals(SD_SRic_am_lga) ~ SR_env[surveyed_sites,"Location_type"])
# Check for outliers
outlierTest(SD_SRic_am_lga)
# Plot residuals
plot(SD_SRic_am_lga)

# Summarize ANCOVA results and calculate pairwise comparisons
summary(SD_SRic_am_dist)
SD_SRic_amp_dist <- emmeans(SD_SRic_am_dist, pairwise ~ Location_type, adjust = "bonferroni")
SD_SRic_amp_dist$contrasts
summary(SD_SRic_am_mxd)
SD_SRic_amp_mxd <- emmeans(SD_SRic_am_mxd, pairwise ~ Location_type, adjust = "bonferroni")
SD_SRic_amp_mxd$contrasts
summary(SD_SRic_am_lga)
SD_SRic_amp_lga <- emmeans(SD_SRic_am_lga, pairwise ~ Location_type, adjust = "bonferroni")
SD_SRic_amp_lga$contrasts
# p-values
dist_p_values <- summary(SD_SRic_am_dist)[[1]][, "Pr(>F)"]
mxd_p_values <- summary(SD_SRic_am_mxd)[[1]][, "Pr(>F)"]
lga_p_values <- summary(SD_SRic_am_lga)[[1]][, "Pr(>F)"]
p_values <- c(dist_p_values[1:3], mxd_p_values[1:3], lga_p_values[1:3])
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Mixed lakes
SD_SRic_geo_M_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[mixed_lakes,])
shapiro.test(residuals(SD_SRic_geo_M_lm))
summary(SD_SRic_geo_M_lm)
car::Anova(SD_SRic_geo_M_lm, type = 3)
p_values <- summary(SD_SRic_geo_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

### Stratified lakes
SD_SRic_geo_S_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[stratified_lakes,])
shapiro.test(residuals(SD_SRic_geo_S_lm))
summary(SD_SRic_geo_S_lm)
car::Anova(SD_SRic_geo_S_lm, type = 3)
p_values <- summary(SD_SRic_geo_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
par(mfrow=c(1,1))
```

## SD beta diversity
### SD beta diversity distance calculations plus dispersion and PERMANOVA
```{r, SD beta distances}
### Regular
# Reference
SD_beta_ref_dist <- BAT::beta(presabs_lake, abund = F)
(SD_beta_ref_BD <- betadisper(SD_beta_ref_dist$Btotal, Location_type_group_ref))
(SD_beta_ref_AOV <- anova(SD_beta_ref_BD))
(SD_beta_ref_THSD <- TukeyHSD(SD_beta_ref_BD))
(SD_beta_ref_PM <- adonis2(SD_beta_ref_dist$Btotal ~ env[,"Location_type"], permutations = 999))
(SD_beta_ref_PM_pair <- pairwise.adonis(SD_beta_ref_dist$Btotal, env[,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites total
SD_beta_dist <- BAT::beta(surveyed_sites_lake, abund = F)
(SD_beta_BD <- betadisper(SD_beta_dist$Btotal, Location_type_group))
(SD_beta_AOV <- anova(SD_beta_BD))
(SD_beta_THSD <- TukeyHSD(SD_beta_BD))
(SD_beta_PM <- adonis2(SD_beta_dist$Btotal ~ env[surveyed_sites,"Location_type"], permutations = 999, method = "euclidean"))
(SD_beta_PM_pair <- pairwise.adonis(SD_beta_dist$Btotal, env[surveyed_sites,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites replacement
(SD_beta_rep_BD <- betadisper(SD_beta_dist$Brepl, Location_type_group))
(SD_beta_rep_AOV <- anova(SD_beta_rep_BD))
(SD_beta_rep_THSD <- TukeyHSD(SD_beta_rep_BD))
(SD_beta_rep_PM <- adonis2(SD_beta_dist$Brepl ~ env[surveyed_sites,"Location_type"], permutations = 999, method = "euclidean"))
(SD_beta_rep_PM_pair <- pairwise.adonis(SD_beta_dist$Brepl, env[surveyed_sites,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites richness
(SD_beta_ric_BD <- betadisper(SD_beta_dist$Brich, Location_type_group))
(SD_beta_ric_AOV <- anova(SD_beta_ric_BD))
(SD_beta_ric_THSD <- TukeyHSD(SD_beta_ric_BD))
(SD_beta_ric_PM <- adonis2(SD_beta_dist$Brich ~ env[surveyed_sites,"Location_type"], permutations = 999, method = "euclidean"))
(SD_beta_ric_PM_pair <- pairwise.adonis(SD_beta_dist$Brich, env[surveyed_sites,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Without LCN
SD_beta_wo_LCN_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_LCN,], abund = F)
(SD_beta_wo_LCN_BD <- betadisper(SD_beta_wo_LCN_dist$Btotal, Location_type_group[-8]))
(SD_beta_wo_LCN_AOV <- anova(SD_beta_wo_LCN_BD))
(SD_beta_wo_LCN_THSD <- TukeyHSD(SD_beta_wo_LCN_BD))
(SD_beta_wo_LCN_PM <- adonis2(SD_beta_wo_LCN_dist$Btotal ~ env[surveyed_sites_wo_LCN,"Location_type"], permutations = 999))
(SD_beta_wo_LCN_PM_pair <- pairwise.adonis(SD_beta_wo_LCN_dist$Btotal, env[surveyed_sites_wo_LCN,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Without OCO
SD_beta_wo_OCO_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_OCO,], abund = F)
(SD_beta_wo_OCO_BD <- betadisper(SD_beta_wo_OCO_dist$Btotal, Location_type_group[-8]))
(SD_beta_wo_OCO_AOV <- anova(SD_beta_wo_OCO_BD))
(SD_beta_wo_OCO_THSD <- TukeyHSD(SD_beta_wo_OCO_BD))
(SD_beta_wo_OCO_PM <- adonis2(SD_beta_wo_OCO_dist$Btotal ~ env[surveyed_sites_wo_OCO,"Location_type"], permutations = 999))
(SD_beta_wo_OCO_PM_pair <- pairwise.adonis(SD_beta_wo_OCO_dist$Btotal, env[surveyed_sites_wo_OCO,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Without TLN and HLM
SD_beta_wo_TLN_HLM_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_TLN_HLM,], abund = F)
(SD_beta_wo_TLN_HLM_BD <- betadisper(SD_beta_wo_TLN_HLM_dist$Btotal, Location_type_group[-c(5,21)]))
(SD_beta_wo_TLN_HLM_AOV <- anova(SD_beta_wo_TLN_HLM_BD))
(SD_beta_wo_TLN_HLM_THSD <- TukeyHSD(SD_beta_wo_TLN_HLM_BD))
(SD_beta_wo_TLN_HLM_PM <- adonis2(SD_beta_wo_TLN_HLM_dist$Btotal ~ env[surveyed_sites_wo_TLN_HLM,"Location_type"], permutations = 999))
(SD_beta_wo_TLN_HLM_PM_pair <- pairwise.adonis(SD_beta_wo_TLN_HLM_dist$Btotal, env[surveyed_sites_wo_TLN_HLM,"Location_type"], p.adjust.m = "bonferroni", perm = 999))
# Mixed and stratified lakes
SD_beta_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], abund = F)
# Ocean sites and mixed lakes
SD_beta_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites,], abund = F)
# Stratified lakes and ocean sites
SD_beta_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites,], abund = F)
# Mixed lakes
SD_beta_M_dist<- BAT::beta(surveyed_sites_lake[mixed_lakes,], abund = F)
# Stratified lakes
SD_beta_S_dist<- BAT::beta(surveyed_sites_lake[stratified_lakes,], abund = F)


### Environmental
# Surveyed sites
SD_beta_env_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_env,], abund = F)
# Mixed and stratified lakes
SD_beta_env_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], abund = F)
# Ocean sites and mixed lakes
SD_beta_env_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites_env,], abund = F)
# Stratified lakes and ocean sites
SD_beta_env_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites_env,], abund = F)


### Geographic
# Surveyed sites
SD_beta_geo_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites,], abund = F)
# Mixed and stratified lakes
SD_beta_geo_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], abund = F)
# Ocean sites and mixed lakes
SD_beta_geo_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites,], abund = F)
# Stratified lakes and ocean sites
SD_beta_geo_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites,], abund = F)
```

### SD beta q values of dispersion statistics
```{r, SD beta dispersion q-values}
# Sites
N <- 22
# Categories
k <- 3
# Sum of squares
SS <- SD_beta_AOV$`Sum Sq`[2]
# Mean of squares
MS <- SD_beta_AOV$`Mean Sq`[2]
# q value
q.value <- qtukey(p = 0.95, nmeans = k, df = N - k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

group <- SD_beta_THSD$group

(OM_q <- (abs(group[1,1]))/USE)
(SO_q <- (abs(group[2,1]))/USE)
(MS_q <- (abs(group[3,1]))/BSE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx
```

### SD beta dispersions
- The dissimilarity between sites of the same Location type
```{r, SD beta dispersion boxplot}
# Dispersion within-groups
SD_beta_BD_dist <- SD_beta_BD$distances
SD_beta_BD_dist <- as.data.frame(SD_beta_BD_dist)
SD_beta_BD_dist$X <- row.names(SD_beta_BD_dist)
SD_beta_BD_dist_env <- merge(SD_beta_BD_dist, env[surveyed_sites,], by = "X", sort = F)

# Add levels to Location_type column
SD_beta_BD_dist_env$Location_type <- factor(SD_beta_BD_dist_env$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Determine outliers based on dispersion of sites by Location type
outlier_SD_beta_BD_dist_env <- SD_beta_BD_dist_env %>%
  group_by(Location_type) %>%
  mutate(
    Q1 = quantile(SD_beta_BD_dist, 0.25),
    Q3 = quantile(SD_beta_BD_dist, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = SD_beta_BD_dist < lower_bound | SD_beta_BD_dist > upper_bound
  )

outlier_SD_beta_BD_dist_env$is_outlier

# Plot dispersion
(SD_beta_BD_dist_env_plot <- ggplot(SD_beta_BD_dist_env, aes(x = Location_type, y = SD_beta_BD_dist, color = Location_type, fill = Location_type)) +
  geom_violin(alpha = 0.6, draw_quantiles = c(0.25, 0.5, 0.75), aes(color = Location_type, fill = Location_type)) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  guides(fill = "none", color = "none") +
  geom_jitter(shape = 21,
    size = 4,
    alpha = 0.8,
    width = 0.1) +
  geom_text_repel(data = SD_beta_BD_dist_env, label = SD_beta_BD_dist_env$X, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme_bw() +
  theme(text = element_text(size = 17), legend.text = element_text(size = 17),
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(), axis.text = element_text(size = 17, color = "black"),
  axis.line = element_line(color = "black")) +
  scale_y_continuous(expand = c(0,0.05)) +
  xlab("Location type") +
  ylab("Distance to Centroid") +
  labs(color = "Location_type", tag = "a"))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_BD_dist_plot.jpg", SD_beta_BD_dist_env_plot, width = 5.4, height = 6, units = "in")
```

### SD beta NMDS
```{r, SD beta NMDS, results=FALSE}
### Regular
# Reference
SD_beta_ref_NMDS <- metaMDS(SD_beta_ref_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Surveyed sites 
SD_beta_NMDS <- metaMDS(SD_beta_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
round(SD_beta_NMDS$stress, digits = 2)
SD_beta_rep_NMDS <- metaMDS(SD_beta_dist$Brepl, try = 1000, parallel = 4, trymax = 500, maxit = 500)
SD_beta_ric_NMDS <- metaMDS(SD_beta_dist$Brich, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Without LCN
SD_beta_wo_LCN_NMDS <- metaMDS(SD_beta_wo_LCN_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Without TLN and HLM
SD_beta_wo_TLN_HLM_NMDS <- metaMDS(SD_beta_wo_TLN_HLM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
SD_beta_MS_NMDS <- metaMDS(SD_beta_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
SD_beta_OM_NMDS <- metaMDS(SD_beta_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
SD_beta_SO_NMDS <- metaMDS(SD_beta_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed lakes
SD_beta_M_NMDS <- metaMDS(SD_beta_M_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes
SD_beta_S_NMDS <- metaMDS(SD_beta_S_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)


### Environmental
# Surveyed sites 
SD_beta_env_NMDS <- metaMDS(SD_beta_env_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
SD_beta_env_MS_NMDS <- metaMDS(SD_beta_env_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
SD_beta_env_OM_NMDS <- metaMDS(SD_beta_env_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
SD_beta_env_SO_NMDS <- metaMDS(SD_beta_env_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)


### Geographic
# Surveyed sites 
SD_beta_geo_NMDS <- metaMDS(SD_beta_geo_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
SD_beta_geo_MS_NMDS <- metaMDS(SD_beta_geo_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
SD_beta_geo_OM_NMDS <- metaMDS(SD_beta_geo_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
SD_beta_geo_SO_NMDS <- metaMDS(SD_beta_geo_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
```

### Load an envfit analysis function for adjusting p-values
- Need to run this function for downstream analyses in this Markdown file
```{r, load envfit p-value function}
# Function p.adjust.envfit
# Calculates adjusted P values for results stored in envfit object,
# created using envfit function from vegan, which fits supplementary variables
# onto axes of unconstrained ordination. 
# Arguments: 
# x - envfit object
# method - method for correction of multiple testing issue (default = 'bonferroni',
#    see ''?p.adjust' for more options)
# n - optional, number of tests for which to correct; if not given, the number is
#    taken as the number of tests conducted by envfit function (for both vectors and factors).
# Author: David Zeleny
p.adjust.envfit <- function (x, method = 'bonferroni', n)
{
  x.new <- x
  if (!is.null (x$vectors)) pval.vectors <- x$vectors$pvals else pval.vectors <- NULL
  if (!is.null (x$factors)) pval.factors <- x$factors$pvals else pval.factors <- NULL
  if (missing (n)) n <- length (pval.vectors) + length (pval.factors)
  if (!is.null (x$vectors)) x.new$vectors$pvals <- p.adjust (x$vectors$pvals, method = method, n = n)
  if (!is.null (x$factors)) x.new$factors$pvals <- p.adjust (x$factors$pvals, method = method, n = n)
  cat ('Adjustment of significance by', method, 'method')
  return (x.new)
}
```

### SD beta varitation partitioning of env and geo variables
```{r}
env_var <- env[surveyed_sites_env,environment]
geo_var <- env[surveyed_sites_env,geography]
SD_beta_varpart <- varpart(SD_beta_env_dist$Btotal, env_var, geo_var)
SD_beta_varpart$part

# Open a jpg device
png("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_varpart.jpg", width = 4.5, height = 4.5, units = "in", res = 300, type = "cairo")
# Plot the variation partitioning results
par(mar = c(3, 3, 3, 3) + 1)  # Increase bottom margin if needed
plot(SD_beta_varpart,
     Xnames = c("Env", "Geo"),
     bg = c("mediumpurple", "orange"), alpha = 80,
     digits = 1,
     asp = 1)
# Close the jpg device
dev.off()

plot(SD_beta_varpart,
     Xnames = c("Env", "Geo"),
     bg = c("mediumpurple", "orange"), alpha = 80,
     digits = 1,
     asp = 1)
```

### SD beta with env and geo correlated variables using envfit
```{r, SD beta envfit}
### Environmental
# Surveyed sites 
# For figure
(SD_beta_env_ef <- envfit(SD_beta_env_NMDS, env[surveyed_sites_env,"S"], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Location_type"]))
(SD_beta_env_efp <- p.adjust.envfit(SD_beta_env_ef))
# Surveyed sites by Location_type
(SD_beta_env_A_ef <- envfit(SD_beta_env_NMDS, env[surveyed_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Location_type"]))
(SD_beta_env_A_efp <- p.adjust.envfit(SD_beta_env_A_ef))
# Mixed and stratified lakes
(SD_beta_env_MS_ef <- envfit(SD_beta_env_MS_NMDS, env[mixed_stratified_lakes,environment], permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
(SD_beta_env_MS_efp <- p.adjust.envfit(SD_beta_env_MS_ef))
# Ocean sites and mixed lakes
(SD_beta_env_OM_ef <- envfit(SD_beta_env_OM_NMDS, env[ocean_mixed_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Location_type"]))
(SD_beta_env_OM_efp <- p.adjust.envfit(SD_beta_env_OM_ef))
# Stratified lakes and ocean sites
(SD_beta_env_SO_ef <- envfit(SD_beta_env_SO_NMDS, env[ocean_stratified_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Location_type"]))
(SD_beta_env_SO_efp <- p.adjust.envfit(SD_beta_env_SO_ef))


# Mixed lakes
(SD_beta_M_ef <- envfit(SD_beta_M_NMDS, env[mixed_lakes,c("temperature_median","salinity_median","oxygen_median","distance_to_ocean_min_m","max_depth","logArea")], permutations = 999, na.rm = TRUE))
(SD_beta_M_efp <- p.adjust.envfit(SD_beta_M_ef))


# Stratified lakes
(SD_beta_S_ef <- envfit(SD_beta_S_NMDS, env[stratified_lakes,c("temperature_median","salinity_median","oxygen_median","distance_to_ocean_min_m","max_depth","logArea")], permutations = 999, na.rm = TRUE))
(SD_beta_S_efp <- p.adjust.envfit(SD_beta_S_ef))


### Geographic
# Surveyed sites
# For figure
(SD_beta_geo_ef <- envfit(SD_beta_geo_NMDS, env[surveyed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Location_type"]))
(SD_beta_geo_efp <- p.adjust.envfit(SD_beta_geo_ef))
# Surveyed sites by Location_type
(SD_beta_geo_A_ef <- envfit(SD_beta_geo_NMDS, env[surveyed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Location_type"]))
(SD_beta_geo_A_efp <- p.adjust.envfit(SD_beta_geo_A_ef))
# Mixed and stratified lakes
(SD_beta_geo_MS_ef <- envfit(SD_beta_geo_MS_NMDS, env[mixed_stratified_lakes,geography], permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
(SD_beta_geo_MS_efp <- p.adjust.envfit(SD_beta_geo_MS_ef))
# Ocean sites and mixed lakes
(SD_beta_geo_OM_ef <- envfit(SD_beta_geo_OM_NMDS, env[ocean_mixed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Location_type"]))
(SD_beta_geo_OM_efp <- p.adjust.envfit(SD_beta_geo_OM_ef))
# Stratified lakes and ocean sites
(SD_beta_geo_SO_ef <- envfit(SD_beta_geo_SO_NMDS, env[ocean_stratified_sites,geography], permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Location_type"]))
(SD_beta_geo_SO_efp <- p.adjust.envfit(SD_beta_geo_SO_ef))
```

### SD beta Mantel correlation tests
```{r, SD beta Mantel tests}
### Environmental
# Surveyed sites
env_dist_t <- dist(scaled_env[surveyed_sites_env,"temperature_median"], method = "euclidean")
(SD_beta_env_mant_t <- mantel(SD_beta_env_dist$Btotal, env_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Location_type"]))
env_dist_s <- dist(scaled_env[surveyed_sites_env,"salinity_median"], method = "euclidean")
(SD_beta_env_mant_s <- mantel(SD_beta_env_dist$Btotal, env_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Location_type"]))
env_dist_o <- dist(scaled_env[surveyed_sites_env,"oxygen_median"], method = "euclidean")
(SD_beta_env_mant_o <- mantel(SD_beta_env_dist$Btotal, env_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,"Location_type"]))
# Adjust p-values
SD_beta_env_mant_pv <- rbind(SD_beta_env_mant_t$signif, SD_beta_env_mant_s$signif, SD_beta_env_mant_o$signif)
SD_beta_env_mant_pv <- SD_beta_env_mant_pv[,1]
(SD_beta_env_mant_pv <- p.adjust(SD_beta_env_mant_pv, method = "bonferroni"))

# Mixed and stratified lakes
env_MS_dist_t <- dist(scaled_env[mixed_stratified_lakes,"temperature_median"], method = "euclidean")
(SD_beta_env_MS_mant_t <- mantel(SD_beta_env_MS_dist$Btotal, env_MS_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
env_MS_dist_s <- dist(scaled_env[mixed_stratified_lakes,"salinity_median"], method = "euclidean")
(SD_beta_env_MS_mant_s <- mantel(SD_beta_env_MS_dist$Btotal, env_MS_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
env_MS_dist_o <- dist(scaled_env[mixed_stratified_lakes,"oxygen_median"], method = "euclidean")
(SD_beta_env_MS_mant_o <- mantel(SD_beta_env_MS_dist$Btotal, env_MS_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
# Adjust p-values
SD_beta_env_MS_mant_pv <- rbind(SD_beta_env_MS_mant_t$signif, SD_beta_env_MS_mant_s$signif, SD_beta_env_MS_mant_o$signif)
SD_beta_env_MS_mant_pv <- SD_beta_env_MS_mant_pv[,1]
(SD_beta_env_MS_mant_pv <- p.adjust(SD_beta_env_MS_mant_pv, method = "bonferroni"))

# Ocean sites and mixed lakes
env_OM_dist_t <- dist(scaled_env[ocean_mixed_sites_env,"temperature_median"], method = "euclidean")
(SD_beta_env_OM_mant_t <- mantel(SD_beta_env_OM_dist$Btotal, env_OM_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Location_type"]))
env_OM_dist_s <- dist(scaled_env[ocean_mixed_sites_env,"salinity_median"], method = "euclidean")
(SD_beta_env_OM_mant_s <- mantel(SD_beta_env_OM_dist$Btotal, env_OM_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Location_type"]))
env_OM_dist_o <- dist(scaled_env[ocean_mixed_sites_env,"oxygen_median"], method = "euclidean")
(SD_beta_env_OM_mant_o <- mantel(SD_beta_env_OM_dist$Btotal, env_OM_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,"Location_type"]))
# Adjust p-values
SD_beta_env_OM_mant_pv <- rbind(SD_beta_env_OM_mant_t$signif, SD_beta_env_OM_mant_s$signif, SD_beta_env_OM_mant_o$signif)
SD_beta_env_OM_mant_pv <- SD_beta_env_OM_mant_pv[,1]
(SD_beta_env_OM_mant_pv <- p.adjust(SD_beta_env_OM_mant_pv, method = "bonferroni"))

# Stratified lakes and ocean sites
env_SO_dist_t <- dist(scaled_env[ocean_stratified_sites_env,"temperature_median"], method = "euclidean")
(SD_beta_env_SO_mant_t <- mantel(SD_beta_env_SO_dist$Btotal, env_SO_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Location_type"]))
env_SO_dist_s <- dist(scaled_env[ocean_stratified_sites_env,"salinity_median"], method = "euclidean")
(SD_beta_env_SO_mant_s <- mantel(SD_beta_env_SO_dist$Btotal, env_SO_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Location_type"]))
env_SO_dist_o <- dist(scaled_env[ocean_stratified_sites_env,"oxygen_median"], method = "euclidean")
(SD_beta_env_SO_mant_o <- mantel(SD_beta_env_SO_dist$Btotal, env_SO_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,"Location_type"]))
# Adjust p-values
SD_beta_env_SO_mant_pv <- rbind(SD_beta_env_SO_mant_t$signif, SD_beta_env_SO_mant_s$signif, SD_beta_env_SO_mant_o$signif)
SD_beta_env_SO_mant_pv <- SD_beta_env_SO_mant_pv[,1]
(SD_beta_env_SO_mant_pv <- p.adjust(SD_beta_env_SO_mant_pv, method = "bonferroni"))


# Mixed lakes
env_M_dist_t <- dist(scaled_env[mixed_lakes,"temperature_median"], method = "euclidean")
(SD_beta_M_mant_t <- mantel(SD_beta_M_dist$Btotal, env_M_dist_t, method = "spearman", permutations = 999, na.rm = TRUE))
env_M_dist_s <- dist(scaled_env[mixed_lakes,"salinity_median"], method = "euclidean")
(SD_beta_M_mant_s <- mantel(SD_beta_M_dist$Btotal, env_M_dist_s, method = "spearman", permutations = 999, na.rm = TRUE))
env_M_dist_o <- dist(scaled_env[mixed_lakes,"oxygen_median"], method = "euclidean")
(SD_beta_M_mant_o <- mantel(SD_beta_M_dist$Btotal, env_M_dist_o, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_dm <- dist(scaled_env[mixed_lakes,"distance_to_ocean_min_m"], method = "euclidean")
(SD_beta_M_mant_dm <- mantel(SD_beta_M_dist$Btotal, geo_M_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_md <- dist(scaled_env[mixed_lakes,"max_depth"], method = "euclidean")
(SD_beta_M_mant_md <- mantel(SD_beta_M_dist$Btotal, geo_M_dist_md, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_la <- dist(scaled_env[mixed_lakes,"logArea"], method = "euclidean")
(SD_beta_M_mant_la <- mantel(SD_beta_M_dist$Btotal, geo_M_dist_la, method = "spearman", permutations = 999, na.rm = TRUE))
# Adjust p-values
SD_beta_M_mant_pv <- rbind(SD_beta_M_mant_t$signif, SD_beta_M_mant_s$signif, SD_beta_M_mant_o$signif, SD_beta_M_mant_dm$signif, SD_beta_M_mant_md$signif, SD_beta_M_mant_la$signif)
SD_beta_M_mant_pv <- SD_beta_M_mant_pv[,1]
(SD_beta_M_mant_pv <- p.adjust(SD_beta_M_mant_pv, method = "bonferroni"))


# Stratified lakes
env_S_dist_t <- dist(scaled_env[stratified_lakes,"temperature_median"], method = "euclidean")
(SD_beta_S_mant_t <- mantel(SD_beta_S_dist$Btotal, env_S_dist_t, method = "spearman", permutations = 999, na.rm = TRUE))
env_S_dist_s <- dist(scaled_env[stratified_lakes,"salinity_median"], method = "euclidean")
(SD_beta_S_mant_s <- mantel(SD_beta_S_dist$Btotal, env_S_dist_s, method = "spearman", permutations = 999, na.rm = TRUE))
env_S_dist_o <- dist(scaled_env[stratified_lakes,"oxygen_median"], method = "euclidean")
(SD_beta_S_mant_o <- mantel(SD_beta_S_dist$Btotal, env_S_dist_o, method = "spearman", permutations = 999, na.rm = TRUE))
geo_S_dist_dm <- dist(scaled_env[stratified_lakes,"distance_to_ocean_min_m"], method = "euclidean")
(SD_beta_S_mant_dm <- mantel(SD_beta_S_dist$Btotal, geo_S_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE))
geo_S_dist_md <- dist(scaled_env[stratified_lakes,"max_depth"], method = "euclidean")
(SD_beta_S_mant_md <- mantel(SD_beta_S_dist$Btotal, geo_S_dist_md, method = "spearman", permutations = 999, na.rm = TRUE))
geo_S_dist_la <- dist(scaled_env[stratified_lakes,"logArea"], method = "euclidean")
(SD_beta_S_mant_la <- mantel(SD_beta_S_dist$Btotal, geo_S_dist_la, method = "spearman", permutations = 999, na.rm = TRUE))
# Adjust p-values
SD_beta_S_mant_pv <- rbind(SD_beta_S_mant_t$signif, SD_beta_S_mant_s$signif, SD_beta_S_mant_o$signif, SD_beta_S_mant_dm$signif, SD_beta_S_mant_md$signif, SD_beta_S_mant_la$signif)
SD_beta_S_mant_pv <- SD_beta_S_mant_pv[,1]
(SD_beta_S_mant_pv <- p.adjust(SD_beta_S_mant_pv, method = "bonferroni"))


### Geographic
# Surveyed sites
geo_dist_dm <- dist(scaled_env[surveyed_sites,"distance_to_ocean_min_m"], method = "euclidean")
(SD_beta_geo_mant_dm <- mantel(SD_beta_geo_dist$Btotal, geo_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Location_type"]))
geo_dist_md <- dist(scaled_env[surveyed_sites,"max_depth"], method = "euclidean")
(SD_beta_geo_mant_md <- mantel(SD_beta_geo_dist$Btotal, geo_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Location_type"]))
geo_dist_la <- dist(scaled_env[surveyed_sites,"logArea"], method = "euclidean")
(SD_beta_geo_mant_la <- mantel(SD_beta_geo_dist$Btotal, geo_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,"Location_type"]))
# Adjust p-values
SD_beta_geo_mant_pv <- rbind(SD_beta_geo_mant_dm$signif, SD_beta_geo_mant_md$signif, SD_beta_geo_mant_la$signif)
SD_beta_geo_mant_pv <- SD_beta_geo_mant_pv[,1]
(SD_beta_geo_mant_pv <- p.adjust(SD_beta_geo_mant_pv, method = "bonferroni"))

# Mixed and stratified lakes 
geo_MS_dist_dm <- dist(scaled_env[mixed_stratified_lakes,"distance_to_ocean_min_m"], method = "euclidean")
(SD_beta_geo_MS_mant_dm <- mantel(SD_beta_geo_MS_dist$Btotal, geo_MS_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
geo_MS_dist_md <- dist(scaled_env[mixed_stratified_lakes,"max_depth"], method = "euclidean")
(SD_beta_geo_MS_mant_md <- mantel(SD_beta_geo_MS_dist$Btotal, geo_MS_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
geo_MS_dist_la <- dist(scaled_env[mixed_stratified_lakes,"logArea"], method = "euclidean")
(SD_beta_geo_MS_mant_la <- mantel(SD_beta_geo_MS_dist$Btotal, geo_MS_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,"Location_type"]))
# Adjust p-values
SD_beta_geo_MS_mant_pv <- rbind(SD_beta_geo_MS_mant_dm$signif, SD_beta_geo_MS_mant_md$signif, SD_beta_geo_MS_mant_la$signif)
SD_beta_geo_MS_mant_pv <- SD_beta_geo_MS_mant_pv[,1]
(SD_beta_geo_MS_mant_pv <- p.adjust(SD_beta_geo_MS_mant_pv, method = "bonferroni"))

# Ocean sites and mixed lakes
geo_OM_dist_dm <- dist(scaled_env[ocean_mixed_sites,"distance_to_ocean_min_m"], method = "euclidean")
(SD_beta_geo_OM_mant_dm <- mantel(SD_beta_geo_OM_dist$Btotal, geo_OM_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Location_type"]))
geo_OM_dist_md <- dist(scaled_env[ocean_mixed_sites,"max_depth"], method = "euclidean")
(SD_beta_geo_OM_mant_md <- mantel(SD_beta_geo_OM_dist$Btotal, geo_OM_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Location_type"]))
geo_OM_dist_la <- dist(scaled_env[ocean_mixed_sites,"logArea"], method = "euclidean")
(SD_beta_geo_OM_mant_la <- mantel(SD_beta_geo_OM_dist$Btotal, geo_OM_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,"Location_type"]))
# Adjust p-values
SD_beta_geo_OM_mant_pv <- rbind( SD_beta_geo_OM_mant_dm$signif, SD_beta_geo_OM_mant_md$signif, SD_beta_geo_OM_mant_la$signif)
SD_beta_geo_OM_mant_pv <- SD_beta_geo_OM_mant_pv[,1]
(SD_beta_geo_OM_mant_pv <- p.adjust(SD_beta_geo_OM_mant_pv, method = "bonferroni"))

# Stratified lakes and ocean sites
geo_SO_dist_dm <- dist(scaled_env[ocean_stratified_sites,"distance_to_ocean_min_m"], method = "euclidean")
(SD_beta_geo_SO_mant_dm <- mantel(SD_beta_geo_SO_dist$Btotal, geo_SO_dist_dm, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Location_type"]))
geo_SO_dist_md <- dist(scaled_env[ocean_stratified_sites,"max_depth"], method = "euclidean")
(SD_beta_geo_SO_mant_md <- mantel(SD_beta_geo_SO_dist$Btotal, geo_SO_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Location_type"]))
geo_SO_dist_la <- dist(scaled_env[ocean_stratified_sites,"logArea"], method = "euclidean")
(SD_beta_geo_SO_mant_la <- mantel(SD_beta_geo_SO_dist$Btotal, geo_SO_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,"Location_type"]))
# Adjust p-values
SD_beta_geo_SO_mant_pv <- rbind( SD_beta_geo_SO_mant_dm$signif, SD_beta_geo_SO_mant_md$signif, SD_beta_geo_SO_mant_la$signif)
SD_beta_geo_SO_mant_pv <- SD_beta_geo_SO_mant_pv[,1]
(SD_beta_geo_SO_mant_pv <- p.adjust(SD_beta_geo_SO_mant_pv, method = "bonferroni"))
```

### SD beta NMDS ordination plots
```{r, SD beta NMDS plots}
# SD beta total NMDS scores
SD_beta_NMDS_data.scores <- as.data.frame(scores(SD_beta_NMDS))
SD_beta_NMDS_data.scores$Location_type <- env[surveyed_sites,"Location_type"]
SD_beta_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
SD_beta_NMDS_data.scores$Location_type <- factor(SD_beta_NMDS_data.scores$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Get significantly correlated environmental variables
SD_beta_env_ef_coord_cont <- as.data.frame(scores(SD_beta_env_ef, "vectors")) * ordiArrowMul(SD_beta_env_ef)
SD_beta_env_row_names <- c("S")
# Assign the new row names to the data frame
SD_beta_env_ef_coord_cont <- data.frame(row.names = SD_beta_env_row_names, SD_beta_env_ef_coord_cont)

# SD_beta_geo_ef_coord_cont <- as.data.frame(scores(SD_beta_geo_ef, "vectors")) * ordiArrowMul(SD_beta_geo_ef)
# SD_beta_geo_row_names <- c("minD")
# # Assign the new row names to the data frame
# SD_beta_geo_ef_coord_cont <- data.frame(row.names = SD_beta_geo_row_names, SD_beta_geo_ef_coord_cont)

# Plot NMDS ordination of SD beta total with CI = 0.95
(SD_beta_ef_plot <- ggplot(data = SD_beta_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.95) +
  geom_point(data = SD_beta_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 2, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
     data = SD_beta_env_ef_coord_cont, linewidth =1, alpha = 0.2, color = env_cont) +
  geom_text(data = SD_beta_env_ef_coord_cont, aes(x = NMDS1, y = NMDS2),
    color = env_cont, label = row.names(SD_beta_env_ef_coord_cont), size = 3) +
  geom_text_repel(data = SD_beta_NMDS_data.scores, label = SD_beta_NMDS_data.scores$Lakes, 
      size = 3, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 12), 
    legend.position = "bottom", 
    legend.title = element_text(size = 10), 
    legend.text = element_text(size = 10), 
    axis.title = element_text(size = 10),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black", size = 10), 
    legend.key = element_blank()) +
  annotate("text", x = -0.5, y = 0.55, size = 3.5,
     label = paste("Stress: ", round(SD_beta_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "a") +
  coord_fixed() +
  guides(color = "none", fill = "none") +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_NMDS.jpg", SD_beta_ef_plot, width = 3.25, height = 3.3, units = "in")


# Plot NMDS ordination of SD beta total with CI = 0.90
SD_beta_ef_plot_CI90 <- ggplot(data = SD_beta_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.90) +
  geom_point(data = SD_beta_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
     data = SD_beta_env_ef_coord_cont, linewidth =1, alpha = 0.2, color = env_cont) +
  geom_text(data = SD_beta_env_ef_coord_cont, aes(x = NMDS1, y = NMDS2),
    color = env_cont, label = row.names(SD_beta_env_ef_coord_cont), size = 5) +
  geom_text_repel(data = SD_beta_NMDS_data.scores, label = SD_beta_NMDS_data.scores$Lakes, 
      size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 17), 
    legend.position = "bottom", 
    legend.title = element_text(size = 17), 
    legend.text = element_text(size = 17), 
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black"), 
    legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
     label = paste("Stress: ", round(SD_beta_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(SD_beta_ef_plot_CI90 <- SD_beta_ef_plot_CI90 + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_NMDS_CI90.jpg", SD_beta_ef_plot_CI90, width = 5.4, height = 6, units = "in")


# SD beta total NMDS scores without LCN
SD_beta_wo_LCN_NMDS_data.scores <- as.data.frame(scores(SD_beta_wo_LCN_NMDS))
SD_beta_wo_LCN_NMDS_data.scores$Location_type <- env[surveyed_sites_wo_LCN,"Location_type"]
SD_beta_wo_LCN_NMDS_data.scores$Lakes <- env[surveyed_sites_wo_LCN,1]
SD_beta_wo_LCN_NMDS_data.scores$Location_type <- factor(SD_beta_wo_LCN_NMDS_data.scores$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta total without LCN with CI = 0.95
SD_beta_wo_LCN_plot <- ggplot(data = SD_beta_wo_LCN_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.95) +
  geom_point(data = SD_beta_wo_LCN_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_wo_LCN_NMDS_data.scores, label = SD_beta_wo_LCN_NMDS_data.scores$Lakes, 
      size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
    legend.position = "bottom", 
    legend.title = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black"), 
    legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
     label = paste("Stress: ", round(SD_beta_wo_LCN_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(SD_beta_wo_LCN_plot <- SD_beta_wo_LCN_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_NMDS_wo_LCN.jpg", SD_beta_wo_LCN_plot, width = 4.88, height = 6, units = "in")


# SD beta total NMDS scores without TLN and HLM
SD_beta_wo_TLN_HLM_NMDS_data.scores <- as.data.frame(scores(SD_beta_wo_TLN_HLM_NMDS))
SD_beta_wo_TLN_HLM_NMDS_data.scores$Location_type <- env[surveyed_sites_wo_TLN_HLM,"Location_type"]
SD_beta_wo_TLN_HLM_NMDS_data.scores$Lakes <- env[surveyed_sites_wo_TLN_HLM,1]
SD_beta_wo_TLN_HLM_NMDS_data.scores$Location_type <- factor(SD_beta_wo_TLN_HLM_NMDS_data.scores$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta total without TLN and HLM with CI = 0.95
SD_beta_wo_TLN_HLM_plot <- ggplot(data = SD_beta_wo_TLN_HLM_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.95) +
  geom_point(data = SD_beta_wo_TLN_HLM_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_wo_TLN_HLM_NMDS_data.scores, label = SD_beta_wo_TLN_HLM_NMDS_data.scores$Lakes, 
      size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
    legend.position = "bottom", 
    legend.title = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black"), 
    legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
     label = paste("Stress: ", round(SD_beta_wo_TLN_HLM_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(SD_beta_wo_TLN_HLM_plot <- SD_beta_wo_TLN_HLM_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_NMDS_wo_TLN_HLM.jpg", SD_beta_wo_TLN_HLM_plot, width = 4.88, height = 6, units = "in")


# SD beta replacement NMDS scores
SD_beta_rep_NMDS_data.scores <- as.data.frame(scores(SD_beta_rep_NMDS))
SD_beta_rep_NMDS_data.scores$Location_type <- env[surveyed_sites,"Location_type"]
SD_beta_rep_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
SD_beta_rep_NMDS_data.scores$Location_type <- factor(SD_beta_rep_NMDS_data.scores$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta replacement with CI = 0.95
SD_beta_rep_plot <- ggplot(data = SD_beta_rep_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.95) +
  geom_point(data = SD_beta_rep_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_rep_NMDS_data.scores, label = SD_beta_rep_NMDS_data.scores$Lakes, 
      size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
    legend.position = "bottom", 
    legend.title = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black"), 
    legend.key = element_blank()) +
  annotate("text", x = -0.7, y = 0.8, size = 5,
     label = paste("Stress: ", round(SD_beta_rep_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(SD_beta_rep_plot <- SD_beta_rep_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_rep_NMDS.jpg", SD_beta_rep_plot, width = 4.88, height = 6, units = "in")


# SD beta richness NMDS scores
SD_beta_ric_NMDS_data.scores <- as.data.frame(scores(SD_beta_ric_NMDS))
SD_beta_ric_NMDS_data.scores$Location_type <- env[surveyed_sites,"Location_type"]
SD_beta_ric_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
SD_beta_ric_NMDS_data.scores$Location_type <- factor(SD_beta_ric_NMDS_data.scores$Location_type, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta richness with CI = 0.95
SD_beta_ric_plot <- ggplot(data = SD_beta_ric_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.95) +
  geom_point(data = SD_beta_ric_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_ric_NMDS_data.scores, label = SD_beta_ric_NMDS_data.scores$Lakes, 
      size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
    legend.position = "bottom", 
    legend.title = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black"), 
    legend.key = element_blank()) +
  annotate("text", x = -0.7, y = 0.8, size = 5,
     label = paste("Stress: ", round(SD_beta_ric_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(SD_beta_ric_plot <- SD_beta_ric_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_ric_NMDS.jpg", SD_beta_ric_plot, width = 4.88, height = 6, units = "in")


# SD beta ref NMDS scores
SD_beta_ref_NMDS_data.scores <- as.data.frame(scores(SD_beta_ref_NMDS))
SD_beta_ref_NMDS_data.scores$Location_type <- env[,"Location_type"]
SD_beta_ref_NMDS_data.scores$Lakes <- env[,1]
SD_beta_ref_NMDS_data.scores$Location_type <- factor(SD_beta_ref_NMDS_data.scores$Location_type, levels = c("Reference", "Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta ref with CI = 0.95
SD_beta_ref_plot <- ggplot(data = SD_beta_ref_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Location_type)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Location_type), type = "t", level = 0.95) +
  geom_point(data = SD_beta_ref_NMDS_data.scores, aes(color = Location_type, fill = Location_type), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_ref_NMDS_data.scores, label = SD_beta_ref_NMDS_data.scores$Lakes, 
      size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
    legend.position = "right", 
    legend.title = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.border = element_rect(fill = NA), 
    axis.text = element_text(color = "black", size = 14), 
    legend.key = element_blank()) +
  annotate("text", x = -0.7, y = 0.8, size = 5,
     label = paste("Stress: ", round(SD_beta_ref_NMDS$stress, digits = 2))) +
  labs(colour = "Location type:  ", fill = "Location type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(SD_beta_ref_plot <- SD_beta_ref_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_ref_NMDS.jpg", SD_beta_ref_plot, width = 6.26, height = 6, units = "in")
```

### SD beta dendrogram
```{r, SD beta dendrogram}
# Cluster communities using average-linkage algorithm
SD_beta_dist_clust <- hclust(SD_beta_dist$Btotal, method = "average")

# The following will create a dendrogram of SD_beta putting things in one dimensional space
# Open a jpg device
png("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_dendrogram.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

# Create your plot using the plot() function
plot(SD_beta_dist_clust, 
   xlab = "Sites",
   ylab = "SD beta",
   main = "",
   sub = "",
   pch = 20,
   col= "black")

# Close the jpg device
dev.off()


# The following will graph and identify the best supported number of clusters of the dendrogram
# Open a jpg device
png("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_dg_silhouette.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

Si <- numeric(nrow(presabs_lake[surveyed_sites,]))
for (k in 2:(nrow(presabs_lake[surveyed_sites,])-1))
{
  sil<-silhouette(cutree(SD_beta_dist_clust, k=k), SD_beta_dist$Btotal)
  Si[k] <- summary(sil)$avg.width
}
k.best<-which.max(Si)
plot(1:nrow(presabs_lake[surveyed_sites,]), Si, type = "h", main = "Silhouette", xlab = "K", ylab = "Width")
axis(1, k.best, paste("optimum", k.best, sep="\n"), col = "red", font = 2, col.axis = "red")

# Close the jpg device
dev.off()


# Partitioning around mediods uses distances matrix to found how points according to silhouette are grouped around a centroid. The greater the number, the greater the distance from the centroid.
# Open a jpg device
png("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/SD_beta_dg_partitioning.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

SD_beta_pam <- pam(SD_beta_dist$Btotal,2)
plot(SD_beta_pam)

# Close the jpg device
dev.off()
```

### SD beta NMDS and distance outliers
```{r, SD beta outliers}
# Determine NMDS1 outliers
ordered_SD_beta_NMDS1_data.scores <- SD_beta_NMDS_data.scores[order(SD_beta_NMDS_data.scores$NMDS1), ]

outlier_SD_beta_NMDS1 <- ordered_SD_beta_NMDS1_data.scores %>%
  group_by(Location_type) %>%
  mutate(
  Q1 = quantile(NMDS1, 0.25),
  Q3 = quantile(NMDS1, 0.75),
  IQR = Q3 - Q1,
  lower_bound = Q1 - 1.5 * IQR,
  upper_bound = Q3 + 1.5 * IQR,
  is_outlier = NMDS1 < lower_bound | NMDS1 > upper_bound
  )

outlier_SD_beta_NMDS1 <- as.data.frame(outlier_SD_beta_NMDS1)

row.names(outlier_SD_beta_NMDS1) <- outlier_SD_beta_NMDS1$Lakes

# Create the plot
(outlier_SD_beta_NMDS1_plot <- ggplot(outlier_SD_beta_NMDS1, aes(x = Location_type, y = NMDS1, fill = Location_type)) +
  geom_violin(alpha = 0.9, draw_quantiles = c(0.25, 0.5, 0.75), aes(fill = Location_type)) +
  geom_jitter(aes(color = is_outlier), width = 0.1, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = outlier_SD_beta_NMDS1, label = outlier_SD_beta_NMDS1$Lakes, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme(text = element_text(size = 16),
    legend.position = "bottom",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16),
    axis.line = element_line(color = "black"),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text = element_text(color = "black", size = 16)) +
  scale_y_continuous(expand = c(0,0.05)) +
  guides(fill = "none") + 
  labs(y = "NMDS1 Distances", x = "Location type", color = "Outlier:", tag = "a"))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/outlier_SD_beta_NMDS1.jpg", outlier_SD_beta_NMDS1_plot, width = 4.88, height = 6, units = "in")


# Determine NMDS2 outliers
ordered_SD_beta_NMDS2_data.scores <- SD_beta_NMDS_data.scores[order(SD_beta_NMDS_data.scores$NMDS2), ]

outlier_SD_beta_NMDS2 <- ordered_SD_beta_NMDS2_data.scores %>%
  group_by(Location_type) %>%
  mutate(
  Q1 = quantile(NMDS2, 0.25),
  Q3 = quantile(NMDS2, 0.75),
  IQR = Q3 - Q1,
  lower_bound = Q1 - 1.5 * IQR,
  upper_bound = Q3 + 1.5 * IQR,
  is_outlier = NMDS2 < lower_bound | NMDS2 > upper_bound
  )

outlier_SD_beta_NMDS2 <- as.data.frame(outlier_SD_beta_NMDS2)

row.names(outlier_SD_beta_NMDS2) <- outlier_SD_beta_NMDS2$Lakes

# Create the plot
(outlier_SD_beta_NMDS2_plot <- ggplot(outlier_SD_beta_NMDS2, aes(x = Location_type, y = NMDS2, fill = Location_type)) +
  geom_violin(alpha = 0.9, draw_quantiles = c(0.25, 0.5, 0.75), aes(fill = Location_type)) +
  geom_jitter(aes(color = is_outlier), width = 0.1, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = outlier_SD_beta_NMDS2, label = outlier_SD_beta_NMDS2$Lakes, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme(text = element_text(size = 16),
    legend.position = "bottom",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16),
    axis.line = element_line(color = "black"),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text = element_text(color = "black", size = 16)) +
  scale_y_continuous(expand = c(0,0.05)) +
  guides(fill = "none") + 
  labs(y = "NMDS2 Distances", x = "Location type", color = "Outlier:", tag = "b"))
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/SD/outlier_SD_beta_NMDS2.jpg", outlier_SD_beta_NMDS2_plot, width = 4.88, height = 6, units = "in")
```

### Package and version info
```{r, session info}
sessionInfo()
```

#### Determine what families of fish are most present in the lakes
```{r, eval=F}
library(rfishbase)
species_sums_tax <- site_sums
species_sums_tax$X <- gsub("_", " ", species_sums_tax$X)
species_sums_tax <- species_sums_tax$X
species_sums_tax <- rfishbase::load_taxa() %>% 
  filter(Species %in% species_sums_tax) %>%
  collect()
species_sums_tax$Species <- gsub(" ", "_", species_sums_tax$Species)
species_sums_tax <- merge(species_sums_tax, species_sums, by.x = "Species", by.y = "X")
species_sums_tax <- species_sums_tax[, -c(2:4, 7:8)]
species_sums_tax <- species_sums_tax[order(species_sums_tax$row_sum),]

species_sums_tax_summary <- species_sums_tax %>%
  group_by(Family) %>%
  summarise(across(.cols = row_sum, .fns = sum, na.rm = TRUE))

family <- ggplot(data = species_sums_tax_summary) + 
  geom_bar(aes(x = reorder(Family, row_sum), y = row_sum),
  stat = 'identity',
  col="green", 
  fill="green4", 
  alpha = .5) + 
  theme_bw() +
  theme(plot.background = element_blank()
  ,panel.grid.major = element_blank()
  ,panel.grid.minor = element_blank()
  ,panel.border = element_blank()) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), axis.line = element_line(color = "black")) +
  labs(x="Family", y="Family Richness")
family
ggsave("family.jpg", family, width = 5, height = 4.5)

order <- ggplot(data = species_sums_tax) + 
  geom_bar(aes(x = reorder(Order, row_sum), y = row_sum),
  stat = 'identity',
  col="green", 
  fill="green4", 
  alpha = .5) + 
  theme_bw() +
  theme(plot.background = element_blank()
  ,panel.grid.major = element_blank()
  ,panel.grid.minor = element_blank()
  ,panel.border = element_blank()) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), axis.line = element_line(color = "black")) +
  labs(x="Order", y="Richness in Lakess")
order
#ggsave("order.jpg", order)

detach("package:rfishbase")
```
