---
title: "Species diversity analyses"
output: github_document
editor_options: 
  chunk_output_type: console
---
### Load packages and files
```{r, load packages and files}
# Load the knitr package if not already loaded
library(knitr)

# Source the R Markdown file
knit("/Users/bailey/Documents/research/fish_biodiversity/src/collection/load_collection_data.Rmd", output = "/Users/bailey/Documents/research/fish_biodiversity/src/collection/load_collection_data.md")

# Packages
library(picante)
library(vegan)
library(phytools)
library(ggplot2)
library(viridis)
library(ggrepel)
library(caret)
library(BAT)
library(pairwiseAdonis)
library(ggvenn)
library(dplyr)

# Site vectors
surveyed_sites <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOM", "OOO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_LCN <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOM", "OOO", "OTM", "RCA", "SLN", "TLN", "ULN")

surveyed_sites_wo_TLN_HLM <- c("BCM", "CLM", "FLK", "GLK", "HLO", "IBK", "LCN", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOM", "OOO", "OTM", "RCA", "SLN", "ULN")

surveyed_sites_env <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OTM", "RCA", "SLN", "TLN", "ULN")

ocean_mixed_sites <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA", "FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

ocean_mixed_sites_env <- c("IBK", "NCN", "RCA", "FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

ocean_stratified_sites <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA", "BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

ocean_stratified_sites_env <- c("IBK", "NCN", "RCA", "BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

mixed_stratified_lakes <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "LLN", "MLN", "NLK", "NLN", "NLU", "OLO", "OTM", "SLN", "TLN", "ULN")

ocean_sites <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA")

ocean_sites_env <- c("IBK", "NCN", "RCA")

mixed_lakes <- c("FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")

stratified_lakes <- c("BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")

# Define your custom colors
custom_colors <- c("Reference" = "black", "Ocean" = "#EE6363", "Mixed" = "#87CEFA", "Stratified" = "#6E8B3D")

env_cont <- "#FFB90F"

trait_cont <- "#68228B"

trait_cat <- "#DEB887"
```

### Edit input files
```{r, edit files}
strat_fish <- stratified_fish[,-c(1,10)]
mix_fish <- mixed_fish[,-c(1,10)]
oc_fish <- ocean_fish[,-c(1,8)]

env <- env[order(env$X),]
presabs_lake <- presabs_lake[order(row.names(presabs_lake)),]
surveyed_sites_lake <- surveyed_sites_lake[order(row.names(surveyed_sites_lake)),]

stratification_group_ref <- env[,19]
stratification_group <- env[surveyed_sites,19]
```

# Species Diversity
## SD alpha diversity
### SD alpha bar plot
```{r, SD alpha bar plot}
#Plot species richness with a bar plot
SR_env$Stratification <- factor(SR_env$Stratification, levels = c("Ocean", "Mixed", "Stratified"))

(SR_plot <- ggplot(data = SR_env[-20,], mapping = aes(x = reorder(X, row_sum, decreasing = T), y = row_sum, color = Stratification, fill = Stratification, )) + 
  geom_bar(stat = 'identity',
    alpha = 1) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16), legend.title = element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(x="Site", y="Species Richness", colour = "Site type", fill = "Site type"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SR_plot.jpg", plot = SR_plot, width = 6, height = 6, units = "in")
```

### SD alpha venn diagram plot
```{r, SD alpha venn diagram plot}
site_type_pres <- strat_presabs_lake[c(24:26),]
site_type_pres <- t(site_type_pres)
site_type_pres <- site_type_pres[which(rowSums(site_type_pres) > 0),]

# Convert numerical values to logical
site_type_pres_logical <- as.data.frame(site_type_pres > 0)

# Check the structure of the transformed data
str(site_type_pres_logical)

# Create the Venn diagram
(venn_plot <- ggvenn(site_type_pres_logical,
                    show_percentage = F,
                    fill_color = c("#EE6363", "#87CEFA", "#6E8B3D"),
                    fill_alpha = 0.7,
                    stroke_alpha = 0,
                    stroke_size = 0.5, 
                    set_name_size = 5,
                    text_size = 5))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/venn_plot.jpg", plot = venn_plot, width = 4.25, height = 4, units = "in")
```

### SD alpha subsample venn diagram plots
```{r, SD alpha venn diagram plots}
# Randomly sample six columns from stratified_fish and mixed_fish
sampled_stratified <- sample(names(strat_fish)[-1], 6)
sampled_mixed <- sample(names(mix_fish)[-1], 6)

# Select sampled columns from stratified_fish and mixed_fish
stratified_subset <- select(stratified_fish, all_of(sampled_stratified))
mixed_subset <- select(mixed_fish, all_of(sampled_mixed))

# Get rid of species not present in the 6 remaining sites
stratified_subset <- stratified_subset[which(rowSums(stratified_subset[, -1]) > 0),]
mixed_subset <- mixed_subset[which(rowSums(mixed_subset[, -1]) > 0),]

# Make site type columns in each dataframe
stratified_subset[,7] <- rowSums(stratified_subset[,c(1:6)])
colnames(stratified_subset)[colnames(stratified_subset) == "V7"] <- "Stratified lakes"

mixed_subset[,7] <- rowSums(mixed_subset[,c(1:6)])
colnames(mixed_subset)[colnames(mixed_subset) == "V7"] <- "Mixed lakes"

oc_fish[,7] <- rowSums(oc_fish[,c(1:6)])
colnames(oc_fish)[colnames(oc_fish) == "V7"] <- "Ocean sites"

# Merge the two data frames, replacing missing values with 0
oc_mix_merge <- merge(oc_fish, mixed_subset, by = 0, all = TRUE)
row.names(oc_mix_merge) <- oc_mix_merge$Row.names
oc_mix_merge <- oc_mix_merge[,-1]
oc_mix_strat_merge <- merge(oc_mix_merge, stratified_subset, by = 0, all = TRUE)
row.names(oc_mix_strat_merge) <- oc_mix_strat_merge$Row.names
oc_mix_strat_merge <- oc_mix_strat_merge[,-1]

# Keep only site type
keep <- c("Ocean sites", "Mixed lakes", "Stratified lakes")
site_type_pres_sample <- oc_mix_strat_merge[,keep]

# Replace NA with 0
site_type_pres_sample[is.na(site_type_pres_sample)] <- 0

## Remove all species not found in any locations
# Identifies which rows are greater than 0
site_type_pres_sample <- site_type_pres_sample[which(rowSums(site_type_pres_sample) > 0),]

site_type_pres_sample_logical <- as.data.frame(site_type_pres_sample > 0)

venn_plot <- ggvenn(site_type_pres_sample_logical,
                    show_percentage = F,
                    fill_color = c("#EE6363", "#87CEFA", "#6E8B3D"),
                    fill_alpha = 0.7,
                    stroke_alpha = 0,
                    stroke_size = 0.5, 
                    set_name_size = 5,
                    text_size = 5)
venn_plot
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/venn_plot_s10.jpg", plot = venn_plot, width = 4.25, height = 4, units = "in")
```

### SD alpha outliers for each stratification type
```{r, SD alpha outliers}
outlier_SD_alpha <- SR_env[surveyed_sites,] %>%
  group_by(Stratification) %>%
  mutate(
    Q1 = quantile(row_sum, 0.25),
    Q3 = quantile(row_sum, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = row_sum < lower_bound | row_sum > upper_bound
  )

outlier_SD_alpha <- as.data.frame(outlier_SD_alpha)
row.names(outlier_SD_alpha) <- outlier_SD_alpha$X

# Create the plot
(outlier_SD_alpha_plot <- ggplot(outlier_SD_alpha, aes(x = Stratification, y = row_sum, fill = Stratification)) +
  geom_violin(trim = FALSE) +
  geom_jitter(aes(color = is_outlier, fill = is_outlier), width = 0.2, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  theme(text = element_text(size = 16),
        legend.position = "right",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  scale_y_continuous(expand = c(0,0)) +
  guides(fill = "none") + 
  labs(y = "Species Richness", x = "Site type", color = "Outlier", fill = "Site type", tag = "a"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/outlier_SD_alpha.jpg", outlier_SD_alpha_plot, width = 6.26, height = 6, units = "in")
```

### Identify VIF of env and geo variables
```{r, env and geo VIFs}
# Determine which env variables have variance
environment <- c("temperature_median", "salinity_median", "oxygen_median", "pH_median")
nzv <- nearZeroVar(env[,environment])
nzv
# All env variables have variance

mod <-  lm(row_sum ~ temperature_median + salinity_median + oxygen_median + pH_median, SR_env)
summary(mod)
anova(mod)
rms::vif(mod)
car::vif(mod)

mod <-  lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env)
summary(mod)
rms::vif(mod)
car::vif(mod)
environment <- c("temperature_median", "salinity_median", "oxygen_median")

# Determine which geo variables have variance
geography <- c("volume_m3_w_chemocline", "volume_m3", "surface_area_m2", "distance_to_ocean_min_m", "distance_to_ocean_mean_m", "distance_to_ocean_median_m", "tidal_lag_time_minutes", "tidal_efficiency", "perimeter_fromSat", "max_depth", "logArea")
nzv <- nearZeroVar(env[,geography])
nzv
# All geo variables have variance

mod <-  lm(row_sum ~ volume_m3_w_chemocline + volume_m3 + surface_area_m2 + distance_to_ocean_min_m + distance_to_ocean_mean_m + distance_to_ocean_median_m + tidal_lag_time_minutes + tidal_efficiency + perimeter_fromSat + max_depth + logArea, SR_env)
summary(mod)
rms::vif(mod)
car::vif(mod)

mod <-  lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env)
summary(mod)
rms::vif(mod)
car::vif(mod)
geography <- c("distance_to_ocean_min_m", "max_depth", "logArea")
```

### SD alpha SRic & stratification
```{r, SD alpha & stratification}
### Stratification
# Run ANOVA
anova_logSRic_result <- aov(log(row_sum) ~ Stratification, data = SR_env[surveyed_sites,])
summary(anova_logSRic_result)

# Perform Tukey's HSD test
tukey_logSRic_result <- TukeyHSD(anova_logSRic_result)
print(tukey_logSRic_result)

# Sites
N <- length(anova_logSRic_result$residuals)
# Categories
k <- length(anova_logSRic_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_logSRic_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Stratification <- tukey_logSRic_result$Stratification

# q-values
(OM_q <- (abs(Stratification[1,1]))/USE)
(MS_q <- (abs(Stratification[3,1]))/BSE)
(SO_q <- (abs(Stratification[2,1]))/USE)


# Run ANOVA
anova_SRic_result <- aov(row_sum ~ Stratification, data = SR_env[surveyed_sites,])
summary(anova_SRic_result)

# Perform Tukey's HSD test
tukey_SRic_result <- TukeyHSD(anova_SRic_result)
print(tukey_SRic_result)

# Sites
N <- length(anova_SRic_result$residuals)
# Categories
k <- length(anova_SRic_result$coefficients)
# Sum of squares
SS <- sum(resid(anova_SRic_result)^2) # from anova, the residual SS
# Mean squares
MS <- SS/(N-k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

# Get differences
Stratification <- tukey_SRic_result$Stratification

# q-values
(OM_q <- (abs(Stratification[1,1]))/USE)
(MS_q <- (abs(Stratification[3,1]))/BSE)
(SO_q <- (abs(Stratification[2,1]))/USE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx
```

### SD alpha SRic env and geo linear models
```{r, SD alpha lm}
### Environmental
# Surveyed sites
SD_logSRic_env_am <- aov(log(row_sum) ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[surveyed_sites_env,])
summary(SD_logSRic_env_am)
p_values <- summary(SD_logSRic_env_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_env_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[surveyed_sites_env,])
summary(SD_logSRic_env_lm)
p_values <- summary(SD_logSRic_env_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_am <- aov(row_sum ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[surveyed_sites_env,])
summary(SD_SRic_env_am)
p_values <- summary(SD_SRic_env_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_env_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[surveyed_sites_env,])
summary(SD_SRic_env_lm)
p_values <- summary(SD_SRic_env_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Mixed and stratified lakes
SD_logSRic_env_MS_am <- aov(log(row_sum) ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_env_MS_am)
p_values <- summary(SD_logSRic_env_MS_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_env_MS_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_env_MS_lm)
anova(SD_logSRic_env_MS_lm)
p_values <- summary(SD_logSRic_env_MS_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_MS_am <- aov(row_sum ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[mixed_stratified_lakes,])
summary(SD_SRic_env_MS_am)
p_values <- summary(SD_SRic_env_MS_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_env_MS_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[mixed_stratified_lakes,])
summary(SD_SRic_env_MS_lm)
anova(SD_SRic_env_MS_lm)
p_values <- summary(SD_SRic_env_MS_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

# Ocean sites and mixed lakes
SD_logSRic_env_OM_am <- aov(log(row_sum) ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_env_OM_am)
p_values <- summary(SD_logSRic_env_OM_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_env_OM_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[ocean_mixed_sites_env,])
summary(SD_logSRic_env_OM_lm)
p_values <- summary(SD_logSRic_env_OM_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_OM_am <- aov(row_sum ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[ocean_mixed_sites_env,])
summary(SD_SRic_env_OM_am)
p_values <- summary(SD_SRic_env_OM_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_env_OM_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[ocean_mixed_sites_env,])
summary(SD_SRic_env_OM_lm)
p_values <- summary(SD_SRic_env_OM_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Stratified lakes and ocean sites
SD_logSRic_env_SO_am <- aov(log(row_sum) ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_env_SO_am)
p_values <- summary(SD_logSRic_env_SO_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_env_SO_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[ocean_stratified_sites_env,])
summary(SD_logSRic_env_SO_lm)
p_values <- summary(SD_logSRic_env_SO_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_SO_am <- aov(row_sum ~ (temperature_median + salinity_median + oxygen_median) * Stratification, data = SR_env[ocean_stratified_sites_env,])
summary(SD_SRic_env_SO_am)
p_values <- summary(SD_SRic_env_SO_am)[[1]][, "Pr(>F)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_env_SO_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[ocean_stratified_sites_env,])
summary(SD_SRic_env_SO_lm)
p_values <- summary(SD_SRic_env_SO_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Ocean sites
SD_logSRic_env_O_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[ocean_sites,])
summary(SD_logSRic_env_O_lm)
anova(SD_logSRic_env_O_lm)
p_values <- summary(SD_logSRic_env_O_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_O_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[ocean_sites,])
summary(SD_SRic_env_O_lm)
anova(SD_SRic_env_O_lm)
p_values <- summary(SD_SRic_env_O_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Mixed lakes
SD_logSRic_env_M_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[mixed_lakes,])
summary(SD_logSRic_env_M_lm)
anova(SD_logSRic_env_M_lm)
p_values <- summary(SD_logSRic_env_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_M_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[mixed_lakes,])
summary(SD_SRic_env_M_lm)
anova(SD_SRic_env_M_lm)
p_values <- summary(SD_SRic_env_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Stratified lakes
SD_logSRic_env_S_lm <- lm(log(row_sum) ~ temperature_median + salinity_median + oxygen_median, SR_env[stratified_lakes,])
summary(SD_logSRic_env_S_lm)
anova(SD_logSRic_env_S_lm)
p_values <- summary(SD_logSRic_env_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_env_S_lm <- lm(row_sum ~ temperature_median + salinity_median + oxygen_median, SR_env[stratified_lakes,])
summary(SD_SRic_env_S_lm)
anova(SD_SRic_env_S_lm)
p_values <- summary(SD_SRic_env_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


### Geographical
# Surveyed sites
SD_logSRic_geo_am <- aov(log(row_sum) ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[surveyed_sites,])
summary(SD_logSRic_geo_am)
p_values <- summary(SD_logSRic_geo_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_geo_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[surveyed_sites,])
summary(SD_logSRic_geo_lm)
p_values <- summary(SD_logSRic_geo_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_am <- aov(row_sum ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[surveyed_sites,])
summary(SD_SRic_geo_am)
p_values <- summary(SD_SRic_geo_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_geo_lm <- lm(row_sum ~  max_depth + logArea * Stratification, SR_env[surveyed_sites,])
summary(SD_SRic_geo_lm)
p_values <- summary(SD_SRic_geo_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Mixed and stratified lakes
SD_logSRic_geo_MS_am <- aov(log(row_sum) ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_geo_MS_am)
p_values <- summary(SD_logSRic_geo_MS_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_geo_MS_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[mixed_stratified_lakes,])
summary(SD_logSRic_geo_MS_lm)
p_values <- summary(SD_logSRic_geo_MS_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_MS_am <- aov(row_sum ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[mixed_stratified_lakes,])
summary(SD_SRic_geo_MS_am)
p_values <- summary(SD_SRic_geo_MS_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_geo_MS_lm <- lm(row_sum ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, SR_env[mixed_stratified_lakes,])
summary(SD_SRic_geo_MS_lm)
p_values <- summary(SD_SRic_geo_MS_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Ocean sites and mixed lakes
SD_logSRic_geo_OM_am <- aov(log(row_sum) ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[ocean_mixed_sites,])
summary(SD_logSRic_geo_OM_am)
p_values <- summary(SD_logSRic_geo_OM_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_geo_OM_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[ocean_mixed_sites,])
summary(SD_logSRic_geo_OM_lm)
p_values <- summary(SD_logSRic_geo_OM_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_OM_am <- aov(row_sum ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[ocean_mixed_sites,])
summary(SD_SRic_geo_OM_am)
p_values <- summary(SD_SRic_geo_OM_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_geo_OM_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[ocean_mixed_sites,])
summary(SD_SRic_geo_OM_lm)
p_values <- summary(SD_SRic_geo_OM_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Stratified lakes and ocean sites
SD_logSRic_geo_SO_am <- aov(log(row_sum) ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[ocean_stratified_sites,])
summary(SD_logSRic_geo_SO_am)
p_values <- summary(SD_logSRic_geo_SO_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_logSRic_geo_SO_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[ocean_stratified_sites,])
summary(SD_logSRic_geo_SO_lm)
p_values <- summary(SD_logSRic_geo_SO_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_SO_am <- aov(row_sum ~ (distance_to_ocean_min_m + max_depth + logArea) * Stratification, data = SR_env[ocean_stratified_sites,])
summary(SD_SRic_geo_SO_am)
p_values <- summary(SD_SRic_geo_SO_am)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
SD_SRic_geo_SO_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[ocean_stratified_sites,])
summary(SD_SRic_geo_SO_lm)
p_values <- summary(SD_SRic_geo_SO_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Ocean sites
SD_logSRic_geo_M_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[ocean_sites,])
summary(SD_logSRic_geo_M_lm)
anova(SD_logSRic_geo_M_lm)
p_values <- summary(SD_logSRic_geo_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_M_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[ocean_sites,])
summary(SD_SRic_geo_M_lm)
anova(SD_SRic_geo_M_lm)
p_values <- summary(SD_SRic_geo_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Mixed lakes
SD_logSRic_geo_M_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[mixed_lakes,])
summary(SD_logSRic_geo_M_lm)
anova(SD_logSRic_geo_M_lm)
p_values <- summary(SD_logSRic_geo_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_M_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[mixed_lakes,])
summary(SD_SRic_geo_M_lm)
anova(SD_SRic_geo_M_lm)
p_values <- summary(SD_SRic_geo_M_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))


# Stratified lakes
SD_logSRic_geo_S_lm <- lm(log(row_sum) ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[stratified_lakes,])
summary(SD_logSRic_geo_S_lm)
anova(SD_logSRic_geo_S_lm)
p_values <- summary(SD_logSRic_geo_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))

SD_SRic_geo_S_lm <- lm(row_sum ~ distance_to_ocean_min_m + max_depth + logArea, SR_env[stratified_lakes,])
summary(SD_SRic_geo_S_lm)
anova(SD_SRic_geo_S_lm)
p_values <- summary(SD_SRic_geo_S_lm)$coefficients[, "Pr(>|t|)"]
(adjusted_p_values <- p.adjust(p_values, method = "bonferroni"))
```

### SD alpha z-scores with env and geo correlated variables
```{r, SD alpha lm plots}
# Log Area
SD_alpha_LA_plot <- ggplot(data = SR_env[surveyed_sites,], mapping = aes(y = log(row_sum), x = logArea, color = Stratification, fill = Stratification)) + 
  geom_point(stat = 'identity',
    size = 4,
    alpha = 1) + 
  geom_text_repel(label = SR_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(x="Log Area (m"^"2"~")", y="Log SRic", colour = "Site type", fill = "Site type", tag = "c")
(SD_alpha_LA_plot <- SD_alpha_LA_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_alpha_LA_plot.jpg", plot = SD_alpha_LA_plot, width = 6.26, height = 6, units = "in")

# Distance from the ocean mean
SD_alpha_D_plot <- ggplot(data = SR_env[surveyed_sites,], mapping = aes(y = row_sum, x = distance_to_ocean_min_m, color = Stratification, fill = Stratification)) + 
  geom_point(stat = 'identity',
    size = 4,
    alpha = 1) + 
  geom_text_repel(label = SR_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(x="Isolation (m)", y="SRic", colour = "Site type", fill = "Site type", tag = "b")
(SD_alpha_D_plot <- SD_alpha_D_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_alpha_D_plot.jpg", plot = SD_alpha_D_plot, width = 6.26, height = 6, units = "in")

# Max depth
SD_alpha_MD_plot <- ggplot(data = SR_env[surveyed_sites,], mapping = aes(y = row_sum, x = max_depth, color = Stratification, fill = Stratification)) + 
  geom_point(stat = 'identity',
    size = 4,
    alpha = 1) + 
  geom_text_repel(label = SR_env[surveyed_sites,1], size = 5, point.padding = 3) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_bw() +
  theme(text = element_text(size = 16),legend.title =  element_text(size = 16), legend.text = element_text(size = 16),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black"),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  labs(x="Age (m)", y="SRic", colour = "Site type", fill = "Site type", tag = "a")
(SD_alpha_MD_plot <- SD_alpha_MD_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_alpha_MD_plot.jpg", plot = SD_alpha_MD_plot, width = 6.26, height = 6, units = "in")
```

## SD beta diversity
### SD beta diversity distance calculations plus dispersion and PERMANOVA
```{r, SD beta distances}
### Regular
# Reference
SD_beta_ref_dist <- BAT::beta(presabs_lake, abund = F)
(SD_beta_ref_BD <- betadisper(SD_beta_ref_dist$Btotal, stratification_group_ref))
(SD_beta_ref_AOV <- anova(SD_beta_ref_BD))
(SD_beta_ref_THSD <- TukeyHSD(SD_beta_ref_BD))
(SD_beta_ref_PM <- adonis2(SD_beta_ref_dist$Btotal ~ env[,19], permutations = 999))
(SD_beta_ref_PM_pair <- pairwise.adonis(SD_beta_ref_dist$Btotal, env[,19], p.adjust.m = "bonferroni", perm = 999))
# Surveyed sites
SD_beta_dist <- BAT::beta(surveyed_sites_lake, abund = F)
(SD_beta_BD <- betadisper(SD_beta_dist$Btotal, stratification_group))
(SD_beta_AOV <- anova(SD_beta_BD))
(SD_beta_THSD <- TukeyHSD(SD_beta_BD))
(SD_beta_PM <- adonis2(SD_beta_dist$Btotal ~ env[surveyed_sites,19], permutations = 999, method = "euclidean"))
(SD_beta_PM_pair <- pairwise.adonis(SD_beta_dist$Btotal, env[surveyed_sites,19], p.adjust.m = "bonferroni", perm = 999))
# Without LCN
SD_beta_wo_LCN_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_LCN,], abund = F)
(SD_beta_wo_LCN_BD <- betadisper(SD_beta_wo_LCN_dist$Btotal, stratification_group[-8]))
(SD_beta_wo_LCN_AOV <- anova(SD_beta_wo_LCN_BD))
(SD_beta_wo_LCN_THSD <- TukeyHSD(SD_beta_wo_LCN_BD))
(SD_beta_wo_LCN_PM <- adonis2(SD_beta_wo_LCN_dist$Btotal ~ env[surveyed_sites_wo_LCN,19], permutations = 999))
(SD_beta_wo_LCN_PM_pair <- pairwise.adonis(SD_beta_wo_LCN_dist$Btotal, env[surveyed_sites_wo_LCN,19], p.adjust.m = "bonferroni", perm = 999))
# Without TLN and HLM
SD_beta_wo_TLN_HLM_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_wo_TLN_HLM,], abund = F)
(SD_beta_wo_TLN_HLM_BD <- betadisper(SD_beta_wo_TLN_HLM_dist$Btotal, stratification_group[-c(5,21)]))
(SD_beta_wo_TLN_HLM_AOV <- anova(SD_beta_wo_TLN_HLM_BD))
(SD_beta_wo_TLN_HLM_THSD <- TukeyHSD(SD_beta_wo_TLN_HLM_BD))
(SD_beta_wo_TLN_HLM_PM <- adonis2(SD_beta_wo_TLN_HLM_dist$Btotal ~ env[surveyed_sites_wo_TLN_HLM,19], permutations = 999))
(SD_beta_wo_TLN_HLM_PM_pair <- pairwise.adonis(SD_beta_wo_TLN_HLM_dist$Btotal, env[surveyed_sites_wo_TLN_HLM,19], p.adjust.m = "bonferroni", perm = 999))
# Mixed and stratified lakes
SD_beta_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], abund = F)
# Ocean sites and mixed lakes
SD_beta_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites,], abund = F)
# Stratified lakes and ocean sites
SD_beta_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites,], abund = F)
# Stratified lakes
SD_beta_S_dist<- BAT::beta(surveyed_sites_lake[stratified_lakes,], abund = F)


### Environmental
# Surveyed sites
SD_beta_env_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites_env,], abund = F)
# Mixed and stratified lakes
SD_beta_env_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], abund = F)
# Ocean sites and mixed lakes
SD_beta_env_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites_env,], abund = F)
# Stratified lakes and ocean sites
SD_beta_env_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites_env,], abund = F)
# Mixed lakes
SD_beta_env_M_dist<- BAT::beta(surveyed_sites_lake[mixed_lakes,], abund = F)


### Geographic
# Surveyed sites
SD_beta_geo_dist <- BAT::beta(surveyed_sites_lake[surveyed_sites,], abund = F)
# Mixed and stratified lakes
SD_beta_geo_MS_dist<- BAT::beta(surveyed_sites_lake[mixed_stratified_lakes,], abund = F)
# Ocean sites and mixed lakes
SD_beta_geo_OM_dist<- BAT::beta(surveyed_sites_lake[ocean_mixed_sites,], abund = F)
# Stratified lakes and ocean sites
SD_beta_geo_SO_dist<- BAT::beta(surveyed_sites_lake[ocean_stratified_sites,], abund = F)
# Mixed lakes
SD_beta_geo_M_dist<- BAT::beta(surveyed_sites_lake[mixed_lakes,], abund = F)
```

### SD beta q values of dispersion statistics
```{r, SD beta dispersion q-values}
# Sites
N <- 22
# Categories
k <- 3
# Sum of squares
SS <- SD_beta_AOV$`Sum Sq`[2]
# Mean of squares
MS <- SD_beta_AOV$`Mean Sq`[2]
# q value
q.value <- qtukey(p = 0.95, nmeans = k, df = N - k)
# Balanced
BSE <- sqrt(MS / 8)
# Unbalanced
USE <- sqrt((MS/2)*((1/6)+(1/8)))

group <- SD_beta_THSD$group

(OM_q <- (abs(group[1,1]))/USE)
(MS_q <- (abs(group[2,1]))/BSE)
(SO_q <- (abs(group[3,1]))/USE)
# Check values here https://www.socscistatistics.com/pvalues/qdistribution.aspx
```

### SD beta richness dispersions
- The dissimilarity between sites of the same site type
```{r, SD beta dispersion boxplot}
# Dispersion within-groups
SD_beta_BD_dist <- SD_beta_BD$distances
SD_beta_BD_dist <- as.data.frame(SD_beta_BD_dist)
SD_beta_BD_dist$X <- row.names(SD_beta_BD_dist)
SD_beta_BD_dist_env <- merge(SD_beta_BD_dist, env[surveyed_sites,], by = "X", sort = F)

SD_beta_BD_dist_env$Stratification <- factor(SD_beta_BD_dist_env$Stratification, levels = c("Ocean", "Mixed", "Stratified"))

(SD_beta_BD_dist_env_plot <- ggplot(SD_beta_BD_dist_env, aes(x = Stratification, y = SD_beta_BD_dist, color = Stratification, fill = Stratification)) +
  geom_violin(alpha = 0.6, draw_quantiles = c(0.25, 0.5, 0.75), aes(color = Stratification, fill = Stratification)) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  guides(fill = "none", color = "none") +
  geom_jitter(shape = 21,
            size = 4,
            alpha = 0.8,
            width = 0.1) +
    geom_text_repel(data = SD_beta_BD_dist_env, label = SD_beta_BD_dist_env$X, size = 5, point.padding = 7, max.overlaps = 30, nudge_x = -0.01) +
  theme_bw() +
  theme(text = element_text(size = 16), legend.text = element_text(size = 16),
    panel.background = element_rect(fill='transparent'),
    plot.background = element_rect(fill='transparent', color=NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(), axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(color = "black")) +
  xlab("Site type") +
  ylab("Distance to Centroid") +
  labs(color = "Stratification", tag = "a"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_BD_dist_plot.jpg", SD_beta_BD_dist_env_plot, width = 4.88, height = 6, units = "in")
```

### SD beta NMDS
```{r, SD beta NMDS}
### Regular
# Reference
SD_beta_ref_NMDS <- metaMDS(SD_beta_ref_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Surveyed sites 
SD_beta_NMDS <- metaMDS(SD_beta_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
round(SD_beta_NMDS$stress, digits = 2)
SD_beta_rep_NMDS <- metaMDS(SD_beta_dist$Brepl, try = 1000, parallel = 4, trymax = 500, maxit = 500)
SD_beta_ric_NMDS <- metaMDS(SD_beta_dist$Brich, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
SD_beta_MS_NMDS <- metaMDS(SD_beta_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
SD_beta_OM_NMDS <- metaMDS(SD_beta_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
SD_beta_SO_NMDS <- metaMDS(SD_beta_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes
SD_beta_S_NMDS <- metaMDS(SD_beta_S_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)


### Environmental
# Surveyed sites 
SD_beta_env_NMDS <- metaMDS(SD_beta_env_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
SD_beta_env_MS_NMDS <- metaMDS(SD_beta_env_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
SD_beta_env_OM_NMDS <- metaMDS(SD_beta_env_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
SD_beta_env_SO_NMDS <- metaMDS(SD_beta_env_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed lakes
SD_beta_env_M_NMDS <- metaMDS(SD_beta_env_M_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)


### Geographic
# Surveyed sites 
SD_beta_geo_NMDS <- metaMDS(SD_beta_geo_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed and stratified lakes
SD_beta_geo_MS_NMDS <- metaMDS(SD_beta_geo_MS_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Ocean sites and mixed lakes
SD_beta_geo_OM_NMDS <- metaMDS(SD_beta_geo_OM_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Stratified lakes and ocean sites
SD_beta_geo_SO_NMDS <- metaMDS(SD_beta_geo_SO_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
# Mixed lakes
SD_beta_geo_M_NMDS <- metaMDS(SD_beta_geo_M_dist$Btotal, try = 1000, parallel = 4, trymax = 500, maxit = 500)
```

### Load an envfit analysis function for adjusting p-values
- Need to run this function for downstream analyses in this Markdown file
```{r, load envfit p-value function}
# Function p.adjust.envfit
# Calculates adjusted P values for results stored in envfit object,
# created using envfit function from vegan, which fits supplementary variables
# onto axes of unconstrained ordination. 
# Arguments: 
# x - envfit object
# method - method for correction of multiple testing issue (default = 'bonferroni',
#          see ''?p.adjust' for more options)
# n - optional, number of tests for which to correct; if not given, the number is
#          taken as the number of tests conducted by envfit function (for both vectors and factors).
# Author: David Zeleny
p.adjust.envfit <- function (x, method = 'bonferroni', n)
{
  x.new <- x
  if (!is.null (x$vectors)) pval.vectors <- x$vectors$pvals else pval.vectors <- NULL
  if (!is.null (x$factors)) pval.factors <- x$factors$pvals else pval.factors <- NULL
  if (missing (n)) n <- length (pval.vectors) + length (pval.factors)
  if (!is.null (x$vectors)) x.new$vectors$pvals <- p.adjust (x$vectors$pvals, method = method, n = n)
  if (!is.null (x$factors)) x.new$factors$pvals <- p.adjust (x$factors$pvals, method = method, n = n)
  cat ('Adjustment of significance by', method, 'method')
  return (x.new)
}
```

### SD beta env and geo correlated variables using envfit
```{r, SD beta envfit}
### Environmental
# Surveyed sites 
# For figure
(SD_beta_env_ef <- envfit(SD_beta_env_NMDS, env[surveyed_sites_env,c(34)], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,19]))
(SD_beta_env_efp <- p.adjust.envfit(SD_beta_env_ef))
# Surveyed sites by stratification
(SD_beta_env_A_ef <- envfit(SD_beta_env_NMDS, env[surveyed_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,19]))
(SD_beta_env_A_efp <- p.adjust.envfit(SD_beta_env_A_ef))
# Mixed and stratified lakes
(SD_beta_env_MS_ef <- envfit(SD_beta_env_MS_NMDS, env[mixed_stratified_lakes,environment], permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
(SD_beta_env_MS_efp <- p.adjust.envfit(SD_beta_env_MS_ef))
# Ocean sites and mixed lakes
(SD_beta_env_OM_ef <- envfit(SD_beta_env_OM_NMDS, env[ocean_mixed_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,19]))
(SD_beta_env_OM_efp <- p.adjust.envfit(SD_beta_env_OM_ef))
# Stratified lakes and ocean sites
(SD_beta_env_SO_ef <- envfit(SD_beta_env_SO_NMDS, env[ocean_stratified_sites_env,environment], permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,19]))
(SD_beta_env_SO_efp <- p.adjust.envfit(SD_beta_env_SO_ef))
# Mixed lakes
(SD_beta_env_M_ef <- envfit(SD_beta_env_M_NMDS, env[mixed_lakes,environment], permutations = 999, na.rm = TRUE))
(SD_beta_env_M_efp <- p.adjust.envfit(SD_beta_env_M_ef))


# Stratified lakes
(SD_beta_S_ef <- envfit(SD_beta_S_NMDS, env[stratified_lakes,c(2,6,8,26,31:32)], permutations = 999, na.rm = TRUE))
(SD_beta_S_efp <- p.adjust.envfit(SD_beta_S_ef))


### Geographic
# Surveyed sites
# For figure
(SD_beta_geo_ef <- envfit(SD_beta_geo_NMDS, env[surveyed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,19]))
(SD_beta_geo_efp <- p.adjust.envfit(SD_beta_geo_ef))
# Surveyed sites by stratification
(SD_beta_geo_A_ef <- envfit(SD_beta_geo_NMDS, env[surveyed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,19]))
(SD_beta_geo_A_efp <- p.adjust.envfit(SD_beta_geo_A_ef))
# Mixed and stratified lakes
(SD_beta_geo_MS_ef <- envfit(SD_beta_geo_MS_NMDS, env[mixed_stratified_lakes,geography], permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
(SD_beta_geo_MS_efp <- p.adjust.envfit(SD_beta_geo_MS_ef))
# Ocean sites and mixed lakes
(SD_beta_geo_OM_ef <- envfit(SD_beta_geo_OM_NMDS, env[ocean_mixed_sites,geography], permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,19]))
(SD_beta_geo_OM_efp <- p.adjust.envfit(SD_beta_geo_OM_ef))
# Stratified lakes and ocean sites
(SD_beta_geo_SO_ef <- envfit(SD_beta_geo_SO_NMDS, env[ocean_stratified_sites,geography], permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,19]))
(SD_beta_geo_SO_efp <- p.adjust.envfit(SD_beta_geo_SO_ef))
# Mixed lakes
(SD_beta_geo_M_ef <- envfit(SD_beta_geo_M_NMDS, env[mixed_lakes,geography], permutations = 999, na.rm = TRUE))
(SD_beta_geo_M_efp <- p.adjust.envfit(SD_beta_geo_M_ef))
```

### SD beta Mantel correlation tests
```{r, SD beta Mantel tests}
### Environmental
# Surveyed sites
env_dist_t <- dist(scaled_env[surveyed_sites_env,c(1)], method = "euclidean")
(SD_beta_env_mant_t <- mantel(SD_beta_env_dist$Btotal, env_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,19]))
env_dist_s <- dist(scaled_env[surveyed_sites_env,c(2)], method = "euclidean")
(SD_beta_env_mant_s <- mantel(SD_beta_env_dist$Btotal, env_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,19]))
env_dist_o <- dist(scaled_env[surveyed_sites_env,c(3)], method = "euclidean")
(SD_beta_env_mant_o <- mantel(SD_beta_env_dist$Btotal, env_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites_env,19]))
# Adjust p-values
SD_beta_env_mant_pv <- rbind(SD_beta_env_mant_t$signif, SD_beta_env_mant_s$signif, SD_beta_env_mant_o$signif)
SD_beta_env_mant_pv <- SD_beta_env_mant_pv[,1]
(SD_beta_env_mant_pv <- p.adjust(SD_beta_env_mant_pv, method = "bonferroni"))

# Mixed and stratified lakes
env_MS_dist_t <- dist(scaled_env[mixed_stratified_lakes,c(1)], method = "euclidean")
(SD_beta_env_MS_mant_t <- mantel(SD_beta_env_MS_dist$Btotal, env_MS_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
env_MS_dist_s <- dist(scaled_env[mixed_stratified_lakes,c(2)], method = "euclidean")
(SD_beta_env_MS_mant_s <- mantel(SD_beta_env_MS_dist$Btotal, env_MS_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
env_MS_dist_o <- dist(scaled_env[mixed_stratified_lakes,c(3)], method = "euclidean")
(SD_beta_env_MS_mant_o <- mantel(SD_beta_env_MS_dist$Btotal, env_MS_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
# Adjust p-values
SD_beta_env_MS_mant_pv <- rbind(SD_beta_env_MS_mant_t$signif, SD_beta_env_MS_mant_s$signif, SD_beta_env_MS_mant_o$signif)
SD_beta_env_MS_mant_pv <- SD_beta_env_MS_mant_pv[,1]
(SD_beta_env_MS_mant_pv <- p.adjust(SD_beta_env_MS_mant_pv, method = "bonferroni"))

# Ocean sites and mixed lakes
env_OM_dist_t <- dist(scaled_env[ocean_mixed_sites_env,c(1)], method = "euclidean")
(SD_beta_env_OM_mant_t <- mantel(SD_beta_env_OM_dist$Btotal, env_OM_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,19]))
env_OM_dist_s <- dist(scaled_env[ocean_mixed_sites_env,c(2)], method = "euclidean")
(SD_beta_env_OM_mant_s <- mantel(SD_beta_env_OM_dist$Btotal, env_OM_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,19]))
env_OM_dist_o <- dist(scaled_env[ocean_mixed_sites_env,c(3)], method = "euclidean")
(SD_beta_env_OM_mant_o <- mantel(SD_beta_env_OM_dist$Btotal, env_OM_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites_env,19]))
# Adjust p-values
SD_beta_env_OM_mant_pv <- rbind(SD_beta_env_OM_mant_t$signif, SD_beta_env_OM_mant_s$signif, SD_beta_env_OM_mant_o$signif)
SD_beta_env_OM_mant_pv <- SD_beta_env_OM_mant_pv[,1]
(SD_beta_env_OM_mant_pv <- p.adjust(SD_beta_env_OM_mant_pv, method = "bonferroni"))

# Stratified lakes and ocean sites
env_SO_dist_t <- dist(scaled_env[ocean_stratified_sites_env,c(1)], method = "euclidean")
(SD_beta_env_SO_mant_t <- mantel(SD_beta_env_SO_dist$Btotal, env_SO_dist_t, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,19]))
env_SO_dist_s <- dist(scaled_env[ocean_stratified_sites_env,c(2)], method = "euclidean")
(SD_beta_env_SO_mant_s <- mantel(SD_beta_env_SO_dist$Btotal, env_SO_dist_s, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,19]))
env_SO_dist_o <- dist(scaled_env[ocean_stratified_sites_env,c(3)], method = "euclidean")
(SD_beta_env_SO_mant_o <- mantel(SD_beta_env_SO_dist$Btotal, env_SO_dist_o, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites_env,19]))
# Adjust p-values
SD_beta_env_SO_mant_pv <- rbind(SD_beta_env_SO_mant_t$signif, SD_beta_env_SO_mant_s$signif, SD_beta_env_SO_mant_o$signif)
SD_beta_env_SO_mant_pv <- SD_beta_env_SO_mant_pv[,1]
(SD_beta_env_SO_mant_pv <- p.adjust(SD_beta_env_SO_mant_pv, method = "bonferroni"))

# Mixed lakes
env_M_dist_t <- dist(scaled_env[mixed_lakes,c(1)], method = "euclidean")
(SD_beta_env_M_mant_t <- mantel(SD_beta_env_M_dist$Btotal, env_M_dist_t, method = "spearman", permutations = 999, na.rm = TRUE))
env_M_dist_s <- dist(scaled_env[mixed_lakes,c(2)], method = "euclidean")
(SD_beta_env_M_mant_s <- mantel(SD_beta_env_M_dist$Btotal, env_M_dist_s, method = "spearman", permutations = 999, na.rm = TRUE))
env_M_dist_o <- dist(scaled_env[mixed_lakes,c(3)], method = "euclidean")
(SD_beta_env_M_mant_o <- mantel(SD_beta_env_M_dist$Btotal, env_M_dist_o, method = "spearman", permutations = 999, na.rm = TRUE))
# Adjust p-values
SD_beta_env_M_mant_pv <- rbind(SD_beta_env_M_mant_t$signif, SD_beta_env_M_mant_s$signif, SD_beta_env_M_mant_o$signif)
SD_beta_env_M_mant_pv <- SD_beta_env_M_mant_pv[,1]
(SD_beta_env_M_mant_pv <- p.adjust(SD_beta_env_M_mant_pv, method = "bonferroni"))


# Stratified lakes
env_geo_S_dist <- dist(scaled_env[stratified_lakes,c(1:3,9,14:15)], method = "euclidean")
(SD_beta_S_mant <- mantel(SD_beta_S_dist$Btotal, env_geo_S_dist, method = "spearman", permutations = 999, na.rm = TRUE))


### Geographic
# Surveyed sites
geo_dist_dmean <- dist(scaled_env[surveyed_sites,c(9)], method = "euclidean")
(SD_beta_geo_mant_dmean <- mantel(SD_beta_geo_dist$Btotal, geo_dist_dmean, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,19]))
geo_dist_md <- dist(scaled_env[surveyed_sites,c(14)], method = "euclidean")
(SD_beta_geo_mant_md <- mantel(SD_beta_geo_dist$Btotal, geo_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,19]))
geo_dist_la <- dist(scaled_env[surveyed_sites,c(15)], method = "euclidean")
(SD_beta_geo_mant_la <- mantel(SD_beta_geo_dist$Btotal, geo_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[surveyed_sites,19]))
# Adjust p-values
SD_beta_geo_mant_pv <- rbind(SD_beta_geo_mant_dmean$signif, SD_beta_geo_mant_md$signif, SD_beta_geo_mant_la$signif)
SD_beta_geo_mant_pv <- SD_beta_geo_mant_pv[,1]
(SD_beta_geo_mant_pv <- p.adjust(SD_beta_geo_mant_pv, method = "bonferroni"))

# Mixed and stratified lakes 
geo_MS_dist_dmean <- dist(scaled_env[mixed_stratified_lakes,c(9)], method = "euclidean")
(SD_beta_geo_MS_mant_dmean <- mantel(SD_beta_geo_MS_dist$Btotal, geo_MS_dist_dmean, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
geo_MS_dist_md <- dist(scaled_env[mixed_stratified_lakes,c(14)], method = "euclidean")
(SD_beta_geo_MS_mant_md <- mantel(SD_beta_geo_MS_dist$Btotal, geo_MS_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
geo_MS_dist_la <- dist(scaled_env[mixed_stratified_lakes,c(15)], method = "euclidean")
(SD_beta_geo_MS_mant_la <- mantel(SD_beta_geo_MS_dist$Btotal, geo_MS_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[mixed_stratified_lakes,19]))
# Adjust p-values
SD_beta_geo_MS_mant_pv <- rbind(SD_beta_geo_MS_mant_dmean$signif, SD_beta_geo_MS_mant_md$signif, SD_beta_geo_MS_mant_la$signif)
SD_beta_geo_MS_mant_pv <- SD_beta_geo_MS_mant_pv[,1]
(SD_beta_geo_MS_mant_pv <- p.adjust(SD_beta_geo_MS_mant_pv, method = "bonferroni"))

# Ocean sites and mixed lakes
geo_OM_dist_dmean <- dist(scaled_env[ocean_mixed_sites,c(9)], method = "euclidean")
(SD_beta_geo_OM_mant_dmean <- mantel(SD_beta_geo_OM_dist$Btotal, geo_OM_dist_dmean, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,19]))
geo_OM_dist_md <- dist(scaled_env[ocean_mixed_sites,c(14)], method = "euclidean")
(SD_beta_geo_OM_mant_md <- mantel(SD_beta_geo_OM_dist$Btotal, geo_OM_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,19]))
geo_OM_dist_la <- dist(scaled_env[ocean_mixed_sites,c(15)], method = "euclidean")
(SD_beta_geo_OM_mant_la <- mantel(SD_beta_geo_OM_dist$Btotal, geo_OM_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_mixed_sites,19]))
# Adjust p-values
SD_beta_geo_OM_mant_pv <- rbind( SD_beta_geo_OM_mant_dmean$signif, SD_beta_geo_OM_mant_md$signif, SD_beta_geo_OM_mant_la$signif)
SD_beta_geo_OM_mant_pv <- SD_beta_geo_OM_mant_pv[,1]
(SD_beta_geo_OM_mant_pv <- p.adjust(SD_beta_geo_OM_mant_pv, method = "bonferroni"))

# Stratified lakes and ocean sites
geo_SO_dist_dmean <- dist(scaled_env[ocean_stratified_sites,c(9)], method = "euclidean")
(SD_beta_geo_SO_mant_dmean <- mantel(SD_beta_geo_SO_dist$Btotal, geo_SO_dist_dmean, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,19]))
geo_SO_dist_md <- dist(scaled_env[ocean_stratified_sites,c(14)], method = "euclidean")
(SD_beta_geo_SO_mant_md <- mantel(SD_beta_geo_SO_dist$Btotal, geo_SO_dist_md, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,19]))
geo_SO_dist_la <- dist(scaled_env[ocean_stratified_sites,c(15)], method = "euclidean")
(SD_beta_geo_SO_mant_la <- mantel(SD_beta_geo_SO_dist$Btotal, geo_SO_dist_la, method = "spearman", permutations = 999, na.rm = TRUE, strata = env[ocean_stratified_sites,19]))
# Adjust p-values
SD_beta_geo_SO_mant_pv <- rbind( SD_beta_geo_SO_mant_dmean$signif, SD_beta_geo_SO_mant_md$signif, SD_beta_geo_SO_mant_la$signif)
SD_beta_geo_SO_mant_pv <- SD_beta_geo_SO_mant_pv[,1]
(SD_beta_geo_SO_mant_pv <- p.adjust(SD_beta_geo_SO_mant_pv, method = "bonferroni"))

# Mixed lakes
geo_M_dist_dmean <- dist(scaled_env[mixed_lakes,c(9)], method = "euclidean")
(SD_beta_geo_M_mant_dmean <- mantel(SD_beta_geo_M_dist$Btotal, geo_M_dist_dmean, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_md <- dist(scaled_env[mixed_lakes,c(14)], method = "euclidean")
(SD_beta_geo_M_mant_md <- mantel(SD_beta_geo_M_dist$Btotal, geo_M_dist_md, method = "spearman", permutations = 999, na.rm = TRUE))
geo_M_dist_la <- dist(scaled_env[mixed_lakes,c(15)], method = "euclidean")
(SD_beta_geo_M_mant_la <- mantel(SD_beta_geo_M_dist$Btotal, geo_M_dist_la, method = "spearman", permutations = 999, na.rm = TRUE))
# Adjust p-values
SD_beta_geo_M_mant_pv <- rbind( SD_beta_geo_M_mant_dmean$signif, SD_beta_geo_M_mant_md$signif, SD_beta_geo_M_mant_la$signif)
SD_beta_geo_M_mant_pv <- SD_beta_geo_M_mant_pv[,1]
(SD_beta_geo_M_mant_pv <- p.adjust(SD_beta_geo_M_mant_pv, method = "bonferroni"))
```

### SD beta NMDS ordination plots
```{r, SD beta NMDS plots}
# SD beta total NMDS scores
SD_beta_NMDS_data.scores <- as.data.frame(scores(SD_beta_NMDS))
SD_beta_NMDS_data.scores$Stratification <- env[surveyed_sites,19]
SD_beta_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
SD_beta_NMDS_data.scores$Stratification <- factor(SD_beta_NMDS_data.scores$Stratification, levels = c("Ocean", "Mixed", "Stratified"))

# Get significantly correlated environmental variables
SD_beta_env_ef_coord_cont <- as.data.frame(scores(SD_beta_env_ef, "vectors")) * ordiArrowMul(SD_beta_env_ef)
SD_beta_env_row_names <- c("S")
# Assign the new row names to the data frame
SD_beta_env_ef_coord_cont <- data.frame(row.names = SD_beta_env_row_names, SD_beta_env_ef_coord_cont)

# SD_beta_geo_ef_coord_cont <- as.data.frame(scores(SD_beta_geo_ef, "vectors")) * ordiArrowMul(SD_beta_geo_ef)
# SD_beta_geo_row_names <- c("minD")
# # Assign the new row names to the data frame
# SD_beta_geo_ef_coord_cont <- data.frame(row.names = SD_beta_geo_row_names, SD_beta_geo_ef_coord_cont)

# Plot NMDS ordination of SD beta total with CI = 0.95
SD_beta_ef_plot <- ggplot(data = SD_beta_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Stratification)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Stratification), type = "t", level = 0.95) +
  geom_point(data = SD_beta_NMDS_data.scores, aes(color = Stratification, fill = Stratification), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               data = SD_beta_env_ef_coord_cont, linewidth =1, alpha = 0.2, color = env_cont) +
  geom_text(data = SD_beta_env_ef_coord_cont, aes(x = NMDS1, y = NMDS2),
            color = env_cont, label = row.names(SD_beta_env_ef_coord_cont), size = 5) +
  geom_text_repel(data = SD_beta_NMDS_data.scores, label = SD_beta_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 16), 
        legend.position = "bottom", 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black", size = 16), 
        legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
           label = paste("Stress: ", round(SD_beta_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(SD_beta_ef_plot <- SD_beta_ef_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_NMDS.jpg", SD_beta_ef_plot, width = 6, height = 6, units = "in")


# Plot NMDS ordination of SD beta total with CI = 0.90
SD_beta_ef_plot_CI90 <- ggplot(data = SD_beta_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Stratification)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Stratification), type = "t", level = 0.90) +
  geom_point(data = SD_beta_NMDS_data.scores, aes(color = Stratification, fill = Stratification), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               data = SD_beta_env_ef_coord_cont, linewidth =1, alpha = 0.2, color = env_cont) +
  geom_text(data = SD_beta_env_ef_coord_cont, aes(x = NMDS1, y = NMDS2),
            color = env_cont, label = row.names(SD_beta_env_ef_coord_cont), size = 5) +
  geom_text_repel(data = SD_beta_NMDS_data.scores, label = SD_beta_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 16), 
        legend.position = "bottom", 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 16), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black"), 
        legend.key = element_blank()) +
  annotate("text", x = -0.6, y = 0.5, size = 5,
           label = paste("Stress: ", round(SD_beta_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-0.81,0.81))
(SD_beta_ef_plot_CI90 <- SD_beta_ef_plot_CI90 + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_NMDS_CI90.jpg", SD_beta_ef_plot_CI90, width = 6, height = 6, units = "in")


# SD beta replacement NMDS scores
SD_beta_rep_NMDS_data.scores <- as.data.frame(scores(SD_beta_rep_NMDS))
SD_beta_rep_NMDS_data.scores$Stratification <- env[surveyed_sites,19]
SD_beta_rep_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
SD_beta_rep_NMDS_data.scores$Stratification <- factor(SD_beta_rep_NMDS_data.scores$Stratification, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta replacement with CI = 0.95
SD_beta_rep_plot <- ggplot(data = SD_beta_rep_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Stratification)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Stratification), type = "t", level = 0.95) +
  geom_point(data = SD_beta_rep_NMDS_data.scores, aes(color = Stratification, fill = Stratification), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_rep_NMDS_data.scores, label = SD_beta_rep_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black"), 
        legend.key = element_blank()) +
  annotate("text", x = -0.7, y = 0.8, size = 5,
           label = paste("Stress: ", round(SD_beta_rep_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(SD_beta_rep_plot <- SD_beta_rep_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_rep_NMDS.jpg", SD_beta_rep_plot, width = 4.88, height = 6, units = "in")


# SD beta richness NMDS scores
SD_beta_ric_NMDS_data.scores <- as.data.frame(scores(SD_beta_ric_NMDS))
SD_beta_ric_NMDS_data.scores$Stratification <- env[surveyed_sites,19]
SD_beta_ric_NMDS_data.scores$Lakes <- env[surveyed_sites,1]
SD_beta_ric_NMDS_data.scores$Stratification <- factor(SD_beta_ric_NMDS_data.scores$Stratification, levels = c("Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta richness with CI = 0.95
SD_beta_ric_plot <- ggplot(data = SD_beta_ric_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Stratification)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Stratification), type = "t", level = 0.95) +
  geom_point(data = SD_beta_ric_NMDS_data.scores, aes(color = Stratification, fill = Stratification), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_ric_NMDS_data.scores, label = SD_beta_ric_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black"), 
        legend.key = element_blank()) +
  annotate("text", x = -0.7, y = 0.8, size = 5,
           label = paste("Stress: ", round(SD_beta_ric_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "b") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(SD_beta_ric_plot <- SD_beta_ric_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_ric_NMDS.jpg", SD_beta_ric_plot, width = 4.88, height = 6, units = "in")


# SD beta ref NMDS scores
SD_beta_ref_NMDS_data.scores <- as.data.frame(scores(SD_beta_ref_NMDS))
SD_beta_ref_NMDS_data.scores$Stratification <- env[,19]
SD_beta_ref_NMDS_data.scores$Lakes <- env[,1]
SD_beta_ref_NMDS_data.scores$Stratification <- factor(SD_beta_ref_NMDS_data.scores$Stratification, levels = c("Reference", "Ocean", "Mixed", "Stratified"))

# Plot NMDS ordination of SD beta ref with CI = 0.95
SD_beta_ref_plot <- ggplot(data = SD_beta_ref_NMDS_data.scores, aes(x = NMDS1, y = NMDS2, color = Stratification)) + 
  stat_ellipse(aes(x = NMDS1, y = NMDS2, color = Stratification), type = "t", level = 0.95) +
  geom_point(data = SD_beta_ref_NMDS_data.scores, aes(color = Stratification, fill = Stratification), size = 4, alpha = 1) + 
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  geom_text_repel(data = SD_beta_ref_NMDS_data.scores, label = SD_beta_ref_NMDS_data.scores$Lakes, 
                  size = 5, force = 10, point.padding = 0, max.overlaps = 30, nudge_x = -0.05) +
  theme(text = element_text(size = 14), 
        legend.position = "bottom", 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 14), 
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.border = element_rect(fill = NA), 
        axis.text = element_text(color = "black"), 
        legend.key = element_blank()) +
  annotate("text", x = -0.7, y = 0.8, size = 5,
           label = paste("Stress: ", round(SD_beta_ref_NMDS$stress, digits = 2))) +
  labs(colour = "Site type:  ", fill = "Site type:  ", tag = "a") +
  coord_fixed() +
  scale_x_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05)) +
  scale_y_continuous(breaks = c(-0.5,0,0.5), limits = c(-1.05,1.05))
(SD_beta_ref_plot <- SD_beta_ref_plot + guides(color = guide_legend(override.aes = list(label = ""))))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_ref_NMDS.jpg", SD_beta_ref_plot, width = 4.88, height = 6, units = "in")
```

### SD beta dendrogram
```{r, SD beta dendrogram}
# Cluster communities using average-linkage algorithm
SD_beta_dist_clust <- hclust(SD_beta_dist$Btotal, method = "average")

# The following will create a dendrogram of SD_beta putting things in one dimensional space
# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_dendrogram.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

# Create your plot using the plot() function
plot(SD_beta_dist_clust, 
     xlab = "Sites",
     ylab = "SD beta",
     main = "",
     sub = "",
     pch = 20,
     col= "black")

# Close the jpg device
dev.off()


# The following will graph and identify the best supported number of clusters of the dendrogram
# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_dg_silhouette.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

Si <- numeric(nrow(presabs_lake[surveyed_sites,]))
for (k in 2:(nrow(presabs_lake[surveyed_sites,])-1))
{
  sil<-silhouette(cutree(SD_beta_dist_clust, k=k), SD_beta_dist$Btotal)
  Si[k] <- summary(sil)$avg.width
}
k.best<-which.max(Si)
plot(1:nrow(presabs_lake[surveyed_sites,]), Si, type = "h", main = "Silhouette", xlab = "K", ylab = "Width")
axis(1, k.best, paste("optimum", k.best, sep="\n"), col = "red", font = 2, col.axis = "red")

# Close the jpg device
dev.off()


# Partitioning around mediods uses distances matrix to found how points according to silhouette are grouped around a centroid. The greater the number, the greater the distance from the centroid.
# Open a jpg device
png("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/SD_beta_dg_partitioning.jpg", width = 6.5, height = 3, units = "in", res = 300, type = "cairo")

SD_beta_pam <- pam(SD_beta_dist$Btotal,2)
plot(SD_beta_pam)

# Close the jpg device
dev.off()
```

### SD beta outliers
```{r, SD beta outliers}
# Determine NMDS1 outliers
ordered_SD_beta_NMDS1_data.scores <- SD_beta_NMDS_data.scores[order(SD_beta_NMDS_data.scores$NMDS1), ]

outlier_SD_beta_NMDS1 <- ordered_SD_beta_NMDS1_data.scores %>%
  group_by(Stratification) %>%
  mutate(
    Q1 = quantile(NMDS1, 0.25),
    Q3 = quantile(NMDS1, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = NMDS1 < lower_bound | NMDS1 > upper_bound
  )

outlier_SD_beta_NMDS1 <- as.data.frame(outlier_SD_beta_NMDS1)

row.names(outlier_SD_beta_NMDS1) <- outlier_SD_beta_NMDS1$X

# Create the plot
(outlier_SD_beta_NMDS1_plot <- ggplot(outlier_SD_beta_NMDS1, aes(x = Stratification, y = NMDS1, fill = Stratification)) +
  geom_violin(trim = FALSE) +
  geom_jitter(aes(color = is_outlier), width = 0.2, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  theme(text = element_text(size = 16),
        legend.position = "right",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  labs(y = "NMDS1 Distances", x = "Site type", color = "Outlier", fill = "Site type", tag = "a"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/outlier_SD_beta_NMDS1.jpg", outlier_SD_beta_NMDS1_plot, width = 6.26, height = 6, units = "in")


# Determine NMDS2 outliers
ordered_SD_beta_NMDS2_data.scores <- SD_beta_NMDS_data.scores[order(SD_beta_NMDS_data.scores$NMDS2), ]

outlier_SD_beta_NMDS2 <- ordered_SD_beta_NMDS2_data.scores %>%
  group_by(Stratification) %>%
  mutate(
    Q1 = quantile(NMDS2, 0.25),
    Q3 = quantile(NMDS2, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = NMDS2 < lower_bound | NMDS2 > upper_bound
  )

outlier_SD_beta_NMDS2 <- as.data.frame(outlier_SD_beta_NMDS2)

row.names(outlier_SD_beta_NMDS2) <- outlier_SD_beta_NMDS2$X

# Create the plot
(outlier_SD_beta_NMDS2_plot <- ggplot(outlier_SD_beta_NMDS2, aes(x = Stratification, y = NMDS2, fill = Stratification)) +
  geom_violin(trim = FALSE) +
  geom_jitter(aes(color = is_outlier), width = 0.2, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  theme(text = element_text(size = 16),
        legend.position = "right",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  labs(y = "NMDS2 Distances", x = "Site type", color = "Outlier", fill = "Site type", tag = "b"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/outlier_SD_beta_NMDS2.jpg", outlier_SD_beta_NMDS2_plot, width = 6.26, height = 6, units = "in")


# Make SD_beta distances into a matrix
SD_beta_dist_matrix <- as.matrix(SD_beta_dist$Btotal)

# Set the diagonal elements to NA
diag(SD_beta_dist_matrix) <- NA

# Get the labels of the sites
sites <- attr(SD_beta_dist_matrix, "Labels")

# Calculate the mean dist for each group
SD_beta_mean_dist <- aggregate(SD_beta_dist_matrix, by = list(stratification_group), FUN = mean, na.rm = TRUE)

# Rename rows, delete column, and transpose matrix and make it a data frame
row.names(SD_beta_mean_dist) <- SD_beta_mean_dist$Group.1
SD_beta_mean_dist <- SD_beta_mean_dist[,-1] 
SD_beta_mean_dist <- as.data.frame(t(SD_beta_mean_dist))

# Add Stratification column
SD_beta_mean_dist$Stratification <- env[surveyed_sites,19]
SD_beta_mean_dist$Stratification <- factor(SD_beta_mean_dist$Stratification, levels = c("Ocean", "Mixed", "Stratified"))


# Determine ocean sites outliers based on distance from other ocean sites
outlier_SD_beta_mean_dist_ocean <- SD_beta_mean_dist[ocean_sites,] %>%
  group_by(Stratification) %>%
  mutate(
    Q1 = quantile(Ocean, 0.25),
    Q3 = quantile(Ocean, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = Ocean < lower_bound | Ocean > upper_bound
  )

outlier_SD_beta_mean_dist_ocean <- as.data.frame(outlier_SD_beta_mean_dist_ocean)
row.names(outlier_SD_beta_mean_dist_ocean) <- outlier_SD_beta_mean_dist_ocean$X
outlier_SD_beta_mean_dist_ocean$Distances <- outlier_SD_beta_mean_dist_ocean$Ocean
outlier_SD_beta_mean_dist_ocean <- outlier_SD_beta_mean_dist_ocean[,-c(1:3)]


# Determine mixed lake outliers based on distance from other mixed lakes
outlier_SD_beta_mean_dist_mixed <- SD_beta_mean_dist[mixed_lakes,] %>%
  group_by(Stratification) %>%
  mutate(
    Q1 = quantile(Mixed, 0.25),
    Q3 = quantile(Mixed, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = Mixed < lower_bound | Mixed > upper_bound
  )

outlier_SD_beta_mean_dist_mixed <- as.data.frame(outlier_SD_beta_mean_dist_mixed)
row.names(outlier_SD_beta_mean_dist_mixed) <- outlier_SD_beta_mean_dist_mixed$X
outlier_SD_beta_mean_dist_mixed$Distances <- outlier_SD_beta_mean_dist_mixed$Mixed
outlier_SD_beta_mean_dist_mixed <- outlier_SD_beta_mean_dist_mixed[,-c(1:3)]


# Determine stratified lake outliers based on distance from other stratified lakes
outlier_SD_beta_mean_dist_stratified <- SD_beta_mean_dist[stratified_lakes,] %>%
  group_by(Stratification) %>%
  mutate(
    Q1 = quantile(Stratified, 0.25),
    Q3 = quantile(Stratified, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR,
    is_outlier = Stratified < lower_bound | Stratified > upper_bound
  )

outlier_SD_beta_mean_dist_stratified <- as.data.frame(outlier_SD_beta_mean_dist_stratified)
row.names(outlier_SD_beta_mean_dist_stratified) <- outlier_SD_beta_mean_dist_stratified$X
outlier_SD_beta_mean_dist_stratified$Distances <- outlier_SD_beta_mean_dist_stratified$Stratified
outlier_SD_beta_mean_dist_stratified <- outlier_SD_beta_mean_dist_stratified[,-c(1:3)]

outlier_SD_beta_mean_dist <- rbind(outlier_SD_beta_mean_dist_ocean, outlier_SD_beta_mean_dist_mixed, outlier_SD_beta_mean_dist_stratified)

(outlier_SD_beta_mean_dist_plot <- ggplot(outlier_SD_beta_mean_dist, aes(x = Stratification, y = Distances, fill = Stratification)) +
  geom_violin(trim = FALSE) +
  geom_jitter(aes(color = is_outlier), width = 0.2, size = 4, alpha = 0.7) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "purple")) +
  scale_fill_manual(values = custom_colors) +
  theme(text = element_text(size = 16),
        legend.position = "right",
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.line = element_line(color = "black"),
        panel.background = element_rect(fill='transparent'),
        plot.background = element_rect(fill='transparent', color=NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(color = "black", size = 16)) +
  labs(y = "Average Distance from other sites by site type", x = "Site type", color = "Outlier", fill = "Site type", tag = "c"))
ggsave("/Users/bailey/Documents/research/fish_biodiversity/figures/SD/outlier_SD_beta_mean_dist.jpg", outlier_SD_beta_mean_dist_plot, width = 6.26, height = 6, units = "in")
```

### Package and version info
```{r, session info}
sessionInfo()
```

#### Determine what families of fish are most present in the lakes
```{r, eval=F}
library(rfishbase)
species_sums_tax <- site_sums
species_sums_tax$X <- gsub("_", " ", species_sums_tax$X)
species_sums_tax <- species_sums_tax$X
species_sums_tax <- rfishbase::load_taxa() %>% 
  filter(Species %in% species_sums_tax) %>%
  collect()
species_sums_tax$Species <- gsub(" ", "_", species_sums_tax$Species)
species_sums_tax <- merge(species_sums_tax, species_sums, by.x = "Species", by.y = "X")
species_sums_tax <- species_sums_tax[, -c(2:4, 7:8)]
species_sums_tax <- species_sums_tax[order(species_sums_tax$row_sum),]

species_sums_tax_summary <- species_sums_tax %>%
  group_by(Family) %>%
  summarise(across(.cols = row_sum, .fns = sum, na.rm = TRUE))

family <- ggplot(data = species_sums_tax_summary) + 
  geom_bar(aes(x = reorder(Family, row_sum), y = row_sum),
    stat = 'identity',
    col="green", 
    fill="green4", 
    alpha = .5) + 
  theme_bw() +
  theme(plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), axis.line = element_line(color = "black")) +
  labs(x="Family", y="Family Richness")
family
ggsave("family.jpg", family, width = 5, height = 4.5)

order <- ggplot(data = species_sums_tax) + 
  geom_bar(aes(x = reorder(Order, row_sum), y = row_sum),
    stat = 'identity',
    col="green", 
    fill="green4", 
    alpha = .5) + 
  theme_bw() +
  theme(plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), axis.line = element_line(color = "black")) +
  labs(x="Order", y="Richness in Lakess")
order
#ggsave("order.jpg", order)

detach("package:rfishbase")
```
