---
title: "Fish incidence"
output: github_document
editor_options: 
  chunk_output_type: console
---
#### R Markdown

## Packages
- These are the packages that are needed to extract data from online repositories and manipulate the data
```{r load packages and files}
library(openxlsx) # data manipulation
library(rfishbase) # data manipulation
library(tidyr) # data manipulation
library(dplyr) # data manipulation
library(phytools) # data manipulation
library(taxize)
```

## Reference
- Take the Micronesian fish file and extract the Palau species from it
```{r, Reference survey}
# File from Rob containing the reference pool of all Micronesian fish with a column dedicated to the fish species in Palau.
micronesian_records <- read.xlsx("/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_surveys/micronesian_fish_corrected.xlsx", sheet = 1, startRow = 3, colNames = T, cols =  1:31)

# Extract the three columns of Palauan fish data with genus and species names and whether a fish is present or absent there
palau_fish <- micronesian_records[c(1:2282),c(5:6,8,24,29)]

# Some fish lack the information of whether they occur in Palau. Therefore we need to get rid of any fish species that we are not confident occur in Palau. A 1 means that the species occurs there and anything else is speculation currently.

# Convert all NAs (and a few remaining "x") to 0
palau_fish$X29[palau_fish$X29 == "x"] <- NA; palau_fish$X29[palau_fish$X29 == "o"] <- 1; palau_fish$X29[palau_fish$X29 == "1?"] <- 1; palau_fish$X29[palau_fish$X29 == "?"] <- NA; palau_fish$X29[palau_fish$X29 == "e"] <- NA; palau_fish$X29[palau_fish$X29 == "x?"] <- NA; palau_fish$X29[palau_fish$X29 == "2?"] <- 1; palau_fish$X29[palau_fish$X29 == " "] <- NA; palau_fish$X29[palau_fish$X29 == "3"] <- 1

# Convert occurrence column (apart from species_name) to numeric
palau_fish$X29 <- as.numeric(palau_fish$X29)

# Remove all species that were labeled with an NA which means currently they are not found in Palau.
fish <- subset(palau_fish, X29 > 0.5) 

## Remove species that live below 200m
# Paracaesio stonei
# Gymnothorax neglectus
# Create a logical condition to identify rows to remove
rows_to_remove <- fish$X8 %in% c("Paracaesio stonei", "Gymnothorax neglectus")

# Subset the data frame to exclude rows meeting the condition
fish <- fish[!rows_to_remove, ]

## Combine Genus and Species
# This will be the corrected combined Genus and Species column where as the other column X8 contains the uncorrected names from Rob's list
fish$Species <- paste(fish[, 1], fish[, 2])

## To preserve the original names Rob gave his species I kept the full Genus and species name uncorrected but corrected the Genus or species individually so I could compare the ID Rob gave versus the corrected name that FishBase gave his ID
# This will show original species names and those that have changed due to using FishBase command validate_names which will be executed later on down below
check <- setdiff(fish$Species, fish$X8)
check
# Here you can see that most names that were corrected were due to minimal spelling mistakes
# Currently at 179

check2 <- setdiff(fish$X8, fish$Species)
check2
# check2 includes species names that are obsolete and were changed to something else that was already in the list
compare_ref_species_names <- as.data.frame(cbind(check, check2))
# Will be a different length when compared to "check" if there are duplicates for species names that are now something else that previously existed
# The warning is because the vectors are different lengths
write.csv(compare_ref_species_names, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/compare_ref_species_names.csv")

# Change column name, it will duplicate the column with a new column name
fish$REF <- fish$X29

# Extract only the genus and species names now that we know that this file only contains species from Palau
fish <- fish[,c(6:7)]
```

## Surveys
- Use the community surveys and extract the species and incidence data
```{r, Site surveys}
# A file with the presence and absence of species observed in the marine lakes. Names have been edited to reflect spelling in FishBase.
palau_records <- read.xlsx("/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_surveys/palau_marine_lake_fish_biodiversity_ordered&corrected.xlsx", sheet = 1, startRow = 3, colNames = T)

# Extract the lake columns with species names. The lake columns have 3 letter codes and if a species is present there will be a 1 in the column.
site_fish <- palau_records[-c(254:259), c(2:3,6,11,16,21,26,31,37,43:44,53,58,63,68,73:74,79,84,89,94,101,106,111)]

# Combine Genus and Species
site_fish$Species <- paste(site_fish[, 1], site_fish[, 2])

site_fish[is.na(site_fish)] <- 0; site_fish[site_fish == "x"] <- 0; site_fish[site_fish == "?"] <- 0

# Convert all fields (apart from species_name) to numeric
# Avoid the species column when using this command
site_fish[, -c(1,2,25)] <- lapply(site_fish[, -c(1,2,25)], function(x) as.numeric(x))

# Identify what species are found in the lakes but not the reference pool
site_data_only <- setdiff(site_fish$Species, fish$Species)
site_data_only
# [1] "Acanthurus sp" "Atherinomorus sp" "Epinephelus sp" "Myripristis sp" "Parioglossus sp"
# [6] "Platax sp" "Pomacentrus sp 2" "Pomacentrus sp" "Silhouettea sp"
```

## Rid unwanted species
```{r, Check Between sites and ref}
# Bring together marine lake presence/absence surveys and palau survey
pres_abs <- merge(site_fish, fish, by = "Species", all = TRUE)

# Create the presence/absence dataframe
pres_abs <- pres_abs[, c(1, 4:26)]

# Convert all NAs (and a few remaining "x" and "?") to 0
pres_abs[is.na(pres_abs)] <- 0; pres_abs[pres_abs == "x"] <- 0; pres_abs[pres_abs == "?"] <- 0

# Convert all fields (apart from species_name) to numeric
# Avoid the species column when using this command
pres_abs[, -1] <- lapply(pres_abs[, -1], function(x) as.numeric(x))

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
pres_abs <- pres_abs[which(rowSums(pres_abs[, -1]) > 0),]

# Removes the species that are not present in any of the environments
pres_abs <- aggregate(pres_abs[,-1], list(names = pres_abs$Species), max, na.rm = TRUE)

pres_abs$REF[pres_abs$REF > "-1"] <- 1

# Extract only the species name
species_names <- pres_abs[, 1]
# 1724 fish species found in Palau and the lakes surveyed but 9 species from the site surveys were only identified to the genus
```

## FishBase validation
```{r, validate species names on rfishbase}
# Run to check for duplicates, you will delete these duplicates in the next line
duplicates_check <- which(duplicated(species_names))
duplicates_check 

## Numbers may change from the above code identifying duplicates, so update each time
# Remove duplicates if need be
#species_names <- unique(species_names)

## Validate the species names against FishBase database
# This file is used to revise the original files to prevent having to update species names every time
val_species_names <- validate_names(species_names)

## Get rid of any species that are not valid by FishBase
# Be careful because a species may be misspelled or not updated or added by FishBase leading to an NA result
val_species_names <- na.omit(val_species_names)
# 1674 of the 1724 species names are validated by FishBase

# For some reason these species are not validate in FishBase even though they show up there.
Ci <- "Cheilodipterus isostigma"
Cr <- "Cirrhilabrus ryukyuensis"
Tm <- "Trimma macrophthalmus"
Zl <- "Zoramia leptacanthus"

# Add this species to the validated group because you can still download its traits.
val_species_names <- append(val_species_names, values = c(Ci, Cr, Tm, Zl))
# Now we have 1678 validated

# For some reason FishBase adds Genus Species to some of the species that are not in its database when you validate species names
blanks <- setdiff(val_species_names, species_names)
blanks
word_to_delete <- "Genus Species"
val_species_names <- val_species_names[val_species_names != word_to_delete]
# This shows us that only 1675 species were validated

# Create a file of species that were only identified to genus level. They will be lost by the validate_names command above and will need to be added back in
incomplete_to_add <- setdiff(species_names, val_species_names)
incomplete_to_add
# 48 species were not validated by FishBase. Some have Genus and Species but have yet to be added to FishBase. Others with an sp cannot be validated
write.csv(incomplete_to_add, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/incomplete_to_add.csv")

# Add back in species that were lost by validate_names command
existing_species <- c(val_species_names, incomplete_to_add)

# Checking for duplicates and comparing files
duplicates_check <- which(duplicated(existing_species))
duplicates_check

# Check differences
diff_check <- setdiff(existing_species, species_names)
diff_check

# Resort your vector since we added some species/changed names. Will come in handy later mainly
existing_species <- sort(existing_species)

# Create these files for when revising the original file so that you have all the correct names and the NAs that now replace lost species
# Once you have loaded the packages and loaded the file containing all species from survey data, you can download trait data from fishbase using fish_traits.Rmd. #From trait-beta-div-processing-fish-traits.R

write.csv(species_names, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/species_names.csv")

write.csv(val_species_names, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/val_species_names.csv")

write.csv(existing_species, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/species_lists/existing_species.csv")
```

## Aesthetics
- Edit occurrence data from the locations that were sampled. Isolate just the occurrence data from everything else that is in the file. This is for future when we compare it with the environmental data. Take out species not found in any location and edit the file for formatting purposes. #From trait-beta-div-processing.R
```{r, Edit presence absence matrix}
# 1. BCM (S), 2. CLM (S), 3. FLK (M), 4. GLK (S), 5. HLM (S), 6. HLO (M), 7. IBK (O), 8. LLN (M), 9. LCN (O), 10. MLN (M), 11. NCN (O), 12. NLK (S), 13. NLN (M), 14. NLU (M), 15. OLO (M), 16. OOO (O), 17. OTM (S), 18. OOM (O), 19. RCA (O), 20. REF (R), 21. SLN (S), 22. TLN (S), 23. ULN (M)

# Sort names
pres_abs$names <- sort(pres_abs$names)

# This should hopefully return a file without duplicates unless things are named differently
pres_abs <- pres_abs[match(existing_species, pres_abs$names),]

# Replace the space in the species name with an underscore
pres_abs$names <- gsub(" ", "_", pres_abs$names)

duplicates_check <- which(duplicated(pres_abs))
duplicates_check 

# Rename rows
row.names(pres_abs) <- pres_abs$names

# Order species names by alphabetical order
pres_abs <- pres_abs[order(row.names(pres_abs)),]

# Reorder lake names
# Define the desired column order
desired_order <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "LCN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOO", "OTM", "OOM", "RCA", "REF", "SLN", "TLN", "ULN")

# Reorder the columns based on the desired order
pres_abs <- pres_abs[, desired_order]

# Transpose matrix to generate lake-by-species matrix
pres_abs_t <- as.data.frame(t(pres_abs))
```

## Community type incidence matrices
- Create a presence absence matrix for each community/stratification type
```{r, Create presence absence matrices for each type}
keep <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "LCN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOO", "OTM", "OOM", "RCA", "SLN", "TLN", "ULN")
surveyed_site_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
surveyed_site_fish <- surveyed_site_fish[which(rowSums(surveyed_site_fish) > 0),]


keep <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "LLN", "MLN", "NLK", "NLN", "NLU", "OLO", "OTM", "SLN", "TLN", "ULN")
marine_lake_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
marine_lake_fish <- marine_lake_fish[which(rowSums(marine_lake_fish) > 0),]


keep <- c("IBK", "LCN", "NCN", "OOO", "OOM", "RCA")
ocean_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
ocean_fish <- ocean_fish[which(rowSums(ocean_fish) > 0),]

keep <- c("FLK", "HLM", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")
holomictic_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
holomictic_fish <- holomictic_fish[which(rowSums(holomictic_fish) > 0),]

keep <- c("BCM", "CLM", "GLK", "NLK", "OTM", "SLN", "TLN")
meromictic_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
meromictic_fish <- meromictic_fish[which(rowSums(meromictic_fish) > 0),]

keep <- c("FLK", "HLO", "LLN", "MLN", "NLN", "NLU", "OLO", "ULN")
mixed_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
mixed_fish <- mixed_fish[which(rowSums(mixed_fish) > 0),]

keep <- c("BCM", "CLM", "GLK", "HLM", "NLK", "OTM", "SLN", "TLN")
stratified_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
stratified_fish <- stratified_fish[which(rowSums(stratified_fish) > 0),]


# Create extra column to be able to extract out lake specific incidences
pres_abs$zero <- 0

keep <- c("BCM", "zero")
BCM_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
BCM_fish <- BCM_fish[which(rowSums(BCM_fish) > 0),]
BCM_fish <- BCM_fish["BCM"]

keep <- c("CLM", "zero")
CLM_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
CLM_fish <- CLM_fish[which(rowSums(CLM_fish) > 0),]
CLM_fish <- CLM_fish["CLM"]

keep <- c("FLK", "zero")
FLK_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
FLK_fish <- FLK_fish[which(rowSums(FLK_fish) > 0),]
FLK_fish <- FLK_fish["FLK"]

keep <- c("GLK", "zero")
GLK_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
GLK_fish <- GLK_fish[which(rowSums(GLK_fish) > 0),]
GLK_fish <- GLK_fish["GLK"]

keep <- c("HLM", "zero")
HLM_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
HLM_fish <- HLM_fish[which(rowSums(HLM_fish) > 0),]
HLM_fish <- HLM_fish["HLM"]

keep <- c("HLO", "zero")
HLO_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
HLO_fish <- HLO_fish[which(rowSums(HLO_fish) > 0),]
HLO_fish <- HLO_fish["HLO"]

keep <- c("IBK", "zero")
IBK_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
IBK_fish <- IBK_fish[which(rowSums(IBK_fish) > 0),]
IBK_fish <- IBK_fish["IBK"]

keep <- c("LLN", "zero")
LLN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
LLN_fish <- LLN_fish[which(rowSums(LLN_fish) > 0),]
LLN_fish <- LLN_fish["LLN"]

keep <- c("LCN", "zero")
LCN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
LCN_fish <- LCN_fish[which(rowSums(LCN_fish) > 0),]
LCN_fish <- LCN_fish["LCN"]

keep <- c("MLN", "zero")
MLN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
MLN_fish <- MLN_fish[which(rowSums(MLN_fish) > 0),]
MLN_fish <- MLN_fish["MLN"]

keep <- c("NCN", "zero")
NCN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
NCN_fish <- NCN_fish[which(rowSums(NCN_fish) > 0),]
NCN_fish <- NCN_fish["NCN"]

keep <- c("NLK", "zero")
NLK_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
NLK_fish <- NLK_fish[which(rowSums(NLK_fish) > 0),]
NLK_fish <- NLK_fish["NLK"]

keep <- c("NLN", "zero")
NLN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
NLN_fish <- NLN_fish[which(rowSums(NLN_fish) > 0),]
NLN_fish <- NLN_fish["NLN"]

keep <- c("NLU", "zero")
NLU_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
NLU_fish <- NLU_fish[which(rowSums(NLU_fish) > 0),]
NLU_fish <- NLU_fish["NLU"]

keep <- c("OLO", "zero")
OLO_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
OLO_fish <- OLO_fish[which(rowSums(OLO_fish) > 0),]
OLO_fish <- OLO_fish["OLO"]

keep <- c("OOO", "zero")
OOO_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
OOO_fish <- OOO_fish[which(rowSums(OOO_fish) > 0),]
OOO_fish <- OOO_fish["OOO"]

keep <- c("OTM", "zero")
OTM_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
OTM_fish <- OTM_fish[which(rowSums(OTM_fish) > 0),]
OTM_fish <- OTM_fish["OTM"]

keep <- c("OOM", "zero")
OOM_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
OOM_fish <- OOM_fish[which(rowSums(OOM_fish) > 0),]
OOM_fish <- OOM_fish["OOM"]

keep <- c("RCA", "zero")
RCA_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
RCA_fish <- RCA_fish[which(rowSums(RCA_fish) > 0),]
RCA_fish <- RCA_fish["RCA"]

keep <- c("SLN", "zero")
SLN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
SLN_fish <- SLN_fish[which(rowSums(SLN_fish) > 0),]
SLN_fish <- SLN_fish["SLN"]

keep <- c("TLN", "zero")
TLN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
TLN_fish <- TLN_fish[which(rowSums(TLN_fish) > 0),]
TLN_fish <- TLN_fish["TLN"]

keep <- c("ULN", "zero")
ULN_fish <- pres_abs[,keep]

## Remove all species not found in any locations and duplicates
# Identifies which rows are greater than 0
# Avoid the species column when using this command
ULN_fish <- ULN_fish[which(rowSums(ULN_fish) > 0),]
ULN_fish <- ULN_fish["ULN"]

keep <- c("BCM", "CLM", "FLK", "GLK", "HLM", "HLO", "IBK", "LLN", "LCN", "MLN", "NCN", "NLK", "NLN", "NLU", "OLO", "OOO", "OTM", "OOM", "RCA", "REF", "SLN", "TLN", "ULN")
pres_abs <- pres_abs[,keep]
```

## Load out files
```{r,Load out pesabs files and session info}
# Fish presence by lake
write.csv(pres_abs_t, "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/fish_presence_matrix_by_lake.csv")

# Fish presence by species
write.csv(pres_abs,  "/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/fish_presence_matrix_by_species.csv")

# Marine lake fish presence by species 
write.csv(surveyed_site_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/surveyed_site_presence_by_species.csv")

# Marine lake fish presence by species 
write.csv(marine_lake_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/marine_lake_presence_by_species.csv")

# Ocean site fish presence by species 
write.csv(ocean_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/ocean_site_presence_by_species.csv")

# Holomictic fish presence by species 
write.csv(holomictic_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/holomictic_presence_by_species.csv")

# Meromictic fish presence by species 
write.csv(meromictic_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/meromictic_presence_by_species.csv")

# Mixed fish presence by species 
write.csv(mixed_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/mixed_presence_by_species.csv")

# Stratified fish presence by species 
write.csv(stratified_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/stratified_presence_by_species.csv")

# Surveyed sites fish presence by species each in separate files
write.csv(BCM_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/BCM_fish.csv")
write.csv(CLM_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/CLM_fish.csv")
write.csv(FLK_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/FLK_fish.csv")
write.csv(GLK_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/GLK_fish.csv")
write.csv(HLM_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/HLM_fish.csv")
write.csv(HLO_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/HLO_fish.csv")
write.csv(IBK_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/IBK_fish.csv")
write.csv(LLN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/LLN_fish.csv")
write.csv(LCN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/LCN_fish.csv")
write.csv(MLN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/MLN_fish.csv")
write.csv(NCN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/NCN_fish.csv")
write.csv(NLK_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/NLK_fish.csv")
write.csv(NLN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/NLN_fish.csv")
write.csv(NLU_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/NLU_fish.csv")
write.csv(OLO_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/OLO_fish.csv")
write.csv(OOO_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/OOO_fish.csv")
write.csv(OTM_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/OTM_fish.csv")
write.csv(OOM_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/OOM_fish.csv")
write.csv(RCA_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/RCA_fish.csv")
write.csv(SLN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/SLN_fish.csv")
write.csv(TLN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/TLN_fish.csv")
write.csv(ULN_fish,"/Users/bailey/Documents/research/fish_biodiversity/data/collection/fish/pres_abs/ULN_fish.csv")

sessionInfo()
```
