---
title: "Trait plots"
output: github_document
editor_options: 
  chunk_output_type: console
---
### Load packages
```{r, Load Location_type weighted mean trait plots packages}
library(dplyr)
library(ggplot2)
library(viridis)
library(dunn.test)
library(car)

# Load the knitr package if not already loaded
library(knitr)

# Source the R Markdown file
knit("/Users/bailey/Documents/research/Fish_Biodiversity/src/collection/load_collection_data.Rmd", output = "/Users/bailey/Documents/research/Fish_Biodiversity/src/collection/load_collection_data.md")

# Define your custom colors
custom_shades <- c("R" = "black", "O" = "#EE6363", "M" = "#87CEFA", "S" = "#6E8B3D")

# Reference use Location_type_at_weighted
# Surveyed sites use Location_type_st_weighted
```

## Location trait data tests
```{r, Location trait data tests}
# Create function
run_tests <- function(trait, data) {
    # Ensure the trait is numeric
    if (!is.numeric(data[[trait]])) {
        return(c(ANOVA_p = NA, Shapiro_p = NA, Levene_p = NA))}  # Return NA if not numeric
    formula <- as.formula(paste(trait, "~ Location_type"))  # Dynamically create formula
    anova_model <- aov(formula, data = data)  # Run ANOVA
    anova_p <- summary(anova_model)[[1]][["Pr(>F)"]][1]  # Extract p-value
    # Run Shapiro-Wilk test on residuals (normality assumption)
    shapiro_p <- tryCatch(
        shapiro.test(residuals(anova_model))$p.value,
        error = function(e) NA)  # Return NA if Shapiro fails
    # Run Levene's test for homogeneity of variance
    levene_p <- tryCatch(
        leveneTest(formula, data = data)$`Pr(>F)`[1],
        error = function(e) NA)  # Return NA if Levene fails
    return(c(ANOVA_p = anova_p, Shapiro_p = shapiro_p, Levene_p = levene_p))
}

# Run tests on numeric traits
traits <- colnames(Locations_at)[1:15]  # Extract trait names
results <- sapply(traits, run_tests, data = Locations_at)

# Convert to a tidy dataframe
results_df <- as.data.frame(t(results))
results_df$Trait <- rownames(results_df)
rownames(results_df) <- NULL

# Print results
print(results_df)


# Identify traits that are numeric
numerical <- c("MaxLengthTL", "Troph", "DepthMin", "DepthMax", "TempPrefMin", "TempPrefMax", "DorsalSpinesMean")

## MaxLengthTL
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$MaxLengthTL, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)


## Troph
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$Troph, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
median(Ocean_at$Troph)
median(Mixed_at$Troph)
median(Stratified_at$Troph)

## DepthMin
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$DepthMin, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
median(Ocean_at$DepthMin)
median(Mixed_at$DepthMin)
median(Stratified_at$DepthMin)

## DepthMax
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$DepthMax, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)


## TempPrefMin
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$TempPrefMin, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)


## TempPrefMax
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$TempPrefMax, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)


## DorsalSpinesMean
# Perform Dunn's test
dunn_result <- dunn.test(Locations_at$DorsalSpinesMean, Locations_at$Location_type, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)


## # Identify traits that are factors
factor <- c("BodyShapeI", "WaterPref", "DemersPelag", "OperculumPresent", "FeedingPath", "RepGuild2", "ParentalCare")

## BodyShapeI
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$BodyShapeI)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)


## WaterPref
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$WaterPref)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)


## DemersPelag
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$DemersPelag)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)


## OperculumPresent
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$OperculumPresent)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)


## FeedingPath
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$FeedingPath)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)


## RepGuild2
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$RepGuild2)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)


## ParentalCare
# Convert to contingency table
contingency_table <- table(Locations_at$Location_type, Locations_at$ParentalCare)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]
# Perform Fisher's test
fisher_test_result <- fisher.test(contingency_table_clean, simulate.p.value=TRUE)  # Adjust workspace size
print(fisher_test_result)
```

## Biotic traits
### Trophic level
```{r, Trophic level}
Troph_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= Troph, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 2,
            alpha = 0.6,
            width = 0.1) +  
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 1, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 12), 
        legend.text = element_text(size = 10),
        axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Trophic Level") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "a")
Troph_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/Troph_plot_weighted.jpg", Troph_plot_weighted, width = 3.25, height = 2.95, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(Troph ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$Troph, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

### Reproductive Guild2
```{r, Reproductive Guild2}
Location_type_st_props_RepGuild2 <- Location_type_st %>%
  group_by(Strat, RepGuild2) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

Location_type_st_props_RepGuild2 <- na.omit(Location_type_st_props_RepGuild2)

RepGuild2_plot_weighted <- ggplot(Location_type_st_props_RepGuild2, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = RepGuild2)) + 
  geom_point(position = "identity", size = 4, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = RepGuild2), linewidth = 1, linetype = "dotted") +
  scale_shape_manual(name = "Egg Strategy:", values = c(11, 21:25), labels = c('1ib' = 'bearers', '6s' = 'scatterers', '3n' = 'nesters', '5h' = 'hiders', '4t' = 'tenders', '2eb' = 'brooders')) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    legend.title = element_text(size = 8),  # Adjust legend title size
    legend.text = element_text(size = 8),  # Adjust legend text size
    legend.key.size = unit(0, "cm"),  # Adjust legend key size
    legend.position = "bottom",
    legend.spacing = unit(0, "cm"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none", shape = guide_legend(nrow = 2, byrow = TRUE, title.position = "top")) +
  ylim(c(0,0.6)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "c")
RepGuild2_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/RepGuild2_plot_weighted.jpg", RepGuild2_plot_weighted, width = 3.25, height = 3.81, units = "in")

# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$RepGuild2)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

### Dorsal Spines Mean
```{r, Dorsal Spines Mean}
DorsalSpinesMean_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= DorsalSpinesMean, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 2,
            alpha = 0.6,
            width = 0.1) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 1, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 12), 
        legend.text = element_text(size = 10),
        axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Dorsal Spines") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "b")
DorsalSpinesMean_plot_weighted
# Warning messages:
# 1: Removed 10 rows containing non-finite values (`stat_ydensity()`). 
# 2: Removed 10 rows containing missing values (`geom_point()`). 
# Twelve values with NA
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/DorsalSpinesMean_plot_weighted.jpg", DorsalSpinesMean_plot_weighted, width = 3.25, height = 2.95, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(DorsalSpinesMean ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$DorsalSpinesMean, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

### Parental Care
```{r, Parental Care}
Location_type_st_props_ParentalCare <- Location_type_st %>%
  group_by(Strat, ParentalCare) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

Location_type_st_props_ParentalCare <- na.omit(Location_type_st_props_ParentalCare)

ParentalCare_plot_weighted <- ggplot(Location_type_st_props_ParentalCare, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = ParentalCare)) + 
  geom_point(position = "identity", size = 4, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = ParentalCare), linewidth = 1, linetype = "dotted") +
  scale_shape_manual(name = "Parental Care:", values = c(21:25), labels = c('4n' = 'none', '3p' = 'paternal', '2m' = 'maternal', '1b' = 'biparental')) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    legend.title = element_text(size = 8),  # Adjust legend title size
    legend.text = element_text(size = 8),  # Adjust legend text size
    legend.key.size = unit(0, "cm"),  # Adjust legend key size
    legend.position = "bottom",
    legend.spacing = unit(0, "cm"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none", shape = guide_legend(nrow = 2, byrow = TRUE, title.position = "top")) +
  ylim(c(0,0.6)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "d")
ParentalCare_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/ParentalCare_plot_weighted.jpg", ParentalCare_plot_weighted, width = 3.25, height = 3.81, units = "in")


# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$ParentalCare)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

## Abiotic traits
### Temperature Preference Minimum
```{r, Temperature Preference Minimum}
TempPrefMin_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= TempPrefMin, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 2,
            alpha = 0.6,
            width = 0.1) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 1, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 12), 
        legend.text = element_text(size = 10),
        axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Temp Min (Cº)") +
  scale_y_continuous(breaks = c(20,22,24,26,28)) +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "a")
TempPrefMin_plot_weighted
# Warning messages:
# 1: Removed 9 rows containing non-finite values (`stat_ydensity()`). 
# 2: Removed 9 rows containing missing values (`geom_point()`). 
# Two values with NA
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/TempPrefMin_plot_weighted.jpg", TempPrefMin_plot_weighted, width = 3.25, height = 2.95, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(TempPrefMin ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$TempPrefMin, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

### Water
- Fresh/Brack/Salt
```{r, Water}
Location_type_st_props_Water <- Location_type_st %>%
  group_by(Strat, WaterPref) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

#Location_type_st_props_Water$Water <- factor(Location_type_st_props_Water$Water, levels = c("all", "fresh", "fresh-brack", "brack", "brack-salt", "salt"))

Water_plot_weighted <- ggplot(Location_type_st_props_Water, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = WaterPref)) + 
  geom_point(position = "identity", size = 4, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = WaterPref), linewidth = 1, linetype = "dotted") +
  scale_shape_manual(name = "Water:", values = c(21:23,11,24:25), labels = c('3a' = 'all', '1s' = 'salt', '2bs' = 'brackish-salt', '4b' = 'brack', '5fb' = 'fresh-brackish', '6f' = 'fresh')) + 
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    legend.title = element_text(size = 8),  # Adjust legend title size
    legend.text = element_text(size = 8),  # Adjust legend text size
    legend.key.size = unit(0, "cm"),  # Adjust legend key size
    legend.position = "bottom",
    legend.spacing = unit(0, "cm"),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none", shape = guide_legend(nrow = 2, byrow = TRUE, title.position = "top")) +
  ylab("Proportion") +
  scale_y_continuous(breaks = c(0,0.5,1), limits = c(0,1)) +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "c")
Water_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/Water_plot_weighted.jpg", Water_plot_weighted, width = 3.25, height = 3.81, units = "in")


# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$WaterPref)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

### Temperature Preference Maximum
```{r, Temperature Preference Maximum}
TempPrefMax_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= TempPrefMax, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 2,
            alpha = 0.6,
            width = 0.1) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 1, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 12), 
        legend.text = element_text(size = 10),
        axis.title = element_text(size = 10),
    axis.text = element_text(size = 10, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Temp Max (Cº)") +
  scale_y_continuous(breaks = c(27,28,29,30)) +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "b")
TempPrefMax_plot_weighted
# Warning messages:
# 1: Removed 11 rows containing non-finite values (`stat_ydensity()`). 
# 2: Removed 11 rows containing missing values (`geom_point()`). 
# Two values with NA
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/TempPrefMax_plot_weighted.jpg", TempPrefMax_plot_weighted, width = 3.25, height = 2.95, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(TempPrefMax ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$TempPrefMax, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

## Supplementary traits
### Operculum Present
```{r, Operculum Present}
Location_type_st_props_OperculumPresent <- Location_type_st %>%
  group_by(Strat, OperculumPresent) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

OperculumPresent_plot_weighted <- ggplot(Location_type_st_props_OperculumPresent, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = OperculumPresent)) + 
  geom_point(position = "identity", size = 10, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = OperculumPresent), linewidth = 2, linetype = "dotted") +
  scale_shape_manual(name = "Operculum:", values = c(21:22)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 30),
    legend.title = element_text(size = 30),  # Adjust legend title size
    legend.text = element_text(size = 30),  # Adjust legend text size
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none") + # Remove the legends for Location type
  ylim(c(0,0.75)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "a")
OperculumPresent_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/OperculumPresent_plot_weighted.jpg", OperculumPresent_plot_weighted,  width = 13, height = 10, units = "in")


# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$OperculumPresent)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

### Max Length (TL)
```{r, Max Length (TL)}
MaxLengthTL_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= MaxLengthTL, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 4,
            alpha = 0.6,
            width = 0.1) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 2, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 30), legend.text = element_text(size = 30),
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Length (cm)") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "b")
MaxLengthTL_plot_weighted
# Warning messages:
# 1: Removed 3 rows containing non-finite values (`stat_ydensity()`). 
# 2: Removed 3 rows containing missing values (`geom_point()`). 
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/MaxLengthTL_plot_weighted.jpg", MaxLengthTL_plot_weighted,  width = 13, height = 10, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(MaxLengthTL ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$MaxLengthTL, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

### Body Shape
```{r, Body Shape}
Location_type_st_props_BodyShape <- Location_type_st %>%
  group_by(Strat, BodyShapeI) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

BodyShape_plot_weighted <- ggplot(Location_type_st_props_BodyShape, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = BodyShapeI)) + 
  geom_point(position = "identity", size = 10, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = BodyShapeI), linewidth = 2, linetype = "dotted") +
  scale_shape_manual(name = "Body Shape:", values = c(11, 21:25), labels = c("1o" = "other", "2s" = "short deep", "3f" = "fusiform", "4e" = "elongated", "5l" = "eel-like")) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 30),
    legend.title = element_text(size = 30),  # Adjust legend title size
    legend.text = element_text(size = 30),  # Adjust legend text size
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none") + # Remove the legends for Location type
  ylim(c(0,0.6)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "c")
BodyShape_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/BodyShape_plot_weighted.jpg", BodyShape_plot_weighted,  width = 13, height = 10, units = "in")


# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$BodyShapeI)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

### Depth Maximum
```{r, Depth Maximum}
DepthMax_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= DepthMax, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 4,
            alpha = 0.6,
            width = 0.1) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 2, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 30), legend.text = element_text(size = 30),
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Depth Max (m)") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "d")
DepthMax_plot_weighted
# Warning messages:
# 1: Removed 10 rows containing non-finite values (`stat_ydensity()`). 
# 2: Removed 10 rows containing missing values (`geom_point()`). 
# One value with NA
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/DepthMax_plot_weighted.jpg", DepthMax_plot_weighted,  width = 13, height = 10, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(DepthMax ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$DepthMax, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

### Feeding Pathway
```{r, Feeding Pathway}
Location_type_st_props_FeedingPath <- Location_type_st %>%
  group_by(Strat, FeedingPath) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

Location_type_st_props_FeedingPath <- na.omit(Location_type_st_props_FeedingPath)

FeedingPath_plot_weighted <- ggplot(Location_type_st_props_FeedingPath, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = FeedingPath)) + 
  geom_point(position = "identity", size = 10, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = FeedingPath), linewidth = 2, linetype = "dotted") +
  scale_shape_manual(name = "Diet Source:", values = c(21:22), labels = c("b" = "benthic", "p" = "pelagic")) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 30),
    legend.title = element_text(size = 30),  # Adjust legend title size
    legend.text = element_text(size = 30),  # Adjust legend text size
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none") + # Remove the legends for Location type
  ylim(c(0,1)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "e")
FeedingPath_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/FeedingPath_plot_weighted.jpg", FeedingPath_plot_weighted,  width = 13, height = 10, units = "in")


# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$FeedingPath)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

### Depth Minimum
```{r, Depth Minimum}
DepthMin_plot_weighted <- ggplot(Location_type_st_weighted, mapping = aes(x= Strat, y= DepthMin, fill = Strat)) +
  geom_jitter(shape = 21,
            size = 4,
            alpha = 0.6,
            width = 0.1) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75), linewidth = 2, aes(group = Strat, fill = Strat)) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  guides(fill = "none") +
  theme_bw() +
  theme(text = element_text(size = 30), legend.text = element_text(size = 30),
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()) + 
  ylab("Depth Min (m)") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "f")
DepthMin_plot_weighted
# Warning messages:
# 1: Removed 8 rows containing non-finite values (`stat_ydensity()`). 
# 2: Removed 8 rows containing missing values (`geom_point()`). 
# One value with NA
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/DepthMin_plot_weighted.jpg", DepthMin_plot_weighted,  width = 13, height = 10, units = "in")


# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(DepthMin ~ Strat, data = Location_type_st_weighted)
# Print result
print(kruskal_result)
# Perform Dunn's test
dunn_result <- dunn.test(Location_type_st_weighted$DepthMin, Location_type_st_weighted$Strat, method = "Bonferroni")
# Print Dunn's test result
print(dunn_result)
```

### DemersPelag
```{r, DemersPelag}
Location_type_st_props_DemersPelag <- Location_type_st %>%
  group_by(Strat, DemersPelag) %>%
  summarise(count = sum(LocationSums)) %>%
  mutate(proportion = count/sum(count))

DemersPelag_plot_weighted <- ggplot(Location_type_st_props_DemersPelag, mapping = aes(x= Strat, y= proportion, fill = Strat, shape = DemersPelag)) + 
  geom_point(position = "identity", size = 10, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = DemersPelag), linewidth = 2, linetype = "dotted") +
  scale_shape_manual(name = "Demersal Pelagic:", values = c(21:25, 12:13), 
                     labels = c("1r" = "reef-associated", "2pn" = "pelagic-neritic", "3p" = "pelagic", "4po" = "pelagic-oceanic", "5d" = "demersal", '6bp' = 'benthopelagic', '7bd' = 'bathydemersal')) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 30),
    legend.title = element_text(size = 30),  # Adjust legend title size
    legend.text = element_text(size = 30),  # Adjust legend text size
    axis.text = element_text(size = 30, color = "black"),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none") + # Remove the legends for Location type
  ylim(c(0,1)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(fill = "Location type:", tag = "g")
DemersPelag_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/DemersPelag_plot_weighted.jpg", DemersPelag_plot_weighted,  width = 13, height = 10, units = "in")


# Convert to contingency table
contingency_table <- table(Location_type_st_weighted$Strat, Location_type_st_weighted$DemersPelag)
contingency_table_clean <- contingency_table[rowSums(contingency_table) > 0, colSums(contingency_table) > 0]

fisher_test_result <- fisher.test(contingency_table_clean, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result)

# Subset for groups O and M
pairwise_OM <- contingency_table_clean[c("O", "M"), ]
fisher_test_result_OM <- fisher.test(pairwise_OM, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OM)

# Subset for groups O and S
pairwise_OS <- contingency_table_clean[c("O", "S"), ]
fisher_test_result_OS <- fisher.test(pairwise_OS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_OS)

# Subset for groups M and S
pairwise_MS <- contingency_table_clean[c("M", "S"), ]
fisher_test_result_MS <- fisher.test(pairwise_MS, workspace = 2e8)  # Adjust workspace size
print(fisher_test_result_MS)

# Adjust for multiple comparisons
p_values <- c(fisher_test_result_OM$p.value, fisher_test_result_OS$p.value, fisher_test_result_MS$p.value)
adjusted_p_values <- p.adjust(p_values, method = "bonferroni")
print(adjusted_p_values)
```

### Reproductive Guild1
```{r, Reproductive Guild1}
Location_type_st_props_RepGuild1 <- Location_type_st %>%
  group_by(Strat, RepGuild1) %>%
  summarise(count = sum(LocationSums)) %>%
  group_by(Strat) %>%
  mutate(proportion = count/sum(count))

Location_type_st_props_RepGuild1 <- na.omit(Location_type_st_props_RepGuild1)

RepGuild1_plot_weighted <- ggplot(Location_type_st_props_RepGuild1, mapping = aes(x = Strat, y = proportion, shape = RepGuild1)) + 
  geom_point(position = "identity", size = 10, aes(group = Strat, fill = Strat)) +
  geom_line(position = "identity", aes(group = RepGuild1), linewidth = 2, linetype = "dotted") +
  scale_shape_manual(name = "Egg Care:", values = c(21, 22, 23), labels = c('2g' = 'guarders', '1b' = 'bearers', '3n' = 'nonguarders')) +
  # scale_color_manual(values = custom_shades) +
  scale_fill_manual(values = custom_shades) +
  theme_bw() +
  theme(
    text = element_text(size = 30),
    legend.title = element_text(size = 30),  # Adjust legend title size
    legend.text = element_text(size = 30),  # Adjust legend text size
    axis.text = element_text(size = 30, color = "black"),
    axis.line = element_line(color = "black"),
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill = "none") + # Remove the legends for Location type
  ylim(c(0, 0.6)) +
  ylab("Proportion") +
  xlab("Location type") +
  labs(tag = "none")

RepGuild1_plot_weighted
ggsave("/Users/bailey/Documents/research/Fish_Biodiversity/figures/FD/trait_plots/RepGuild1_plot_weighted.jpg", RepGuild1_plot_weighted,  width = 13, height = 10, units = "in")
```

```{r, Location_type abundance weighted mean trait plots session info}
sessionInfo()
```
